---
title: "GCRF_GWAS_Workflow"
output:
  html_document:
    toc: true
    toc_float: true
date: "2024-01-04"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Desktop/GCRF/")
```

# INTRODUCTION
Here I will be going through the steps for running a GWAS on a variety of morphological traits. I'm starting with vcf files that have been filtered as follows
-only biallilic alleles
-only SNPs
-fraction of missing genotypes is <25%
-min allele freq is 0.05
-max allele freq is .95
-depth greater than or equal to 4

# WORKFLOW
## 1. VQSR filtering
This filtering is to remove errors in genotypes due to sequencing issues. The script I was given actually wants this to be run on an already merged vcf file, but when I tried this the filtering went fine but downstream the imputation kept failing due to a weird error. So I regrouped and did the filtering on each scaffold individually. This was implemented with the vqsr_filt.sh script, which had the following filters:

-filter "QUAL < 30.0" --filter-name "QUAL30" \
-filter "SOR > 3.0" --filter-name "SOR3" \
-filter "FS > 60.0" --filter-name "FS60" \
-filter "MQ < 40.0" --filter-name "MQ40" \
-filter "QD < 2.0" --filter-name "QD2" \
-filter "MQRankSum < -12.5" --filter-name "MQRankSum-12.5" \
-filter "ReadPosRankSum < -8.0" --filter-name "ReadPosRankSum-8"

The last two, MQRankSum and ReadPosRankSum didn't seem to apply, the error was something along the lines of the information not being present which didn't make sense since the header was there, but I opted to move on.

Once all the scaffolds were filtered, I moved the files to a new folder for imputation.

## 2. BEAGLE 4.1 Imputation
I merged the vqsr filtered files into a single GCRF_merged file. Then I applied BEAGLE4.1 to impute missing genotypes. This was done with the script impute.sh.

```{}
BEAGLE_JAR="/projects/ericacnr@colostate.edu/mambaforge/envs/bioinf/java/beagle.27Jan18.7e1.jar"
j=GCRF_merged.vcf.gz

java -Xmx180G -jar $BEAGLE_JAR \
gt="$j" \
out=GCRF_merged_imputed4.1 \
nthreads=48
```

The log had this as the final output: 
Number of markers:             7942760
Total time for building model: 2 hours 18 minutes 11 seconds
Total time for sampling:       35 minutes 10 seconds
Total run time:                3 hours 3 minutes 55 seconds

Which suggests it was running the correct number of SNPs (7942760) however upon inspecting the file there was over 9mil lines, which doesn't make sense and suggested that some of the markers (~2mil) were being duplicated. Additionally, there was an error with the imputed files where the Chromosome position was missing in the header. So to fix that issue first, I cropped the header from the original merged file and added it to the imputed file.

```{}
(bioinf) [ericacnr@colostate.edu@login12 impute]$ zcat GCRF_merged_impute4.1.vcf.gz |grep -v "^#" > GCRF_merged_impute4.1.vcf_nohead
(bioinf) [ericacnr@colostate.edu@login12 impute]$ zcat GCRF_merged.vcf.gz | grep "^#" > header.txt
(bioinf) [ericacnr@colostate.edu@login12 impute]$ cat GCRF_merged_impute4.1.vcf_nohead | wc -l
9943613
(bioinf) [ericacnr@colostate.edu@login12 impute]$ cat header.txt GCRF_merged_impute4.1.vcf_nohead | bgzip -c > GCRF_merged_imputed4.1_newhead.vcf.gz
```

I then attempted to apply the bcftools norms function on this corrected file...

```{}
(bioinf) [ericacnr@colostate.edu@login12 impute]$ bcftools norm -d none -f ../../Reference/leucosticte_australis_final_assembly.fasta  GCRF_merged_imputed4.1_newhead.vcf.gz > GCRF_merged_impute4.1.rm_dup.vcf.gz
Lines   total/split/joined/realigned/skipped:	9943613/0/0/0/0
(bioinf) [ericacnr@colostate.edu@login12 impute]$ cat GCRF_merged_impute4.1.rm_dup.vcf.gz | grep -v "^#" | wc -l
9943500
```

This still left me with too many lines, but I tried to apply plink anyway... it threw an error though, something about split chromosomes?

```{}
(GWAS) [ericacnr@colostate.edu@login12 impute]$ plink --vcf GCRF_merged_imputed4.1_newhead.vcf.gz --aec --recode --out GCRF_newhead
PLINK v1.90b6.21 64-bit (19 Oct 2020)          www.cog-genomics.org/plink/1.9/
(C) 2005-2020 Shaun Purcell, Christopher Chang   GNU General Public License v3
Logging to GCRF_newhead.log.
Options in effect:
  --allow-extra-chr
  --out GCRF_newhead
  --recode
  --vcf GCRF_merged_imputed4.1_newhead.vcf.gz

15884 MB RAM detected; reserving 7942 MB for main workspace.
--vcf: GCRF_newhead-temporary.bed + GCRF_newhead-temporary.bim +
GCRF_newhead-temporary.fam written.
Error: .bim file has a split chromosome.  Use --make-bed by itself to
remedy this.
```

So I tried sorting the vcf file.

```{}
(bioinf) [ericacnr@colostate.edu@login12 impute]$ bcftools sort GCRF_merged_imputed4.1_newhead.vcf.gz > GCRF_merged_imputed4.1_newhead.sort.vcf
Writing to /tmp/bcftools.zBleDc
Merging 7 temporary files
Cleaning
Done
```

This sorting works! So doing the bcftools norm again on the sorted file.

```{}
(GWAS) [ericacnr@colostate.edu@login12 impute]$ conda activate bioinf
(bioinf) [ericacnr@colostate.edu@login12 impute]$ bcftools norm -d none -f ../../Reference/leucosticte_australis_final_assembly.fasta  GCRF_merged_imputed4.1_newhead.sort.vcf > GCRF_merged_impute4.1.sort.rm_dup.vcf.gz
Lines   total/split/joined/realigned/skipped:	9943613/0/0/0/0
(bioinf) [ericacnr@colostate.edu@login12 impute]$ cat GCRF_merged_impute4.1.sort.rm_dup.vcf.gz | grep -v "^#" |wc -l
7942760
(bioinf) [ericacnr@colostate.edu@login12 impute]$ mv GCRF_merged_impute4.1.sort.rm_dup.vcf.gz GCRF_merged_impute4.1.sort.rm_dup.vcf
```

## 3. PLINK to make .ped and .map files

This worked and the number of SNPs are replaced. I renamed it because I messed that up. Now trying plink again!

```{}
(bioinf) [ericacnr@colostate.edu@login12 impute]$ conda activate GWAS
(GWAS) [ericacnr@colostate.edu@login12 impute]$ plink --vcf GCRF_merged_impute4.1.sort.rm_dup.vcf --aec --recode --out ../input_files/GCRF_base_file
PLINK v1.90b6.21 64-bit (19 Oct 2020)          www.cog-genomics.org/plink/1.9/
(C) 2005-2020 Shaun Purcell, Christopher Chang   GNU General Public License v3
Logging to ../input_files/GCRF_base_file.log.
Options in effect:
  --allow-extra-chr
  --out ../input_files/GCRF_base_file
  --recode
  --vcf GCRF_merged_impute4.1.sort.rm_dup.vcf

15884 MB RAM detected; reserving 7942 MB for main workspace.
--vcf: ../input_files/GCRF_base_file-temporary.bed +
../input_files/GCRF_base_file-temporary.bim +
../input_files/GCRF_base_file-temporary.fam written.
7942760 variants loaded from .bim file.
89 people (0 males, 0 females, 89 ambiguous) loaded from .fam.
Ambiguous sex IDs written to ../input_files/GCRF_base_file.nosex .
Using 1 thread (no multithreaded calculations invoked).
Before main variant filters, 89 founders and 0 nonfounders present.
Calculating allele frequencies... done.
7942760 variants and 89 people pass filters and QC.
Note: No phenotypes present.
--recode ped to ../input_files/GCRF_base_file.ped +
../input_files/GCRF_base_file.map ... done.
```

## 4. Fixing the .map files and the .ped files to have morphology

Fixing the .map file so that the SNP name is the scaffold and the position together. This file shows the chromosome and SNP location for each SNP.

```{}
(GWAS) [ericacnr@colostate.edu@login12 impute]$ cd ../input_files/
(GWAS) [ericacnr@colostate.edu@login12 input_files]$ cat GCRF_base_file.map | head
Scaffold_1__1_contigs__length_151072562	.	0	91
Scaffold_1__1_contigs__length_151072562	.	0	303
Scaffold_1__1_contigs__length_151072562	.	0	323
Scaffold_1__1_contigs__length_151072562	.	0	348
Scaffold_1__1_contigs__length_151072562	.	0	460
Scaffold_1__1_contigs__length_151072562	.	0	473
Scaffold_1__1_contigs__length_151072562	.	0	520
Scaffold_1__1_contigs__length_151072562	.	0	751
Scaffold_1__1_contigs__length_151072562	.	0	821
Scaffold_1__1_contigs__length_151072562	.	0	1271
(GWAS) [ericacnr@colostate.edu@login12 input_files]$ cat GCRF_base_file.map | awk '{print$1"\t"$1"."$4"\t"$3"\t"$4}' > GCRF_base_file.map_loc.map
(GWAS) [ericacnr@colostate.edu@login12 input_files]$ cat GCRF_base_file.map_loc.map | head
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.91	0	91
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.303	0	303
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.323	0	323
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.348	0	348
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.460	0	460
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.473	0	473
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.520	0	520
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.751	0	751
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.821	0	821
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.1271	0	1271
```

The .ped file is the FID, IID, (other stuff 3-5), phenotype data, genotypes. I need to replace the first column in the .ped file to have the FID be the location. So the plan for this is to make a separate .txt file in R with the location and the sample names, then add it to the original ped file to replace the first two columns.

According to Marina: "In your ped file, this means the first columns should be all "1"s, the second column should be your sample ID, then 3 columns which are usually "0"s, and finally, the 6th column is your phenotype column"

I'm going to keep the FID as WMNT=1 and PIPA=2 so that there's some distinction, and see if this will run in GEMMA.

First lets make the .txt file, we'll do that here.
```{r}
morpho <- read.csv("~/Desktop/GCRF/R data/GCRF_morphology.csv")
morpho <- morpho[1:189,]
locID <- as.data.frame(cbind(morpho$site, morpho$BGP_ID))
colnames(locID) <- c("site", "BGP_ID")

#getting the correct sample IDs
locID <- distinct(locID)
locID <- locID[order(locID$BGP_ID), ]
rownames(locID) <- NULL
locID <- locID[38:128,]
rownames(locID) <- NULL
locID <- locID[c(-15,-18),]

#fixing the site data to be numerical
locID <- replace(locID, locID == "WMNT", 1)
locID <- replace(locID, locID == "PIPA", 2)

write.table(locID, file="locID_num.txt", row.names=F, col.names=F)
```

Alright, I moved the .txt file into the folder with the other input files. Now I'm going to replace the first column of the .ped file with the first column of the locID_num.txt. file. I have to remove the " characters first though...

```{}
#removing the " from the file
sed 's/"//g' locID_num.txt > locID_num_clean.txt
#adding the locID to the .ped file as the first two columns
paste locID_num_clean.txt GCRF_base_file.ped > GCRF_intermed.ped
#reformatting the file so it's the first two columns from the locID file and the rest are from the original file, this takes a while
cat GCRF_intermed.ped | awk 'BEGIN {OFS="\t"} {printf "%s\t%s\t%s", $1, $2, $5; for (i=6; i<=15885534; i++) printf "\t%s", $i; print ""}' > GCRF_base_file_fin.ped
#checking that it went correctly (-S command keeps it from wrapping, so it can show the full line without going crazy and feeding out all the genotypes onto the screen)
less -S GCRF_base_file_fin.ped
#checking that it has the correct number of columns, looks good!
awk '{print NF; exit}' GCRF_base_file_fin.ped
>15885526
```

## 5. PLINK for .bim .bam .fam files
Okay so now that the .map file and .ped file are (hopefully) correctly formatted, I'm going to try and run plink to get the files I need to input into the GEMMA program.

So I'm going to use these base files without the morphology to make base .bim .bam and .fam files. I think I can then edit the .fam file to have the phenotypes information since I haven't been able to get plink's --pheno options to work in the past.

```{}
plink --ped GCRF_base_file_fin.ped --map GCRF_base_file_loc.map --out GCRF_base_file --make-bed --aec --allow-no-sex
```
Awesome, now I have these files:

GCRF_base_file.bed #can't view this as it's the binary file
GCRF_base_file.bim #this has the correct SNP ID's
GCRF_base_file.fam #this has all the info and -9 values for the morphology, so this is what I'll fiddle with...

Okay so I'm going to take these base files and the GCRF_bill_morpho.txt and copy them into the bill_morpho folder so I can work with them within there. I'm going to go ahead and delete everything that was in the folder as I'm starting from scratch. 

I'm going to make a GCRF_bill_morpho.txt file that is formatted just like the .fam file but will all the morphology, then I'm going to rename it and see if I can just use that for the mlmm.

Adding all the morphology I want for the beak stuff into one file to use as a .fam files. 

```{r}
morpho <- read.csv("~/Desktop/GCRF/R data/GCRF_morphology.csv")
#getting rid of the empty cells that came from excel
morpho <- morpho[1:189,]

bill_morpho <- as.data.frame(cbind(morpho$BGP_ID, morpho$nare_length, morpho$beak_depth, morpho$beak_width, morpho$culmen_end_length))
bill_morpho <- bill_morpho[order(bill_morpho$V1), ]
rownames(bill_morpho) <- NULL
bill_morpho <- bill_morpho[48:145,]
rownames(bill_morpho) <- NULL
bill_morpho <- bill_morpho[c(-15,-18,-20,-36,-43,-59,-69,-70,-76,-89,-90),]
rownames(bill_morpho) <- NULL
#missing phenotype data but have genetic data, will impute with GEMMA
bill_morpho <- rbind(bill_morpho, c("23N00557", NA, NA, NA, NA), c("23N00459", NA, NA, NA, NA))
bill_morpho <- bill_morpho[order(bill_morpho$V1), ]
rownames(bill_morpho) <- NULL
#Now I've got the right number of rows! Next, replacing the "-" with NAs as this will export best
bill_morpho <- replace(bill_morpho, bill_morpho == "-", NA)
bill_morpho <- replace(bill_morpho, bill_morpho == "", NA)

FID_file <- read.table("~/Desktop/GCRF/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))
#adding columns for the empty values that are normally in the .ped file
FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
#adding the bill_length as the 6th row, skipping tarsus for now
bill_ped_file <- cbind(FID_file[,1:5], bill_morpho[,2:5])
write.table(bill_ped_file, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/GCRF_bill_morpho.txt", row.names=F, col.names=F)
```

Moving that file over onto the cluster and cleaning it up. Then renamed it to GCRF_bill_morpho.fam

```{}
sed 's/"//g' GCRF_bill_morpho.txt > GCRF_bill_morpho_clean.txt
```

# RUNNING THE GWAS
## 6. Relatedness Matrix

First step is to create a relatedness matrix that can be used for all of the following GWAS analyses.

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_base_file -gk 1 -o GCRF.relate > gemma.relate.err 2> gemma.relate.out
```

Okay so that kept failing because:
ERROR: Enforce failed for number of analyzed individuals equals 0. in src/param.cpp at line 2075 in ProcessCvtPhen

Seems like this might be because the phenotype file doesn't have anything.

So I just copies the relatedness matrix from the bill_length stuff from earlier.

## 7. MLMM models
So Marina got eigen values first and included those in the mlmm so I'm going to do that also.

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho \
	-k output/GCRF_relate.cXX.txt \
	-eigen -o GCRF_bill_morpho_eigen
```

Easy enough. I now have:
GCRF_bill_morpho_eigen.eigenD.txt
GCRF_bill_morpho_eigen.eigenU.txt
GCRF_bill_morpho_eigen.log.txt

Okay so now the files are ready to run a mlmm but there are missing phenotypes. I'm going to just run it with those missing phenotypes just being ignored.

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho -k output/GCRF_relate.cXX.txt -d output/GCRF_bill_morpho.eigen.eigenD.txt -u output/GCRF_bill_morpho.eigen.eigenU.txt -lmm 1 -o GCRF_bill_morpho.mlmm -n 1 2 3 4 -w 500000 -s 5000000
```

Got this error:
GSL ERROR: matrix is singular in lu.c at line 449 errno 1

Which might be due to one of the variables being highly correlated with another, probably beak width and beak length. So dropping beak width as a variable. Putting this into a script so it had enough memory and stuff.

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho -k output/GCRF_relate.cXX.txt -d output/GCRF_bill_morpho.eigen.eigenD.txt -u output/GCRF_bill_morpho.eigen.eigenU.txt -lmm 1 -o GCRF_bill_morpho.mlmm -n 1 3 4 -w 500000 -s 5000000
```

This has been running for 20 hours now, I only allotted 24 hours total so hopefully it completes soon, but this is the farthest I've gotten with this!

#########################################
I'm going to prep the feather morphology stuff. I've already make a data frame with the residuals after a body size correction and that's what I'll put into the GWAS. So I'm going to fix that file so that it can be used at the .fam file in the GWAS.

```{r}
FID_file <- read.table("~/Desktop/GCRF/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))
#adding columns for the empty values that are normally in the .ped file
FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
colnames(FID_file) <- c("FID", "ID", "V3", "V4", "V5")

feather_res_gwas <- read.table("~/Desktop/GCRF/GenomicPipeline/GWAS input/feather_residuals.txt", header=T, sep=" ")
feather_res_gwas <- data.frame(FID_file[,1:5], feather_res_gwas)
write.table(feather_res_gwas, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/feather_residuals_loc.txt", row.names=F, col.names=F)
missing_ids <- setdiff(feather_res_gwas$ID, FID_file$ID)
```

```{}
sed 's/"//g' feather_residuals_loc.txt > feather_residuals_clean.txt

mv GCRF_base_file.bed GCRF_feather.bed
mv GCRF_base_file.bim GCRF_feather.bim
mv feather_residuals_clean.txt GCRF_feather.fam
```

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_feather \
	-k output/GCRF_relate.cXX.txt \
	-eigen -o GCRF_feather_eigen
```

GCRF_feather_eigen.eigenD.txt 
GCRF_feather_eigen.eigenU.txt
GCRF_feather_eigen.log.txt

Testing this outside of a script to make sure it runs. Looks like it didn't throw the correlation error, so putting it into a script and letting it cook!
```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_feather -k\
output/GCRF_relate.cXX.txt -d output/GCRF_feather_eigen.eigenD.txt \
-u output/GCRF_feather_eigen.eigenU.txt -lmm 1 \
-o GCRF_bill_morpho.mlmm -n 1 2 3 4 5 -w 500000 -s 5000000
```
Gosh, seems like this is running too! I'm just going to adjust the time to a day and a half...

So that ran but after processing it said I have ~257,000 significant SNPs, so I don't trust it at all. So back to lmm and bslmm models for individuals phenotypes I guess, what a bummer.



## 8. BSLMM and LMM
```{r, message=F, LIBRARIES}
library(MASS)
library(caret)
library(ggplot2)
library(dplyr)
library(FactoMineR)
library(factoextra)
library(corrplot)
library(ggcorrplot)
library(rstatix)
library(dplyr)
```
### Feather PCA and BSLMM
What I can do for the feather stuff at least is to run a PCA with the feather morphology and then use the first PC as the phenotype in the GWAS, that was it's one variable that is still accounting for all of the feather morphology...

This whole first part is just to get the site ID information arranged correctly so I can plug it into the feather data more easily. Just ripped from the GCRF_lms.rmd file.
```{r}
morpho <- read.csv("~/Desktop/GCRF/R data/GCRF_morphology.csv")
morpho <- morpho[1:189,]

morpho_num <- as.data.frame(cbind(as.numeric(morpho$wing), as.numeric(morpho$tarsus), as.numeric(morpho$weight), as.numeric(morpho$tail_length), as.numeric(morpho$nare_length), as.numeric(morpho$bristle_length), as.numeric(morpho$beak_depth), as.numeric(morpho$beak_width), as.numeric(morpho$nare_end_length), as.numeric(morpho$culmen_end_length)))

colnames(morpho_num) <- c("wing", "tarsus", "weight", "tail_length", "nare_length", "bristle_length", "beak_depth", "beak_width", "nare_end_length", "culmen_end_length")
morpho2 <- as.data.frame(cbind(morpho$BGP_ID, as.factor(morpho$site),morpho_num))
colnames(morpho2) <- c("BGP_ID", "site", "wing", "tarsus", "weight", "tail_length", "nare_length", "bristle_length", "beak_depth", "beak_width", "nare_end_length", "culmen_end_length")

morpho_avg <- morpho2 %>% group_by(BGP_ID, site) %>% reframe(wing=mean(wing), tarsus=mean(tarsus), weight=mean(weight), tail_length=mean(tail_length), nare_length=mean(nare_length), bristle_length=mean(bristle_length), beak_depth=(beak_depth), beak_width=mean(beak_width), nare_end_length=mean(nare_end_length), culmen_end_length=mean(culmen_end_length))
```
```{r}
feather <- read.csv("~/Desktop/GCRF/R data/GCRF_feather_morpho.csv")
site <- data.frame(morpho_avg$site, morpho_avg$BGP_ID, morpho_avg$tarsus)
colnames(site) <- c("site", "ID", "tarsus")

feather_avg <- feather %>% group_by(ID) %>% summarize(pen_length=mean(Penn_length), pen_num=mean(Penn_number), plum_length=mean(Plum_length), plum_number=mean(Plum_number), nodes=mean(Nodes))

feather_avg <- merge(feather_avg, site, by="ID")
#reordering the columns so they're tidy
feather_avg <- data.frame(feather_avg$ID, feather_avg$site, feather_avg$pen_length, feather_avg$pen_num, feather_avg$plum_length, feather_avg$plum_number, feather_avg$nodes, feather_avg$tarsus)
colnames(feather_avg) <- c("BGP_ID", "site", "pen_length", "pen_num", "plum_length", "plum_number", "nodes", "tarsus")
```


```{r}
#feather_avg <- feather_avg[!duplicated(feather_avg$ID),]

#removing and NAs, only losing 3 individuals
feather_avg <- na.omit(feather_avg)
colSums(is.na(feather_avg))
#including the tarsus variable to maybe account for body size differences?
morpho_num <- feather_avg[,3:8]

corr_m <- cor(morpho_num)
ggcorrplot(corr_m)
#none of them have a value higher than |.53|, so I'm going to say that any correlation is fine
#actually going to remove the tarsus because its so weakly correlated with everything...
morpho <- feather_avg[,3:7]
```

```{r}
feather_avg #the df with individuals (character) and groups (factor)
morpho #the df with just the variables
#nothing has been normalized

#taking the active variable, ie individuals and the measurements that go into the PCA
morpho.active <- feather_avg[,c(1, 3:7)]

#now need to standardize the variables, but using the PCA function that's already done so don't need to do it prior to running the PCA in this case, otherwise use scale()... x-mean(x)/sd(x)
#need to remove the column for individual though...
morpho.active.1 <- morpho.active[,-1]
morpho.pca <- PCA(morpho.active.1,  scale.unit=T, graph=F)

eig.val <- get_eig(morpho.pca)
eig.val
fviz_eig(morpho.pca, addlabels = T)
#this is awesome, the first PC explains over half the variation which suggests it'll be a representative variable to run in the GWAS
```

```{r}
var <- get_pca_var(morpho.pca)
head(var$coord)

```

```{r}

fviz_pca_var(morpho.pca, col.var="black")
fviz_cos2(morpho.pca, choice="var", axes=1:4)
library("corrplot")
corrplot(var$contrib, is.corr=F)
#so I'm intrepreting this as PC1 representing the structure overall, as opposed to one variable more strongly than any others
```

```{r}
#fviz_pca_ind(morpho.pca)
#fviz_pca_ind(morpho.pca, col.ind = "cos2", 
#             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#             repel = TRUE # Avoid text overlapping (slow if many points)
#             )
```

```{r}
fviz_pca_ind(morpho.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = feather_avg$site, # color by groups
             palette = c("#00AFBB", "#E7B800"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups",
             )
ind <- morpho.pca$ind
coord <- as.data.frame(ind$coord)
PC1 <- data.frame(feather_avg$BGP_ID, coord$Dim.1)
colnames(PC1) <- c("BGP_ID", "PC1")

```

Awesome, so now that I have that set up, I can add the missing BGP_ID's and there are some random duplicates...

```{r}
PC1 <- PC1[c(-17, -31, -47, -60, -73, -74),]
feather_PC1 <- rbind(PC1,
                     c("23N00444", NA),
                     c("23N00447", NA), 
                     c("23N00448", NA), 
                     c("23N00450", NA), 
                     c("23N00451", NA), 
                     c("23N00452", NA),
                     c("23N00454", NA), 
                     c("23N00456", NA), 
                     c("23N00457", NA), 
                     c("23N00459", NA), 
                     c("23N00462", NA), 
                     c("23N00557", NA))
#missing phenotype data but have genetic data, will impute with GEMMA
feather_PC1 <- feather_PC1[order(feather_PC1$BGP_ID), ]
rownames(feather_PC1) <- NULL
```

```{r}

FID_file <- read.table("~/Desktop/GCRF/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))
#adding columns for the empty values that are normally in the .ped file
FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
#adding the bill_length as the 6th row, skipping tarsus for now
feather.PC_ped_file <- cbind(FID_file[,1:5], feather_PC1[,2])
write.table(feather.PC_ped_file, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/GCRF_feather_PC.txt", row.names=F, col.names=F)
```

```{}
sed 's/"//g' GCRF_feather_PC.txt > GCRF_feather_PC_clean.txt
paste GCRF_feather_PC_clean.txt ../GCRF_genotypes.ped > GCRF_feather_PC.ped

plink --ped GCRF_feather_PC.ped --map GCRF_base_file_loc.map --out GCRF_feather_PC --make-bed --aec --allow-no-sex 
```

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_feather_PC \
  -gk 1 \ 
	-o GCRF_relate_PC.cXX.txt \
```

```{}
#in a script
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_feather_PC \
-k output/GCRF_relate_PC.cXX.txt -bslmm 1 -o GCRF_feather_PC.bslmm \
-w 500000 -s 5000000

```
GCRF_feather_PC.bslmm.gamma.txt
GCRF_feather_PC.bslmm.param.txt
GCRF_feather_PC.bslmm.hyp.txt
GCRF_feather_PC.bslmm.log.txt

Processed this with GEMMA_bslmm_output_analysis_working.R and got at least two snps that flag as being important.


######################################################################################################################################################

First step is making files for every phenotypes...bleh. Lets do beak length, nare length, bristles... wing chord and all the feather morphology. The beak length stuff should already be set up, so let's use that same code to make the other ones.


### Bill PCA and BSLMM
Repeating the steps from the feather PCA but with the bill data.

```{r}
morpho <- read.csv("~/Desktop/GCRF/GCRF-Present/R data/GCRF_morphology.csv")
morpho <- morpho[1:189,]

morpho_num <- as.data.frame(cbind(as.numeric(morpho$nare_length), as.numeric(morpho$beak_depth), as.numeric(morpho$beak_width), as.numeric(morpho$culmen_end_length)))

colnames(morpho_num) <- c("nare_length", "beak_depth", "beak_width", "culmen_end_length")
morpho2 <- as.data.frame(cbind(morpho$BGP_ID, as.factor(morpho$site), morpho_num))
colnames(morpho2) <- c("BGP_ID", "site", "nare_length", "beak_depth", "beak_width", "culmen_end_length")

morpho_avg <- morpho2 %>% group_by(BGP_ID, site) %>% reframe(nare_length=mean(nare_length), beak_depth=(beak_depth), beak_width=mean(beak_width), culmen_end_length=mean(culmen_end_length))
morpho_avg <- morpho_avg[48:145,]
morpho_avg <- morpho_avg[c(-15,-18,-19,-43,-60,-70,-75,-88,-89),]
```

```{r}
#feather_avg <- feather_avg[!duplicated(feather_avg$ID),]

#removing and NAs, losing 14 individuals
morpho_avg <- na.omit(morpho_avg)
colSums(is.na(morpho_avg))
morpho_num <- morpho_avg[,3:6]


corr_m <- cor(morpho_num)
ggcorrplot(corr_m)
```

```{r}
morpho_avg #the df with individuals (character) and groups (factor)
morpho_num #the df with just the variables
#nothing has been normalized

#taking the active variable, ie individuals and the measurements that go into the PCA
morpho.active <- morpho_avg[,c(1, 3:6)]

#now need to standardize the variables, but using the PCA function that's already done so don't need to do it prior to running the PCA in this case, otherwise use scale()... x-mean(x)/sd(x)
#need to remove the column for individual though...
morpho.active.1 <- morpho.active[,-1]
morpho.pca <- PCA(morpho.active.1,  scale.unit=T, graph=F)

eig.val <- get_eig(morpho.pca)
eig.val
fviz_eig(morpho.pca, addlabels = T)
#this is awesome, the first PC explains over half the variation which suggests it'll be a representative variable to run in the GWAS
```

```{r}
var <- get_pca_var(morpho.pca)
var$coord
```

```{r}
fviz_pca_var(morpho.pca, col.var="black")
fviz_cos2(morpho.pca, choice="var", axes=1:4)
library("corrplot")
corrplot(var$contrib, is.corr=F)
#so I'm intrepreting this as PC1 representing the beak morphology overall, as opposed to one variable more strongly than any others
```

```{r}
fviz_pca_ind(morpho.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = morpho_avg$site, # color by groups
             palette = c("#00AFBB", "#E7B800"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups",
             )
ind <- morpho.pca$ind
coord <- as.data.frame(ind$coord)
PC1 <- data.frame(morpho_avg$BGP_ID, coord$Dim.1)
colnames(PC1) <- c("BGP_ID", "PC1")

```

Awesome, so now that I have that set up, I can add the missing BGP_ID's and there are some random duplicates...

```{r}
bill_PC1 <- rbind(PC1,
                  c("23N00437", NA),
                  c("23N00444", NA), 
                  c("23N00448", NA), 
                  c("23N00460", NA), 
                  c("23N00464", NA), 
                  c("23N00467", NA),
                  c("23N00475", NA), 
                  c("23N00479", NA), 
                  c("23N00480", NA), 
                  c("23N00485", NA), 
                  c("23N00488", NA),
                  c("23N00489", NA), 
                  c("23N00490", NA),
                  c("23N00491", NA))
#missing phenotype data but have genetic data, will impute with GEMMA
bill_PC1 <- bill_PC1[order(bill_PC1$BGP_ID), ]
rownames(bill_PC1) <- NULL
```

```{r}
FID_file <- read.table("~/Desktop/GCRF/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))
#adding columns for the empty values that are normally in the .ped file
FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
#adding the bill_length as the 6th row, skipping tarsus for now
bill.PC_ped_file <- cbind(FID_file[,1:5], bill_PC1[,2])
write.table(bill.PC_ped_file, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/GCRF_bill_PC.txt", row.names=F, col.names=F)
```

Moving it over to the cluster...

```{}
sed 's/"//g' GCRF_bill_PC.txt > GCRF_bill_PC_clean.txt
paste GCRF_bill_PC_clean.txt ../GCRF_genotypes.ped > GCRF_bill_PC.ped

plink --ped GCRF_bill_PC.ped --map GCRF_base_file_loc.map --out GCRF_bill_PC --make-bed --aec --allow-no-sex 
```

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_PC -gk 1 -o GCRF_relate_PC.cXX.txt
```

```{}
#in a script
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_PC \
-k output/GCRF_relate_PC.cXX.txt -bslmm 1 -o GCRF_bill_PC.bslmm \
-w 500000 -s 5000000

```
GCRF_feather_PC.bslmm.gamma.txt
GCRF_feather_PC.bslmm.param.txt
GCRF_feather_PC.bslmm.hyp.txt
GCRF_feather_PC.bslmm.log.txt



## nare length


Running this in a script so it doesn't time out or run out of space. Using one that Marina wrote, again not sure about the -w and -s numbers...


```{}
module load gcc/10.3
module load openmpi
conda activate GWAS2

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho \
-k output/GCRF_relate.cXX.txt -bslmm 1 -o GCRF_bill_morpho_bslmm \
-w 500000 -s 5000000
```

Okay so that seems to have worked! I now have:
GCRF_bill_morpho_bslmm.bv.txt     
GCRF_bill_morpho_bslmm.hyp.txt  
GCRF_bill_morpho_bslmm.param.txt
GCRF_bill_morpho_bslmm.gamma.txt  
GCRF_bill_morpho_bslmm.log.txt


Now I'm going to run each individual feather trait in the bslmm, and do a PCA for bill morphology and run those traits through bslmm too. Oh boy! Might try a lmm on the feather PCA too just to see if I get anything.









########################################################################################################################

So from this I should be able to predict the missing values.

```{}
module load gcc/10.3
module load openmpi
conda activate GWAS2

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho -epm output/GCRF_bill_morpho_bslmm.param.txt -emu output/GCRF_bill_morpho_bslmm.log.txt -ebv output/GCRF_bill_morpho_bslmm.bv.txt -k output/GCRF_bill_length.relate.cXX.txt -predict 1 -o GCRF_bill_morpho_bslmm
```
















## LMM
Setting up a spreadsheet with all of the morphology separately.
```{r}
morpho <- read.csv("~/Desktop/GCRF/GCRF-Present/R data/GCRF_morphology.csv", na.strings = c("", "NA"))
morpho <- morpho[1:189,]
morpho <- morpho[order(morpho$BGP_ID), ]
rownames(morpho) <- NULL
morpho <- morpho[(48:145),]
rownames(morpho) <- NULL
morpho <- morpho[!duplicated(morpho$BGP_ID),]
morpho <- morpho[c(-15, -18,-35,-66),]
morpho <- rbind(morpho,
                c("PIPA", NA, NA, "23N00557", NA, NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA),
                c("PIPA", NA, NA, "23N00459", NA, NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA,NA))
rownames(morpho) <- NULL
morpho <- morpho[order(morpho$BGP_ID), ]

FID_file <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))

FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
#adding the bill_length as the 6th row, skipping tarsus for now
morpho_ped_file <- data.frame(FID_file, morpho[,c(5:18)])
write.table(morpho_ped_file, file="~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_morpho.ped", row.names=F, col.names=F)
```

```{}
sed 's/"//g' GCRF_morpho.ped > GCRF_morpho_clean.ped
sed 's/-/NA/g' GCRF_morpho_clean.ped > GCRF_morpho.ped

cut -d ' ' -f 1-6 GCRF_morpho.ped > wing
cut -d ' ' -f 1-5,7 GCRF_morpho.ped > tarsus
cut -d ' ' -f 1-5,10 GCRF_morpho.ped > tail_length
cut -d ' ' -f 1-5,12 GCRF_morpho.ped > nare_length
cut -d ' ' -f 1-5,13 GCRF_morpho.ped > bristle_length
cut -d ' ' -f 1-5,14 GCRF_morpho.ped > beak_depth
cut -d ' ' -f 1-5,15 GCRF_morpho.ped > beak_width
cut -d ' ' -f 1-5,19 GCRF_morpho.ped > culmen_end_length

paste wing ../GCRF_genotypes.ped > wing.ped
paste tarsus ../GCRF_genotypes.ped > tarsus.ped
paste tail_length ../GCRF_genotypes.ped > tail_length.ped
paste nare_length ../GCRF_genotypes.ped > nare_length.ped
paste bristle_length ../GCRF_genotypes.ped > bristle_length.ped
paste beak_depth ../GCRF_genotypes.ped > beak_depth.ped
paste beak_width ../GCRF_genotypes.ped > beak_width.ped
paste culmen_end_length ../GCRF_genotypes.ped > culmen_end_length.ped
```

```{}
plink --ped wing_output/wing.ped --map GCRF_base_file_loc.map --out wing_output/wing --make-bed --aec --allow-no-sex

plink --ped tarsus_output/tarsus.ped --map GCRF_base_file_loc.map --out tarsus_output/tarsus --make-bed --aec --allow-no-sex

plink --ped tail_length_output/tail_length.ped --map GCRF_base_file_loc.map --out tail_length_output/tail_length --make-bed --aec --allow-no-sex

plink --ped nare_length_output/nare_length.ped --map GCRF_base_file_loc.map --out nare_length_output/nare_length --make-bed --aec --allow-no-sex

plink --ped bristle_length_output/bristle_length.ped --map GCRF_base_file_loc.map --out bristle_length_output/bristle_length --make-bed --aec --allow-no-sex

plink --ped beak_depth_output/beak_depth.ped --map GCRF_base_file_loc.map --out beak_depth_output/beak_depth --make-bed --aec --allow-no-sex

plink --ped beak_width_output/beak_width.ped --map GCRF_base_file_loc.map --out beak_width_output/beak_width --make-bed --aec --allow-no-sex

plink --ped culmen_end_length_output/culmen_end_length.ped --map GCRF_base_file_loc.map --out culmen_end_length_output/culmen_end_length --make-bed --aec --allow-no-sex

```

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile wing -k output/relate.cXX.txt.cXX.txt -lmm 1 -o wing_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile tarsus -gk 1 -o relate
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile tarsus -k output/relate.cXX.txt -lmm 1 -o tarsus_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile tail_length -gk 1 -o relate
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile tail_length -k output/relate.cXX.txt -lmm 1 -o tail_length_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile nare_length -gk 1 -o relate
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile nare_length -k output/relate.cXX.txt -lmm 1 -o nare_length_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile bristle_length -gk 1 -o relate
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile bristle_length -k output/relate.cXX.txt -lmm 1 -o bristle_length_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile beak_depth -gk 1 -o relate
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile beak_depth -k output/relate.cXX.txt -lmm 1 -o beak_depth_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile beak_width -gk 1 -o relate
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile beak_width -k output/relate.cXX.txt -lmm 1 -o beak_width_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile culmen_end_length -gk 1 -o relate
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile culmen_end_length -k output/relate.cXX.txt -lmm 1 -o culmen_end_length_lmm -w 500000 -s 5000000
```



# GETTING ALLELE FREQUENCIES FOR THE FLAGGED SNPS

## BSLMM SNPs
So with filtering using the bslmm output for feathers and bills, I was able to identify some relevant SNPS.
For the bill morphology: (above a .1 threshold for PIP values, per Chaves 2016).
chr	rs	ps	n_miss	alpha	beta	gamma	eff
Scaffold_2__1_contigs__length_114774641	Scaffold_2__1_contigs__length_114774641.41203916	41203916	0	-6.108376e-06	-0.8424626	0.114384	0.0963642420384

For the feather morphology:
chr	rs	ps	n_miss	alpha	beta	gamma	eff
Scaffold_28__1_contigs__length_11240829	Scaffold_28__1_contigs__length_11240829.4798538	4798538	0	-3.014029e-06	-1.487668	0.389188	0.578982533584
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.91081692	91081692	0	-2.156909e-06	-1.643631	0.25956	0.42662086236
Scaffold_18__1_contigs__length_25274884	Scaffold_18__1_contigs__length_25274884.2074884	2074884	0	2.328089e-06	0.9979339	0.23866	0.238166904574

So the goal now is to use plink to pull out these specific snps, get recode the alleles to 0 (homozygous reference), 1 (heterozygous), and 2 (homozygous alternative) for each individual. The idea then is to run a regression with those categorical variables again the continuous bill of feather PC values and see if there's a relationship.

So first, we're going to be pull the SNP info from the vcf that was put into the GWAS, so the imputed file: GCRF_merged_impute4.1.sort.rm_dup.vcf. PLINK should be able to pull out a SNP based on it's name, but because the SNPS are namemd after the scaffold and then given a position, I'm not sure how to extract the one I'm looking for. 

This is how the SNP is stored in the vcf file:
Scaffold_292__1_contigs__length_11690	8856	.	A	G	.	PASS	.	GT	1|1	0|0	0|0	0|0	0|1	0|1	0|0	1|1	1|1	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|1	0|1	1|1	0|0	0|1	0|0	1|1	0|1	0|0	1|1	0|1	1|1	0|0	0|0	0|0	0|1	0|0	0|0	0|1	0|0	0|1	0|0	0|0	0|0	0|0	0|0	0|1	0|0	1|1	0|0	0|0	0|1	0|0	0|0	0|0	0|1	0|1	0|0	0|0	0|0	1|1	0|0	0|1	0|0	1|1	1|1	0|0	0|0	0|0	0|0	1|1	0|0	0|0	0|0	0|0	1|1	0|0	0|0	0|1	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0

### Using PLINK to extract the values
So the names are all there, just gotta figure out the right command for plink...

We'll start with code the CH helped me with. We'll start with the bill SNP.

```{}
conda activate GWAS2
plink --vcf GCRF_merged_impute4.1.sort.rm_dup.vcf --out Scaffold_2__1_contigs__length_114774641.41203916.bill --aec --recodeA --snp Scaffold_2__1_contigs__length_114774641.41203916
```

As I expected, it couldn't fine it. So I'm going to try and use another approach...


```{}
plink --vcf ../impute/GCRF_merged_impute4.1.sort.rm_dup.vcf --aec --recode --out allele_freq

cat allele_freq.map | awk '{print$1"\t"$1"."$4"\t"$3"\t"$4}' > allele_freq_loc.map
mv allele_freq_loc.map allele_freq.map

plink --ped allele_freq.ped --map allele_freq.map --out allele_freq --make-bed --aec --allow-no-sex

plink --bfile allele_freq --snp Scaffold_2__1_contigs__length_114774641.41203916 --aec --recodeA --out snps/Scaffold_2__1_contigs__length_114774641.41203916.bill
```

That worked!! I now have a .fam file that has the info on the genotype! So I can import that here, merge it with the phenotype info and there ya go. Sick. So I'm just gonna do that again real quick with the feather SNPs...

```{}
plink --bfile allele_freq --snp Scaffold_28__1_contigs__length_11240829.4798538 --aec --recodeA --out snps/Scaffold_28__1_contigs__length_11240829.4798538.feather

plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.91081692 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.91081692.feather

plink --bfile allele_freq --snp Scaffold_18__1_contigs__length_25274884.2074884 --aec --recodeA --out snps/Scaffold_18__1_contigs__length_25274884.2074884.feather

```

### Importing data into R

Moving back into R after I moved the files over, going to organizing everything here.
```{r}
feather.PC_ped_file <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_feather_PC.txt")
colnames(feather.PC_ped_file) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE")
bill.PC_ped_file <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_bill_PC.txt")
colnames(bill.PC_ped_file) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE")


S1.1.91081692.feather <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_1__1_contigs__length_151072562.91081692.feather.raw", sep = " ", header = T)
S1.1.91081692.feather <- data.frame(feather.PC_ped_file, S1.1.91081692.feather[,7])
colnames(S1.1.91081692.feather) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")

S18.1.2074884.feather <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_18__1_contigs__length_25274884.2074884.feather.raw", sep = " ", header = T)
S18.1.2074884.feather <- data.frame(feather.PC_ped_file, S18.1.2074884.feather[,7])
colnames(S18.1.2074884.feather) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")

S28.1.4798539.feather <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_28__1_contigs__length_11240829.4798538.feather.raw", sep = " ", header = T)
S28.1.4798539.feather <- data.frame(feather.PC_ped_file, S28.1.4798539.feather[,7])
colnames(S28.1.4798539.feather) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")

S2.1.41203916.bill <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_2__1_contigs__length_114774641.41203916.bill.raw", sep = " ", header = T)
S2.1.41203916.bill <- data.frame(bill.PC_ped_file, S2.1.41203916.bill[,7])
colnames(S2.1.41203916.bill) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")

```

### Running and plotting linear regressions

Going to try running these in a linear regression, there are some NA values that might throw errors...

```{r, S1.1.91081692}
library(ggplot2)
S1.feather <- lm(PHENOTYPE~GENOTYPE, data=S1.1.91081692.feather)
summary(S1.feather)
#p=5.19e-06!!!
S1.1.91081692.feather.no.na <- na.omit(S1.1.91081692.feather)

S1.feather.results <- data.frame(FID=as.factor(S1.1.91081692.feather.no.na$FID), GENOTYPE = as.numeric(S1.1.91081692.feather.no.na$GENOTYPE), PHENOTYPE = as.numeric(S1.1.91081692.feather.no.na$PHENOTYPE))
S1.feather.results$predicted <- predict(S1.feather)

y_breaks <- pretty(range(S1.feather.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S1.feather.results$GENOTYPE), max(S1.feather.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S1.91081692 <- ggplot(S1.feather.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle("Feather Morphology vs SNP S1.91081692") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("GG", "GA", "AA")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S1.91081692.png", plot=S1.91081692, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r, S18.1.2074884}
S18.feather <- lm(PHENOTYPE~GENOTYPE, data=S18.1.2074884.feather)
summary(S18.feather)
#p=0.001585!!!
S18.1.2074884.feather.no.na <- na.omit(S18.1.2074884.feather)

S18.feather.results <- data.frame(FID=as.factor(S18.1.2074884.feather.no.na$FID), GENOTYPE = as.numeric(S18.1.2074884.feather.no.na$GENOTYPE), PHENOTYPE = as.numeric(S18.1.2074884.feather.no.na$PHENOTYPE))
S18.feather.results$predicted <- predict(S18.feather)

y_breaks <- pretty(range(S18.feather.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S18.feather.results$GENOTYPE), max(S18.feather.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S18.2074884 <- ggplot(S18.feather.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle("Feather Morphology vs SNP S18.2074884") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("CC", "CT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S18.2074884.png", plot=S18.2074884, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r, S28.1.4798539}
S28.feather <- lm(PHENOTYPE~GENOTYPE, data=S28.1.4798539.feather)
summary(S28.feather)
#p=9.482e-10!!!
S28.1.4798539.feather.no.na <- na.omit(S28.1.4798539.feather)

S28.feather.results <- data.frame(FID=as.factor(S28.1.4798539.feather.no.na$FID), GENOTYPE = as.numeric(S28.1.4798539.feather.no.na$GENOTYPE), PHENOTYPE = as.numeric(S28.1.4798539.feather.no.na$PHENOTYPE))
S28.feather.results$predicted <- predict(S28.feather)

y_breaks <- pretty(range(S28.feather.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S28.feather.results$GENOTYPE), max(S28.feather.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S28.4798539 <- ggplot(S28.feather.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle("Feather Morphology vs SNP S28.4798539") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("AA", "AG", "GG")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S28.4798539.png", plot=S28.4798539, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)

```

```{r, S2.1.41203916}
S2.bill <- lm(PHENOTYPE~GENOTYPE, data=S2.1.41203916.bill)
summary(S2.bill)
#p=3.949e-09!!!
S2.1.41203916.bill.no.na <- na.omit(S2.1.41203916.bill)

S2.bill.results <- data.frame(FID=as.factor(S2.1.41203916.bill.no.na$FID), GENOTYPE = as.numeric(S2.1.41203916.bill.no.na$GENOTYPE), PHENOTYPE = as.numeric(S2.1.41203916.bill.no.na$PHENOTYPE))
S2.bill.results$predicted <- predict(S2.bill)

y_breaks <- pretty(range(S2.bill.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S2.bill.results$GENOTYPE), max(S2.bill.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S2.41203916 <- ggplot(S2.bill.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Bill Morphology PC1") +  # Add axis labels
  ggtitle("Bill Morphology vs SNP S2.41203916") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("TT", "TG", "GG")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S2.41203916.png", plot=S2.41203916, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)

```


## LMM SNPs
So I ran some LMM's on all of the morphological traits separately, and got some interesting results. I ran them first with a Bonferonii correction (0.05/7900000). I then also ran it with the widely accept 5e-8 significant threshold.

Here's what I got:
WING > 1754 SNP @<6.25e-9
TARSUS > 0 SNP @<6.25e-9 > 0 SNP @5e-8
TAIL LENGTH > 1 SNP @<6.25e-9 13 SNP @5e-8
NARE LENGTH > 0 SNP @<6.25e-9 2 SNP @5e-8
BRISTLE LENGTH > 0 SNP @<6.25e-9 0 SNP @5e-8
BEAK WIDTH > 1 SNP @<6.25e-9 2 SNP @5e-8
BEAK DEPTH > 0 SNP @<6.25e-9 0 SNP @5e-8
CULMEN-END LENGTH > 0 SNP @<6.25e-9 5 SNP @5e-8

So I don't think I'm going to try the allele regression method for all 1754 SNPs for the Wing cause that'd be crazy, but doing some for the bill dimensions would be cool because it makes it a bit clearer what's happening compared to the bill PC equivalent. So let's play with those two SNPs for Nare Length and Beak Width.

Nare Length: (ref>alt)
Scaffold_6__1_contigs__length_71876329.70598544 G>T
Scaffold_6__1_contigs__length_71876329.70598546 G>C

These are right next to each other which is interesting. 

Beak Width
Scaffold_1__1_contigs__length_151072562.25482370 C>T
Scaffold_2__1_contigs__length_114774641.86583423 C>T

So I'm going to use the same plink command and pull out a new VCF file.

### Extracting with PLINK
```{}
plink --bfile allele_freq --snp Scaffold_6__1_contigs__length_71876329.70598544 --aec --recodeA --out snps/Scaffold_6__1_contigs__length_71876329.70598544.nare_length

plink --bfile allele_freq --snp Scaffold_6__1_contigs__length_71876329.70598546 --aec --recodeA --out snps/Scaffold_6__1_contigs__length_71876329.70598546.nare_length

########################
plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.25482370 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.25482370.beak_width

plink --bfile allele_freq --snp Scaffold_2__1_contigs__length_114774641.86583423 --aec --recodeA --out snps/Scaffold_2__1_contigs__length_114774641.86583423.beak_width
```


### Importing everything into R
Moving stuff over from the cluster

```{r}
morpho.ped <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_morpho.ped")
morpho <- data.frame(morpho.ped$V1, morpho.ped$V2, morpho.ped[,6:19])
colnames(morpho) <- c("FID", "BGP_ID", "wing", "tarsus", "weight", "bill_color", "tail_length", "bristle_cov", "nare_length", "bristle_length", "beak_depth", "beak_width", "tarsus_hair_full", "tarsus_hair_jointdown", "nare_end_length", "culmen_end_length")

#Scaffold_6__1_contigs__length_71876329.70598544 nare length
S6.70598544.nare_length <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_6__1_contigs__length_71876329.70598544.nare_length.raw", sep = " ", header = T)
S6.70598544.nare_length <- data.frame(morpho[,1:2], morpho$nare_length, S6.70598544.nare_length[,7])
colnames(S6.70598544.nare_length) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")

#Scaffold_6__1_contigs__length_71876329.70598546 nare length
S6.70598546.nare_length <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_6__1_contigs__length_71876329.70598546.nare_length.raw", sep = " ", header = T)
S6.70598546.nare_length <- data.frame(morpho[,1:2], morpho$nare_length, S6.70598546.nare_length[,7])
colnames(S6.70598546.nare_length) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")

#Scaffold_1__1_contigs__length_151072562.25482370 beak width
S1.25482370.beak_width <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_1__1_contigs__length_151072562.25482370.beak_width.raw", sep = " ", header = T)
S1.25482370.beak_width <- data.frame(morpho[,1:2], morpho$beak_width, S1.25482370.beak_width[,7])
colnames(S1.25482370.beak_width) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")

#Scaffold_2__1_contigs__length_114774641.86583423 beak width
S2.86583423.beak_width <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_2__1_contigs__length_114774641.86583423.beak_width.raw", sep = " ", header = T)
S2.86583423.beak_width <- data.frame(morpho[,1:2], morpho$beak_width, S2.86583423.beak_width[,7])
colnames(S2.86583423.beak_width) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")
```

```{r, S6.70598544.nare_length}
library(ggplot2)
S6.544.nare <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=S6.70598544.nare_length)
summary(S6.544.nare)
#p=3.417e-08
S6.70598544.nare_length.no.na <- na.omit(S6.70598544.nare_length)

S6.544.nare.results <- data.frame(FID=as.factor(S6.70598544.nare_length.no.na$FID), GENOTYPE = as.numeric(S6.70598544.nare_length.no.na$GENOTYPE), PHENOTYPE = as.numeric(S6.70598544.nare_length.no.na$PHENOTYPE))
S6.544.nare.results$predicted <- predict(S6.544.nare)

y_breaks <- pretty(range(S6.544.nare.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S6.544.nare.results$GENOTYPE), max(S6.544.nare.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S6.70598544 <- ggplot(S6.544.nare.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Nare Length (mm)") +  # Add axis labels
  ggtitle("Nare Length vs SNP S6.70598544") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("GG", "GT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S6.70598544.png", plot=S6.70598544, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r, S6.70598546.nare_length}
S6.546.nare <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=S6.70598546.nare_length)
summary(S6.546.nare)
#p=3.417e-08
S6.70598546.nare_length.no.na <- na.omit(S6.70598546.nare_length)

S6.546.nare.results <- data.frame(FID=as.factor(S6.70598546.nare_length.no.na$FID), GENOTYPE = as.numeric(S6.70598546.nare_length.no.na$GENOTYPE), PHENOTYPE = as.numeric(S6.70598546.nare_length.no.na$PHENOTYPE))
S6.546.nare.results$predicted <- predict(S6.546.nare)

y_breaks <- pretty(range(S6.546.nare.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S6.546.nare.results$GENOTYPE), max(S6.546.nare.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S6.70598546 <- ggplot(S6.546.nare.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Nare Length (mm)") +  # Add axis labels
  ggtitle("Nare Length vs SNP S6.70598546") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("GG", "GC", "CC")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S6.70598546.png", plot=S6.70598546, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

Interestingly, these seem to be identical. I'm going to go ahead and say that's because they're next to each other and completely linked.

```{r, S1.25482370.beak_width}
S1.width <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=S1.25482370.beak_width)
summary(S1.width)
#p=5.87e-09
S1.25482370.beak_width[S1.25482370.beak_width == "-"] <- NA
S1.25482370.beak_width.no.na <- na.omit(S1.25482370.beak_width)

S1.width.results <- data.frame(FID=as.factor(S1.25482370.beak_width.no.na$FID), GENOTYPE = as.numeric(S1.25482370.beak_width.no.na$GENOTYPE), PHENOTYPE = as.numeric(S1.25482370.beak_width.no.na$PHENOTYPE))
S1.width.results$predicted <- predict(S1.width)

y_breaks <- pretty(range(S1.width.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S1.width.results$GENOTYPE), max(S1.width.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S1.25482370 <- ggplot(S1.width.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Beak Width (mm)") +  # Add axis labels
  ggtitle("Beak Width vs SNP S1.25482370") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("CC", "CT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S1.25482370.png", plot=S1.25482370, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r, S2.86583423.beak_width}
S2.width <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=S2.86583423.beak_width)
summary(S2.width)
#p=9.965e-09
S2.86583423.beak_width[S2.86583423.beak_width == "-"] <- NA
S2.86583423.beak_width.no.na <- na.omit(S2.86583423.beak_width)

S2.width.results <- data.frame(FID=as.factor(S2.86583423.beak_width.no.na$FID), GENOTYPE = as.numeric(S2.86583423.beak_width.no.na$GENOTYPE), PHENOTYPE = as.numeric(S2.86583423.beak_width.no.na$PHENOTYPE))
S2.width.results$predicted <- predict(S2.width)

y_breaks <- pretty(range(S2.width.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S2.width.results$GENOTYPE), max(S2.width.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S2.86583423 <- ggplot(S2.width.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Beak Width (mm)") +  # Add axis labels
  ggtitle("Beak Width vs SNP S2.86583423") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("CC", "CT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S2.86583423.png", plot=S2.86583423  , path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

## Investigating SNP S28.4798539.feather

This one (Scaffold_28__1_contigs__length_11240829.4798538) is only showing on individual as a heterozygote. I'm going to investigate a bit deeper to see how well we can trust the calling on this to see if this is real, also if this was an imputed SNP or not.

Comparing 
GCRF/GWAS/impute/GCRF_merged.vcf
GCRF/GWAS/impute/GCRF_merged_impute4.1.sort.rm_dup.vcf

```{}
wc -l GCRF_merged.vcf
7953291 GCRF_merged.vcf

wc -l GCRF_merged_impute4.1.sort.rm_dup.vcf
7953293 GCRF_merged_impute4.1.sort.rm_dup.vcf
```

So that's weird, it's suggesting that the imputation only added 2 extra SNPs??

```{}
cat GCRF_merged.vcf | grep "Scaffold_28__1_contigs__length_11240829" | grep "4798538"
```

Scaffold_28__1_contigs__length_11240829	4798538	.	A	G	28042.20	PASS	AC=161;AF=0.904;AN=178;DP=629;ExcessHet=0;FS=0;InbreedingCoeff=0.44;MLEAC=176;MLEAF=0.989;MQ=60;MQRankSum=0;QD=22.95;SOR=0.261	GT:AD:DP:GQ:PGT:PID:PL:PS	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1/1:0,3:3:9:.:.:125,9,0:.	1|1:0,7:7:21:1|1:4798535_T_C:302,21,0:4798535	1/1:0,9:9:27:.:.:376,27,0:.	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1/1:0,9:9:27:.:.:342,27,0:.	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1|1:0,6:6:18:1|1:4798535_T_C:263,18,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:347,24,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:315,21,0:4798535	1/1:0,5:5:15:.:.:209,15,0:.	1|1:0,6:6:18:1|1:4798535_T_C:260,18,0:4798535	1/1:0,7:7:21:.:.:292,21,0:.	0/0:4,0:4:0:.:.:0,0,0:.	1|1:0,7:7:21:1|1:4798535_T_C:309,21,0:4798535	0/0:2,0:2:0:.:.:0,0,0:.	1/1:0,3:3:9:.:.:125,9,0:.	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1/1:0,6:6:18:.:.:227,18,0:.	1/1:0,4:4:12:.:.:167,12,0:.	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1|1:0,10:10:30:1|1:4798535_T_C:428,30,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:309,21,0:4798535	1/1:0,2:2:6:.:.:84,6,0:.	1|1:0,15:15:45:1|1:4798535_T_C:675,45,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:315,21,0:4798535	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1|1:0,5:5:15:1|1:4798535_T_C:225,15,0:4798535	1|1:0,10:10:30:1|1:4798535_T_C:450,30,0:4798535	1|1:0,5:5:15:1|1:4798535_T_C:225,15,0:4798535	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1/1:0,5:5:15:.:.:209,15,0:.	1|1:0,7:7:21:1|1:4798535_T_C:315,21,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:334,24,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:311,21,0:4798535	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1|1:0,3:3:9:1|1:4798535_T_C:135,9,0:4798535	1/1:0,4:4:12:.:.:167,12,0:.	1|1:0,6:6:18:1|1:4798535_T_C:270,18,0:4798535	1|1:0,13:13:39:1|1:4798535_T_C:582,39,0:4798535	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	0/0:2,0:2:0:.:.:0,0,0:.	1|1:0,7:7:21:1|1:4798535_T_C:299,21,0:4798535	1|1:0,11:11:33:1|1:4798535_T_C:495,33,0:4798535	0/0:4,0:4:0:.:.:0,0,0:.	1|1:0,12:12:36:1|1:4798535_T_C:540,36,0:4798535	1|1:0,5:5:15:1|1:4798535_T_C:225,15,0:4798535	*0|1:1,4:5:26:0|1:4798535_T_C:165,0,26:4798535*	1|1:0,12:12:36:1|1:4798535_T_C:540,36,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:360,24,0:4798535	1/1:0,10:10:30:.:.:418,30,0:.	1/1:0,7:7:21:.:.:292,21,0:.	0/0:2,0:2:0:.:.:0,0,0:.	1|1:0,9:9:27:1|1:4798535_T_C:386,27,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:305,21,0:4798535	1|1:0,6:6:18:1|1:4798535_T_C:260,18,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:315,21,0:4798535	0/0:3,0:3:0:.:.:0,0,0:.	1|1:0,10:10:30:1|1:4798535_T_C:440,30,0:4798535	1|1:0,6:6:18:1|1:4798535_T_C:270,18,0:4798535	1/1:0,2:2:6:.:.:84,6,0:.1|1:0,12:12:36:1|1:4798535_T_C:540,36,0:4798535	1/1:0,1:1:3:.:.:42,3,0:.	0/0:5,0:5:0:.:.:0,0,0:.	1|1:0,6:6:18:1|1:4798535_T_C:270,18,0:4798535	0/0:4,0:4:0:.:.:0,0,0:.	1|1:0,11:11:33:1|1:4798535_T_C:495,33,0:4798535	1|1:0,13:13:39:1|1:4798535_T_C:585,39,0:4798535	1/1:0,4:4:12:.:.:167,12,0:.	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:360,24,0:4798535	1/1:0,4:4:12:.:.:167,12,0:.	1|1:0,11:11:33:1|1:4798535_T_C:463,33,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:360,24,0:4798535	1|1:0,11:11:33:1|1:4798535_T_C:470,33,0:4798535	1/1:0,5:5:15:.:.:209,15,0:.	1|1:0,7:7:21:1|1:4798535_T_C:302,21,0:4798535	1/1:0,11:11:33:.:.:459,33,0:.	1/1:0,6:6:18:.:.:251,18,0:.	1|1:0,13:13:39:1|1:4798535_T_C:547,39,0:4798535	1/1:0,3:3:9:.:.:125,9,0:.	1|1:0,8:8:24:1|1:4798535_T_C:340,24,0:4798535	1|1:0,15:15:45:1|1:4798535_T_C:675,45,0:4798535	1/1:0,5:5:15:.:.:209,15,0:.	1|1:0,12:12:36:1|1:4798535_T_C:540,36,0:4798535	1/1:0,7:7:21:.:.:292,21,0:.	1|1:0,8:8:24:1|1:4798535_T_C:350,24,0:4798535	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:360,24,0:4798535

```{}
bcftools view GCRF_merged.bcf Scaffold_28__1_contigs__length_11240829:4798538
bcftools view GCRF_merged_impute4.1.sort.rm_dup.bcf Scaffold_28__1_contigs__length_11240829:4798538
```
```{r}
weird.snp <- read.table("GCRF-Present/R data/testing", sep="\t")
View(weird.snp)

```

So from this I'm getting a couple things:
It has a very high quality score (28042.20)
The frequency of the Alt Allele (G) here is actually really high, which is weird because in the regression most of the individuals are homozygous reference
Total depth is 629? Which I also don't totally get because my average depth per individual was like 7x or something.
A positive inbreeding score, suggesting an excess of homozygotes (which is what we're seeing) and departure from HWE
Mapping quality if 60, which seems good.
THE QD (quality by depth) is also really higher, also suggesting the variant is a true positive

The stats for the one heterozygous individual are:
0|1:1,4:5:26:0|1:4798535_T_C:165,0,26:4798535

So the depth of the first allele isn't very much, but the second is good. Total read depth is good. Genotype quality is 26, which is good. 

So I'm going to go ahead and say that generally, I trust these calls which makes it seem like there actually is only one heterozygote for this snp, which means something cool might actually be going on.

I'm trying the regression from above as a quadratic instead to see if it fits better. It seems like the R-squared values are a bit higher but the confidence interval is getting a lot wider... not sure what the threshold for "better" is here.

```{r, S28.feather_quad}
# Fit quadratic regression model
S28.feather_quad <- lm(PHENOTYPE ~ GENOTYPE + I(GENOTYPE^2), data = S28.1.4798539.feather)

# Summary of the quadratic regression model
summary(S28.feather_quad)

# Removing NA values
S28.1.4798539.feather.no.na <- na.omit(S28.1.4798539.feather)

# Create a data frame for results
S28.feather.results <- data.frame(FID = as.factor(S28.1.4798539.feather.no.na$FID),
                                   GENOTYPE = as.numeric(S28.1.4798539.feather.no.na$GENOTYPE),
                                   PHENOTYPE = as.numeric(S28.1.4798539.feather.no.na$PHENOTYPE))

# Add predicted values from the quadratic regression model
S28.feather.results$predicted <- predict(S28.feather_quad)

# Define breaks and colors for plotting
y_breaks <- pretty(range(S28.feather.results$PHENOTYPE), n = 5)
x_breaks <- seq(min(S28.feather.results$GENOTYPE), max(S28.feather.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

# Plotting with quadratic regression line
S28.4798539.2 <- ggplot(S28.feather.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = TRUE, color = "grey3") +  # Add quadratic trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle("Feather Morphology vs SNP S28.4798539") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("AA", "AG", "GG")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values = group_colors) +  # Customize legend labels
  theme_classic()

# Display the plot
print(S28.4798539.2)

ggsave("S28.4798539_quad.png", plot=S28.4798539, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r}
# Summary of linear regression model
summary(S28.feather)
#Call:
#lm(formula = PHENOTYPE ~ GENOTYPE, data = S28.1.4798539.feather)

#Residuals:
#    Min      1Q  Median      3Q     Max 
#-4.2729 -0.6994  0.0462  0.7834  4.1695 

#Coefficients:
#            Estimate Std. Error t value Pr(>|t|)    
#(Intercept)   0.3120     0.1586   1.967   0.0529 .  
#GENOTYPE     -1.6949     0.2423  -6.995 9.48e-10 ***
#---
#Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1

#Residual standard error: 1.31 on 75 degrees of freedom
#  (12 observations deleted due to missingness)
#Multiple R-squared:  0.3948,	Adjusted R-squared:  0.3868 
#F-statistic: 48.93 on 1 and 75 DF,  p-value: 9.482e-10

######################################
# Summary of quadratic regression model
summary(S28.feather_quad)

#Call:
#lm(formula = PHENOTYPE ~ GENOTYPE + I(GENOTYPE^2), data = S28.1.4798539.feather)

#Residuals:
#    Min      1Q  Median      3Q     Max 
#-4.2538 -0.6802  0.0654  0.6641  4.1886 

#Coefficients:
#              Estimate Std. Error t value Pr(>|t|)  
#(Intercept)     0.2929     0.1556   1.883   0.0637 .
#GENOTYPE        3.6205     2.5861   1.400   0.1657  
#I(GENOTYPE^2)  -2.6936     1.3050  -2.064   0.0425 *
#---
#Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1

#Residual standard error: 1.283 on 74 degrees of freedom
#  (12 observations deleted due to missingness)
#Multiple R-squared:  0.4278,	Adjusted R-squared:  0.4123 
#F-statistic: 27.66 on 2 and 74 DF,  p-value: 1.072e-09
```
