---
title: "GCRF_GWAS_Workflow"
output:
  html_document:
    toc: true
    toc_float: true
date: "2024-01-04"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Desktop/GCRF/")
```

# INTRODUCTION
Here I will be going through the steps for running a GWAS on a variety of morphological traits. I'm starting with vcf files that have been filtered as follows
-only biallilic alleles
-only SNPs
-fraction of missing genotypes is <25%
-min allele freq is 0.05
-max allele freq is .95
-depth greater than or equal to 4

# PRE-GWAS WORKFLOW
## Workflow post GATK issues
I: *HC_HC_genotypes_unfiltered.vcf.gz

O: *_mk_dup0.vcf


###################
Merging all the mkdp0 files together into one merged vcf file to do the downstream filtering with
I: *_mk_dup0.vcf.gz

picard MergeVcfs -I 1_mk_dup0.vcf.gz -I 2_mk_dup0.vcf.gz -I 3_mk_dup0.vcf.gz -I 4_mk_dup0.vcf.gz -I 5_mk_dup0.vcf.gz -I 6_mk_dup0.vcf.gz -I 7_mk_dup0.vcf.gz -I 8_mk_dup0.vcf.gz -I 9_mk_dup0.vcf.gz -I $

O: GCRF_merged_mkdp0.vcf



###################
Checking the number of lines
cat GCRF_merged_mkdp0.vcf | wc -l
16429366


###################
Initial filtering
I: GCRF_merged_mkdp0.vcf

vcf=GCRF_merged_mkdp0.vcf

out_basic=GCRF.mkdp0_merged_25missing_min.5max.95.vcf.gz

bcftools view -m 2 -M 2 --type snps -i 'F_MISSING < 0.25' --min-af 0.05 --max-af 0.95 ${vcf} -O z -o ../../../called_genotypes/filtered/${out_basic}
bcftools index ../../../called_genotypes/filtered/${out_basic}
bcftools stats ../../../called_genotypes/filtered/${out_basic} > ../../../called_genotypes/filtered/${out_basic}.stats

out_depth=../../../called_genotypes/filtered/GCRF.mkdp0_merged_25missing_min.5max.95_depth4.vcf.gz
bcftools filter -i 'INFO/DP >= 4' ../../../called_genotypes/filtered/${out_basic} -O z -o ${out_depth}
bcftools index ${out_depth}
bcftools stats ${out_depth} > ${out_depth}.stats

O: GCRF.mkdp0_merged_25missing_min.5max.95_depth4.vcf.gz

###################
Checking lines of the filtered files
zcat ../../../called_genotypes/filtered/GCRF.mkdp0_merged_25missing_min.5max.95_depth4.vcf.gz | wc -l
7497657

## 1. VQSR filtering
This is done after marking the bases that has zero depth and were being called as homozygous reference (a stupid thing that gatk 4.0+ does). 

This filtering is to remove errors in genotypes due to sequencing issues. The script I was given actually wants this to be run on an already merged vcf file, but when I tried this the filtering went fine but downstream the imputation kept failing due to a weird error. So I regrouped and did the filtering on each scaffold individually. This was implemented with the vqsr_filt.sh script, which had the following filters:

-filter "QUAL < 30.0" --filter-name "QUAL30" \
-filter "SOR > 3.0" --filter-name "SOR3" \
-filter "FS > 60.0" --filter-name "FS60" \
-filter "MQ < 40.0" --filter-name "MQ40" \
-filter "QD < 2.0" --filter-name "QD2" \
-filter "MQRankSum < -12.5" --filter-name "MQRankSum-12.5" \
-filter "ReadPosRankSum < -8.0" --filter-name "ReadPosRankSum-8"

The last two, MQRankSum and ReadPosRankSum didn't seem to apply, the error was something along the lines of the information not being present which didn't make sense since the header was there, but I opted to move on.

Once all the scaffolds were filtered, I moved the files to a new folder for imputation.

## 2. BEAGLE 4.1 Imputation
I merged the GCRF_merged file and filtered that. Then I applied BEAGLE4.1 to impute missing genotypes. This was done with the script impute.sh.

```{}
BEAGLE_JAR="/projects/ericacnr@colostate.edu/mambaforge/envs/bioinf/java/beagle.27Jan18.7e1.jar"
j=GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr.vcf

java -Xmx180G -jar $BEAGLE_JAR \
gt="$j" \
out=GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr.vcf.gl_impute4.1.vcf.gz \
nthreads=48
```

The log had this as the final output: 
Number of markers:             7487128
Total time for building model: 2 hours 35 minutes 59 seconds
Total time for sampling:       42 minutes 24 seconds
Total run time:                3 hours 28 minutes 15 seconds

mv GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr.vcf.gl_impute4.1.vcf.gz GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr_gl_impute4.1.vcf.gz

Just renamed the output file a little.

#### didn't end up doing all following stuff this the second time around, saving for prosperity. The line number of the imputed file post mkdp0 filtering is 7487139 which is perfect.
Which suggests it was running the correct number of SNPs (7942760) however upon inspecting the file there was over 9mil lines, which doesn't make sense and suggested that some of the markers (~2mil) were being duplicated. Additionally, there was an error with the imputed files where the Chromosome position was missing in the header. So to fix that issue first, I cropped the header from the original merged file and added it to the imputed file.

```{}
(bioinf) [ericacnr@colostate.edu@login12 impute]$ zcat GCRF_merged_impute4.1.vcf.gz |grep -v "^#" > GCRF_merged_impute4.1.vcf_nohead
(bioinf) [ericacnr@colostate.edu@login12 impute]$ zcat GCRF_merged.vcf.gz | grep "^#" > header.txt
(bioinf) [ericacnr@colostate.edu@login12 impute]$ cat GCRF_merged_impute4.1.vcf_nohead | wc -l
9943613
(bioinf) [ericacnr@colostate.edu@login12 impute]$ cat header.txt GCRF_merged_impute4.1.vcf_nohead | bgzip -c > GCRF_merged_imputed4.1_newhead.vcf.gz
```

I then attempted to apply the bcftools norms function on this corrected file...

```{}
(bioinf) [ericacnr@colostate.edu@login12 impute]$ bcftools norm -d none -f ../../Reference/leucosticte_australis_final_assembly.fasta  GCRF_merged_imputed4.1_newhead.vcf.gz > GCRF_merged_impute4.1.rm_dup.vcf.gz
Lines   total/split/joined/realigned/skipped:	9943613/0/0/0/0
(bioinf) [ericacnr@colostate.edu@login12 impute]$ cat GCRF_merged_impute4.1.rm_dup.vcf.gz | grep -v "^#" | wc -l
9943500
```

This still left me with too many lines, but I tried to apply plink anyway... it threw an error though, something about split chromosomes?

```{}
(GWAS) [ericacnr@colostate.edu@login12 impute]$ plink --vcf GCRF_merged_imputed4.1_newhead.vcf.gz --aec --recode --out GCRF_newhead
PLINK v1.90b6.21 64-bit (19 Oct 2020)          www.cog-genomics.org/plink/1.9/
(C) 2005-2020 Shaun Purcell, Christopher Chang   GNU General Public License v3
Logging to GCRF_newhead.log.
Options in effect:
  --allow-extra-chr
  --out GCRF_newhead
  --recode
  --vcf GCRF_merged_imputed4.1_newhead.vcf.gz

15884 MB RAM detected; reserving 7942 MB for main workspace.
--vcf: GCRF_newhead-temporary.bed + GCRF_newhead-temporary.bim +
GCRF_newhead-temporary.fam written.
Error: .bim file has a split chromosome.  Use --make-bed by itself to
remedy this.
```

So I tried sorting the vcf file.

```{}
(bioinf) [ericacnr@colostate.edu@login12 impute]$ bcftools sort GCRF_merged_imputed4.1_newhead.vcf.gz > GCRF_merged_imputed4.1_newhead.sort.vcf
Writing to /tmp/bcftools.zBleDc
Merging 7 temporary files
Cleaning
Done
```

This sorting works! So doing the bcftools norm again on the sorted file.

```{}
(GWAS) [ericacnr@colostate.edu@login12 impute]$ conda activate bioinf
(bioinf) [ericacnr@colostate.edu@login12 impute]$ bcftools norm -d none -f ../../Reference/leucosticte_australis_final_assembly.fasta  GCRF_merged_imputed4.1_newhead.sort.vcf > GCRF_merged_impute4.1.sort.rm_dup.vcf.gz
Lines   total/split/joined/realigned/skipped:	9943613/0/0/0/0
(bioinf) [ericacnr@colostate.edu@login12 impute]$ cat GCRF_merged_impute4.1.sort.rm_dup.vcf.gz | grep -v "^#" |wc -l
7942760
(bioinf) [ericacnr@colostate.edu@login12 impute]$ mv GCRF_merged_impute4.1.sort.rm_dup.vcf.gz GCRF_merged_impute4.1.sort.rm_dup.vcf
```

## 3. PLINK to make .ped and .map files

Now jumping into plink. Working is a whole new folder because I don't want to lose anything from the older one. new_input_files

```{}
$conda activate GWAS2
$plink --vcf GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr_gl_impute4.1.vcf.gz --aec --recode --out ../new_input_files/base_files
PLINK v1.90b6.21 64-bit (19 Oct 2020)          www.cog-genomics.org/plink/1.9/
(C) 2005-2020 Shaun Purcell, Christopher Chang   GNU General Public License v3
Logging to ../new_input_files/base_files.log.
Options in effect:
  --allow-extra-chr
  --out ../new_input_files/base_files
  --recode
  --vcf GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr_gl_impute4.1.vcf.gz

15884 MB RAM detected; reserving 7942 MB for main workspace.
--vcf: ../new_input_files/base_files-temporary.bed +
../new_input_files/base_files-temporary.bim +
../new_input_files/base_files-temporary.fam written.
7487128 variants loaded from .bim file.
89 people (0 males, 0 females, 89 ambiguous) loaded from .fam.
Ambiguous sex IDs written to ../new_input_files/base_files.nosex .
Using 1 thread (no multithreaded calculations invoked).
Before main variant filters, 89 founders and 0 nonfounders present.
Calculating allele frequencies... done.
7487128 variants and 89 people pass filters and QC.
Note: No phenotypes present.
--recode ped to ../new_input_files/base_files.ped +
../new_input_files/base_files.map ... done.
```

## 4. Fixing the .map files and the .ped files to have morphology

Fixing the .map file so that the SNP name is the scaffold and the position together. This file shows the chromosome and SNP location for each SNP.

```{}
$ cd ../new_input_files/
$ cat base_files.map | head
Scaffold_1__1_contigs__length_151072562	.	0	751
Scaffold_1__1_contigs__length_151072562	.	0	2701
Scaffold_1__1_contigs__length_151072562	.	0	3064
Scaffold_1__1_contigs__length_151072562	.	0	4683
Scaffold_1__1_contigs__length_151072562	.	0	4696
Scaffold_1__1_contigs__length_151072562	.	0	5001
Scaffold_1__1_contigs__length_151072562	.	0	5009

$ cat base_files.map | awk '{print$1"\t"$1"."$4"\t"$3"\t"$4}' > GCRF_base_file.map_loc.map
$ cat GCRF_base_file.map_loc.map | head
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.751	0	751
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.2701	0	2701
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.3064	0	3064
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.4683	0	4683
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.4696	0	4696
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.5001	0	5001
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.5009	0	5009
```

Now I'm going to make a separate file for the genotype information from the .ped file for downstream stuff

```{}
cut -f7- GCRF_base_file_fin.ped > GCRF_base_file_genotypes.txt
awk '{print NF; exit}' GCRF_base_file_genotypes.txt
14974256
```

great!

#### don't end up needing to fix the .ped file or plug these base files directly into plink because I will rather be making a new .ped file for each morphology and running plink of that and the proper .map file to make my GWAS inputs.
The .ped file is the FID, IID, (other stuff 3-5), phenotype data, genotypes. I need to replace the first column in the .ped file to have the FID be the location. So the plan for this is to make a separate .txt file in R with the location and the sample names, then add it to the original ped file to replace the first two columns.

According to Marina: "In your ped file, this means the first columns should be all "1"s, the second column should be your sample ID, then 3 columns which are usually "0"s, and finally, the 6th column is your phenotype column"

I'm going to keep the FID as WMNT=1 and PIPA=2 so that there's some distinction, and see if this will run in GEMMA.

First lets make the .txt file, we'll do that here.
```{r}
morpho <- read.csv("~/Desktop/GCRF/R data/GCRF_morphology.csv")
morpho <- morpho[1:189,]
locID <- as.data.frame(cbind(morpho$site, morpho$BGP_ID))
colnames(locID) <- c("site", "BGP_ID")

#getting the correct sample IDs
locID <- distinct(locID)
locID <- locID[order(locID$BGP_ID), ]
rownames(locID) <- NULL
locID <- locID[38:128,]
rownames(locID) <- NULL
locID <- locID[c(-15,-18),]

#fixing the site data to be numerical
locID <- replace(locID, locID == "WMNT", 1)
locID <- replace(locID, locID == "PIPA", 2)

write.table(locID, file="locID_num.txt", row.names=F, col.names=F)
```

Alright, I moved the .txt file into the folder with the other input files. Now I'm going to replace the first column of the .ped file with the first column of the locID_num.txt. file. I have to remove the " characters first though...

```{}
#removing the " from the file
sed 's/"//g' locID_num.txt > locID_num_clean.txt
#adding the locID to the .ped file as the first two columns
paste locID_num_clean.txt base_files.ped > GCRF_intermed.ped
#reformatting the file so it's the first two columns from the locID file and the rest are from the original file, this takes a while
cat GCRF_intermed.ped | awk 'BEGIN {OFS="\t"} {printf "%s\t%s\t%s", $1, $2, $5; for (i=6; i<=14974264; i++) printf "\t%s", $i; print ""}' > GCRF_base_file_fin.ped
#14974264 is the number of rows in the intermed file, so it's saying take just the 1st, 2nd, and 5th columns and then everything else (all the genotypes)
#checking that it went correctly (-S command keeps it from wrapping, so it can show the full line without going crazy and feeding out all the genotypes onto the screen)
less -S GCRF_base_file_fin.ped
#checking that it has the correct number of columns, looks good!
awk '{print NF; exit}' GCRF_base_file_fin.ped
>14974262
```



## 5. PLINK for .bim .bam .fam files
Okay so now that the .map file and .ped file are (hopefully) correctly formatted, I'm going to try and run plink to get the files I need to input into the GEMMA program.

So I'm going to use these base files without the morphology to make base .bim .bam and .fam files. I think I can then edit the .fam file to have the phenotypes information since I haven't been able to get plink's --pheno options to work in the past.

```{}
plink --ped GCRF_base_file_fin.ped --map GCRF_base_file.map_loc.map --out GCRF_base_file --make-bed --aec --allow-no-sex
```
Awesome, now I have these files:

GCRF_base_file.bed #can't view this as it's the binary file
GCRF_base_file.bim #this has the correct SNP ID's
GCRF_base_file.fam #this has all the info and -9 values for the morphology, so this is what I'll fiddle with...

Okay so I'm going to take these base files and the GCRF_bill_morpho.txt and copy them into the bill_morpho folder so I can work with them within there. I'm going to go ahead and delete everything that was in the folder as I'm starting from scratch. 

I'm going to make a GCRF_bill_morpho.txt file that is formatted just like the .fam file but will all the morphology, then I'm going to rename it and see if I can just use that for the mlmm.

Adding all the morphology I want for the beak stuff into one file to use as a .fam files. 

```{r}
morpho <- read.csv("~/Desktop/GCRF/R data/GCRF_morphology.csv")
#getting rid of the empty cells that came from excel
morpho <- morpho[1:189,]

bill_morpho <- as.data.frame(cbind(morpho$BGP_ID, morpho$nare_length, morpho$beak_depth, morpho$beak_width, morpho$culmen_end_length))
bill_morpho <- bill_morpho[order(bill_morpho$V1), ]
rownames(bill_morpho) <- NULL
bill_morpho <- bill_morpho[48:145,]
rownames(bill_morpho) <- NULL
bill_morpho <- bill_morpho[c(-15,-18,-20,-36,-43,-59,-69,-70,-76,-89,-90),]
rownames(bill_morpho) <- NULL
#missing phenotype data but have genetic data, will impute with GEMMA
bill_morpho <- rbind(bill_morpho, c("23N00557", NA, NA, NA, NA), c("23N00459", NA, NA, NA, NA))
bill_morpho <- bill_morpho[order(bill_morpho$V1), ]
rownames(bill_morpho) <- NULL
#Now I've got the right number of rows! Next, replacing the "-" with NAs as this will export best
bill_morpho <- replace(bill_morpho, bill_morpho == "-", NA)
bill_morpho <- replace(bill_morpho, bill_morpho == "", NA)

FID_file <- read.table("~/Desktop/GCRF/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))
#adding columns for the empty values that are normally in the .ped file
FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
#adding the bill_length as the 6th row, skipping tarsus for now
bill_ped_file <- cbind(FID_file[,1:5], bill_morpho[,2:5])
write.table(bill_ped_file, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/GCRF_bill_morpho.txt", row.names=F, col.names=F)
```

Moving that file over onto the cluster and cleaning it up. Then renamed it to GCRF_bill_morpho.fam

```{}
sed 's/"//g' GCRF_bill_morpho.txt > GCRF_bill_morpho_clean.txt
```

# RUNNING THE GWAS
## DOING THIS FOR EACH TRAIT SEPARATELY 6. Relatedness Matrix

First step is to create a relatedness matrix that can be used for all of the following GWAS analyses.

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_base_file -gk 1 -o GCRF.relate > gemma.relate.err 2> gemma.relate.out
```

Okay so that kept failing because:
ERROR: Enforce failed for number of analyzed individuals equals 0. in src/param.cpp at line 2075 in ProcessCvtPhen

Seems like this might be because the phenotype file doesn't have anything.

So I just copies the relatedness matrix from the bill_length stuff from earlier.

## SKIPPING THESE 7. MLMM models
So Marina got eigen values first and included those in the mlmm so I'm going to do that also.

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho \
	-k output/GCRF_relate.cXX.txt \
	-eigen -o GCRF_bill_morpho_eigen
```

Easy enough. I now have:
GCRF_bill_morpho_eigen.eigenD.txt
GCRF_bill_morpho_eigen.eigenU.txt
GCRF_bill_morpho_eigen.log.txt

Okay so now the files are ready to run a mlmm but there are missing phenotypes. I'm going to just run it with those missing phenotypes just being ignored.

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho -k output/GCRF_relate.cXX.txt -d output/GCRF_bill_morpho.eigen.eigenD.txt -u output/GCRF_bill_morpho.eigen.eigenU.txt -lmm 1 -o GCRF_bill_morpho.mlmm -n 1 2 3 4 -w 500000 -s 5000000
```

Got this error:
GSL ERROR: matrix is singular in lu.c at line 449 errno 1

Which might be due to one of the variables being highly correlated with another, probably beak width and beak length. So dropping beak width as a variable. Putting this into a script so it had enough memory and stuff.

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho -k output/GCRF_relate.cXX.txt -d output/GCRF_bill_morpho.eigen.eigenD.txt -u output/GCRF_bill_morpho.eigen.eigenU.txt -lmm 1 -o GCRF_bill_morpho.mlmm -n 1 3 4 -w 500000 -s 5000000
```

This has been running for 20 hours now, I only allotted 24 hours total so hopefully it completes soon, but this is the farthest I've gotten with this!

#########################################
I'm going to prep the feather morphology stuff. I've already make a data frame with the residuals after a body size correction and that's what I'll put into the GWAS. So I'm going to fix that file so that it can be used at the .fam file in the GWAS.

```{r}
FID_file <- read.table("~/Desktop/GCRF/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))
#adding columns for the empty values that are normally in the .ped file
FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
colnames(FID_file) <- c("FID", "ID", "V3", "V4", "V5")

feather_res_gwas <- read.table("~/Desktop/GCRF/GenomicPipeline/GWAS input/feather_residuals.txt", header=T, sep=" ")
feather_res_gwas <- data.frame(FID_file[,1:5], feather_res_gwas)
write.table(feather_res_gwas, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/feather_residuals_loc.txt", row.names=F, col.names=F)
missing_ids <- setdiff(feather_res_gwas$ID, FID_file$ID)
```

```{}
sed 's/"//g' feather_residuals_loc.txt > feather_residuals_clean.txt

mv GCRF_base_file.bed GCRF_feather.bed
mv GCRF_base_file.bim GCRF_feather.bim
mv feather_residuals_clean.txt GCRF_feather.fam
```

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_feather \
	-k output/GCRF_relate.cXX.txt \
	-eigen -o GCRF_feather_eigen
```

GCRF_feather_eigen.eigenD.txt 
GCRF_feather_eigen.eigenU.txt
GCRF_feather_eigen.log.txt

Testing this outside of a script to make sure it runs. Looks like it didn't throw the correlation error, so putting it into a script and letting it cook!
```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_feather -k\
output/GCRF_relate.cXX.txt -d output/GCRF_feather_eigen.eigenD.txt \
-u output/GCRF_feather_eigen.eigenU.txt -lmm 1 \
-o GCRF_bill_morpho.mlmm -n 1 2 3 4 5 -w 500000 -s 5000000
```
Gosh, seems like this is running too! I'm just going to adjust the time to a day and a half...

So that ran but after processing it said I have ~257,000 significant SNPs, so I don't trust it at all. So back to lmm and bslmm models for individuals phenotypes I guess, what a bummer.



## 8. BSLMM and LMM
```{r, message=F, LIBRARIES}
library(MASS)
library(caret)
library(ggplot2)
library(dplyr)
library(FactoMineR)
library(factoextra)
library(corrplot)
library(ggcorrplot)
library(rstatix)
library(dplyr)
```
### Feather PCA and BSLMM
What I can do for the feather stuff at least is to run a PCA with the feather morphology and then use the first PC as the phenotype in the GWAS, that was it's one variable that is still accounting for all of the feather morphology...

This whole first part is just to get the site ID information arranged correctly so I can plug it into the feather data more easily. Just ripped from the GCRF_lms.rmd file.
```{r}
morpho <- read.csv("~/Desktop/GCRF/GCRF-Present/R data/GCRF_morphology.csv")
morpho <- morpho[1:189,]

morpho_num <- as.data.frame(cbind(as.numeric(morpho$wing), as.numeric(morpho$tarsus), as.numeric(morpho$weight), as.numeric(morpho$tail_length), as.numeric(morpho$nare_length), as.numeric(morpho$bristle_length), as.numeric(morpho$beak_depth), as.numeric(morpho$beak_width), as.numeric(morpho$nare_end_length), as.numeric(morpho$culmen_end_length)))

colnames(morpho_num) <- c("wing", "tarsus", "weight", "tail_length", "nare_length", "bristle_length", "beak_depth", "beak_width", "nare_end_length", "culmen_end_length")
morpho2 <- as.data.frame(cbind(morpho$BGP_ID, as.factor(morpho$site),morpho_num))
colnames(morpho2) <- c("BGP_ID", "site", "wing", "tarsus", "weight", "tail_length", "nare_length", "bristle_length", "beak_depth", "beak_width", "nare_end_length", "culmen_end_length")

morpho_avg <- morpho2 %>% group_by(BGP_ID, site) %>% reframe(wing=mean(wing), tarsus=mean(tarsus), weight=mean(weight), tail_length=mean(tail_length), nare_length=mean(nare_length), bristle_length=mean(bristle_length), beak_depth=(beak_depth), beak_width=mean(beak_width), nare_end_length=mean(nare_end_length), culmen_end_length=mean(culmen_end_length))
```

```{r}
feather <- read.csv("~/Desktop/GCRF/GCRF-Present/R data/GCRF_feather_morpho.csv")
site <- data.frame(morpho_avg$site, morpho_avg$BGP_ID, morpho_avg$tarsus)
colnames(site) <- c("site", "ID", "tarsus")

feather_avg <- feather %>% group_by(ID) %>% summarize(pen_length=mean(Penn_length), pen_num=mean(Penn_number), plum_length=mean(Plum_length), plum_number=mean(Plum_number), nodes=mean(Nodes))

feather_avg <- merge(feather_avg, site, by="ID")
#reordering the columns so they're tidy
feather_avg <- data.frame(feather_avg$ID, feather_avg$site, feather_avg$pen_length, feather_avg$pen_num, feather_avg$plum_length, feather_avg$plum_number, feather_avg$nodes, feather_avg$tarsus)
colnames(feather_avg) <- c("BGP_ID", "site", "pen_length", "pen_num", "plum_length", "plum_number", "nodes", "tarsus")
```

```{r}
#feather_avg <- feather_avg[!duplicated(feather_avg$ID),]

#removing and NAs, only losing 3 individuals
feather_avg <- na.omit(feather_avg)
colSums(is.na(feather_avg))
#including the tarsus variable to maybe account for body size differences?
morpho_num <- feather_avg[,3:8]

corr_m <- cor(morpho_num)
ggcorrplot(corr_m)
#none of them have a value higher than |.53|, so I'm going to say that any correlation is fine
#actually going to remove the tarsus because its so weakly correlated with everything...
morpho <- feather_avg[,3:7]
```

```{r}
feather_avg #the df with individuals (character) and groups (factor)
morpho #the df with just the variables
#nothing has been normalized

#taking the active variable, ie individuals and the measurements that go into the PCA
morpho.active <- feather_avg[,c(1, 3:7)]

#now need to standardize the variables, but using the PCA function that's already done so don't need to do it prior to running the PCA in this case, otherwise use scale()... x-mean(x)/sd(x)
#need to remove the column for individual though...
morpho.active.1 <- morpho.active[,-1]
morpho.pca <- PCA(morpho.active.1,  scale.unit=T, graph=F)

eig.val <- get_eig(morpho.pca)
eig.val
fviz_eig(morpho.pca, addlabels = T)
#this is awesome, the first PC explains over half the variation which suggests it'll be a representative variable to run in the GWAS
```

```{r}
var <- get_pca_var(morpho.pca)
var$coord

```

```{r}

fviz_pca_var(morpho.pca, col.var="black")
fviz_cos2(morpho.pca, choice="var", axes=1:4)
library("corrplot")
corrplot(var$contrib, is.corr=F)
#so I'm intrepreting this as PC1 representing the structure overall, as opposed to one variable more strongly than any others
```

```{r}
#fviz_pca_ind(morpho.pca)
#fviz_pca_ind(morpho.pca, col.ind = "cos2", 
#             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#             repel = TRUE # Avoid text overlapping (slow if many points)
#             )
```

```{r}
fviz_pca_ind(morpho.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = feather_avg$site, # color by groups
             palette = c("#00AFBB", "#E7B800"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups",
             )
ind <- morpho.pca$ind
coord <- as.data.frame(ind$coord)
PC1 <- data.frame(feather_avg$BGP_ID, coord$Dim.1)
colnames(PC1) <- c("BGP_ID", "PC1")

```

Awesome, so now that I have that set up, I can add the missing BGP_ID's and there are some random duplicates...

```{r}
PC1 <- PC1[c(-17, -31, -47, -60, -73, -74),]
feather_PC1 <- rbind(PC1,
                     c("23N00444", NA),
                     c("23N00447", NA), 
                     c("23N00448", NA), 
                     c("23N00450", NA), 
                     c("23N00451", NA), 
                     c("23N00452", NA),
                     c("23N00454", NA), 
                     c("23N00456", NA), 
                     c("23N00457", NA), 
                     c("23N00459", NA), 
                     c("23N00462", NA), 
                     c("23N00557", NA))
#missing phenotype data but have genetic data, will impute with GEMMA
feather_PC1 <- feather_PC1[order(feather_PC1$BGP_ID), ]
rownames(feather_PC1) <- NULL
```

```{r}
FID_file <- read.table("~/Desktop/GCRF/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))
#adding columns for the empty values that are normally in the .ped file
FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
#adding the bill_length as the 6th row, skipping tarsus for now
feather.PC_ped_file <- cbind(FID_file[,1:5], feather_PC1[,2])
write.table(feather.PC_ped_file, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/GCRF_feather_PC.txt", row.names=F, col.names=F)
```

```{}
sed 's/"//g' GCRF_feather_PC.txt > GCRF_feather_PC_clean.txt
paste GCRF_feather_PC_clean.txt ../GCRF_base_file_genotypes.txt > GCRF_feather_PC.ped
awk '{print NF; exit}' GCRF_feather_PC.ped
14974262
#great! That's the number we want

plink --ped GCRF_feather_PC.ped --map ../GCRF_base_file.map_loc.map --out GCRF_feather_PC --make-bed --aec --allow-no-sex 
```

Generate a relatedness matrix.

```{}
acompile
module load gcc/10.3.0
module load openmpi
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_feather_PC -gk 1 -o GCRF_relate_PC.cXX.txt
```

```{}
#in a script
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_feather_PC \
-k output/GCRF_relate_PC.cXX.txt -bslmm 1 -o GCRF_feather_PC.bslmm \
-w 500000 -s 5000000

```
GCRF_feather_PC.bslmm.gamma.txt
GCRF_feather_PC.bslmm.param.txt
GCRF_feather_PC.bslmm.hyp.txt
GCRF_feather_PC.bslmm.log.txt

Processed this with GEMMA_bslmm_output_analysis_bill.R

### Bill PCA and BSLMM
Repeating the steps from the feather PCA but with the bill data.

```{r}
morpho <- read.csv("~/Desktop/GCRF/GCRF-Present/R data/GCRF_morphology.csv")
morpho <- morpho[1:189,]

morpho_num <- as.data.frame(cbind(as.numeric(morpho$nare_length), as.numeric(morpho$beak_depth), as.numeric(morpho$beak_width), as.numeric(morpho$culmen_end_length)))

colnames(morpho_num) <- c("nare_length", "beak_depth", "beak_width", "culmen_end_length")
morpho2 <- as.data.frame(cbind(morpho$BGP_ID, as.factor(morpho$site), morpho_num))
colnames(morpho2) <- c("BGP_ID", "site", "nare_length", "beak_depth", "beak_width", "culmen_end_length")

morpho_avg <- morpho2 %>% group_by(BGP_ID, site) %>% reframe(nare_length=mean(nare_length), beak_depth=(beak_depth), beak_width=mean(beak_width), culmen_end_length=mean(culmen_end_length))
morpho_avg <- morpho_avg[48:145,]
morpho_avg <- morpho_avg[c(-15,-18,-19,-43,-60,-70,-75,-88,-89),]
```

```{r}
#feather_avg <- feather_avg[!duplicated(feather_avg$ID),]

#removing and NAs, losing 14 individuals
morpho_avg <- na.omit(morpho_avg)
colSums(is.na(morpho_avg))
morpho_num <- morpho_avg[,3:6]


corr_m <- cor(morpho_num)
corr_m
ggcorrplot(corr_m)
```

```{r}
morpho_avg #the df with individuals (character) and groups (factor)
morpho_num #the df with just the variables
#nothing has been normalized

#taking the active variable, ie individuals and the measurements that go into the PCA
morpho.active <- morpho_avg[,c(1, 3:6)]

#now need to standardize the variables, but using the PCA function that's already done so don't need to do it prior to running the PCA in this case, otherwise use scale()... x-mean(x)/sd(x)
#need to remove the column for individual though...
morpho.active.1 <- morpho.active[,-1]
morpho.pca <- PCA(morpho.active.1,  scale.unit=T, graph=F)

eig.val <- get_eig(morpho.pca)
eig.val
fviz_eig(morpho.pca, addlabels = T)
#this is awesome, the first PC explains over half the variation which suggests it'll be a representative variable to run in the GWAS
```

```{r}
var <- get_pca_var(morpho.pca)
var$coord
```

```{r}
fviz_pca_var(morpho.pca, col.var="black")
fviz_cos2(morpho.pca, choice="var", axes=1:4)
library("corrplot")
corrplot(var$contrib, is.corr=F)
#so I'm intrepreting this as PC1 representing the beak morphology overall, as opposed to one variable more strongly than any others
```

```{r}
fviz_pca_ind(morpho.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = morpho_avg$site, # color by groups
             palette = c("#00AFBB", "#E7B800"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups",
             )
ind <- morpho.pca$ind
coord <- as.data.frame(ind$coord)
PC1 <- data.frame(morpho_avg$BGP_ID, coord$Dim.1)
colnames(PC1) <- c("BGP_ID", "PC1")

```

Awesome, so now that I have that set up, I can add the missing BGP_ID's and there are some random duplicates...

```{r}
bill_PC1 <- rbind(PC1,
                  c("23N00437", NA),
                  c("23N00444", NA), 
                  c("23N00448", NA), 
                  c("23N00460", NA), 
                  c("23N00464", NA), 
                  c("23N00467", NA),
                  c("23N00475", NA), 
                  c("23N00479", NA), 
                  c("23N00480", NA), 
                  c("23N00485", NA), 
                  c("23N00488", NA),
                  c("23N00489", NA), 
                  c("23N00490", NA),
                  c("23N00491", NA))
#missing phenotype data but have genetic data, will impute with GEMMA
bill_PC1 <- bill_PC1[order(bill_PC1$BGP_ID), ]
rownames(bill_PC1) <- NULL
```

```{r}
FID_file <- read.table("~/Desktop/GCRF/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))
#adding columns for the empty values that are normally in the .ped file
FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
#adding the bill_length as the 6th row, skipping tarsus for now
bill.PC_ped_file <- cbind(FID_file[,1:5], bill_PC1[,2])
write.table(bill.PC_ped_file, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/GCRF_bill_PC.txt", row.names=F, col.names=F)
```

Moving it over to the cluster...

```{}
sed 's/"//g' GCRF_bill_PC.txt > GCRF_bill_PC_clean.txt
paste GCRF_bill_PC_clean.txt ../GCRF_base_file_genotypes.txt > GCRF_bill_PC.ped
awk '{print NF; exit}' GCRF_bill_PC.ped

plink --ped GCRF_bill_PC.ped --map ../GCRF_base_file.map_loc.map --out GCRF_bill_PC --make-bed --aec --allow-no-sex 
```
Generate relatedness matrix
```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_PC -gk 1 -o GCRF_relate_PC.cXX.txt
```

```{}
#in a script
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_PC \
-k output/GCRF_relate_PC.cXX.txt -bslmm 1 -o GCRF_bill_PC.bslmm \
-w 500000 -s 5000000

```
GCRF_feather_PC.bslmm.gamma.txt
GCRF_feather_PC.bslmm.param.txt
GCRF_feather_PC.bslmm.hyp.txt
GCRF_feather_PC.bslmm.log.txt

## misc stuff


Running this in a script so it doesn't time out or run out of space. Using one that Marina wrote, again not sure about the -w and -s numbers...


```{}
module load gcc/10.3
module load openmpi
conda activate GWAS2

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho \
-k output/GCRF_relate.cXX.txt -bslmm 1 -o GCRF_bill_morpho_bslmm \
-w 500000 -s 5000000
```

Okay so that seems to have worked! I now have:
GCRF_bill_morpho_bslmm.bv.txt     
GCRF_bill_morpho_bslmm.hyp.txt  
GCRF_bill_morpho_bslmm.param.txt
GCRF_bill_morpho_bslmm.gamma.txt  
GCRF_bill_morpho_bslmm.log.txt


Now I'm going to run each individual feather trait in the bslmm, and do a PCA for bill morphology and run those traits through bslmm too. Oh boy! Might try a lmm on the feather PCA too just to see if I get anything.









########################################################################################################################

So from this I should be able to predict the missing values.

```{}
module load gcc/10.3
module load openmpi
conda activate GWAS2

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile GCRF_bill_morpho -epm output/GCRF_bill_morpho_bslmm.param.txt -emu output/GCRF_bill_morpho_bslmm.log.txt -ebv output/GCRF_bill_morpho_bslmm.bv.txt -k output/GCRF_bill_length.relate.cXX.txt -predict 1 -o GCRF_bill_morpho_bslmm
```
















## LMM
Setting up a spreadsheet with all of the morphology separately. I am going to run this on the residual results from the linear models to correct for body size...

So here's the thing, I separated the morphology by sex when I did my initial linear regressions, and I found that there wasn't any difference in general slopes or what variables came out as significant (wing chord is the only one for both sexes), so I'm going to go ahead and group everything by sex. Out of caution I'm going to run linear regressions on each variable (not just wing chord) just to cover my bases although I wouldn't say it seems necessary. The annoying thing here is going to be removing NA's for variables that don't have measurements then adding them back in so I can put everything back into a single dataframe for making the .ped files later.

First let's just organize and clean up some stuff.
```{r}
morpho <- read.csv("~/Desktop/GCRF/GCRF-Present/R data/GCRF_morphology.csv", na.strings = c("", "NA"))
morpho <- morpho[1:189,]
morpho <- morpho[order(morpho$BGP_ID), ]
rownames(morpho) <- NULL
morpho <- morpho[(48:145),]
rownames(morpho) <- NULL
morpho <- morpho[!duplicated(morpho$BGP_ID),]
morpho <- morpho[c(-15, -18),]

morpho <- data.frame(morpho$site, morpho$BGP_ID, morpho$tarsus, morpho$wing, morpho$tail_length, morpho$nare_length, morpho$bristle_length, morpho$beak_depth, morpho$beak_width, morpho$culmen_end_length, col.names=T)
colnames(morpho) <- c("site", "BGP_ID", "tarsus", "wing", "tail_length", "nare_length", "bristle_length", "beak_depth", "beak_width", "culmen_end_length")
morpho[morpho == "-"] <- NA
```

```{r}
wing <- data.frame(site = morpho$site, BGP_ID = morpho$BGP_ID, wing = as.numeric(morpho$wing),tarsus = as.numeric(morpho$tarsus))
wing.ID <- wing %>% na.omit() %>% pull(BGP_ID)

tail <- data.frame(site = morpho$site, BGP_ID = morpho$BGP_ID, tail = as.numeric(morpho$tail),tarsus = as.numeric(morpho$tarsus))
tail.ID <- tail %>% na.omit() %>% pull(BGP_ID)

nare <- data.frame(site = morpho$site, BGP_ID = morpho$BGP_ID, nare = as.numeric(morpho$nare_length),tarsus = as.numeric(morpho$tarsus))
nare.ID <- nare %>% na.omit() %>% pull(BGP_ID)

bristle_l <- data.frame(site = morpho$site, BGP_ID = morpho$BGP_ID, bristle_length = as.numeric(morpho$bristle_length),tarsus = as.numeric(morpho$tarsus))
bristle_l.ID <- bristle_l %>% na.omit() %>% pull(BGP_ID)

bill_l <- data.frame(site = morpho$site, BGP_ID = morpho$BGP_ID, bill_length = as.numeric(morpho$culmen_end_length),tarsus = as.numeric(morpho$tarsus))
bill_l.ID <- bill_l %>% na.omit() %>% pull(BGP_ID)

bill_d <- data.frame(site = morpho$site, BGP_ID = morpho$BGP_ID, bill_depth = as.numeric(morpho$beak_depth),tarsus = as.numeric(morpho$tarsus))
bill_d.ID <- bill_d %>% na.omit() %>% pull(BGP_ID)

bill_w <- data.frame(site = morpho$site, BGP_ID = morpho$BGP_ID, bill_width = as.numeric(morpho$beak_width),tarsus = as.numeric(morpho$tarsus))
bill_w.ID <- bill_w %>% na.omit() %>% pull(BGP_ID)
```

```{r}
wing.lm <- lm(as.numeric(wing) ~ as.numeric(tarsus), data=wing)
summary(wing.lm)
ggplot(wing, aes(x=as.numeric(wing), y=as.numeric(tarsus)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(wing.lm$residuals)
wing.res <- data.frame(wing.lm$model,wing.lm$residuals,wing.ID)
colnames(wing.res) <- c("wing", "tarsus", "wing.res", "BGP_ID")
wing.res.all <- full_join(wing, wing.res, by=c("BGP_ID", "wing", "tarsus"))

tail.lm <- lm(as.numeric(tail) ~ as.numeric(tarsus), data=tail)
summary(tail.lm)
ggplot(tail, aes(x=as.numeric(tail), y=as.numeric(tarsus)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(tail.lm$residuals)
tail.res <- data.frame(tail.lm$model,tail.lm$residuals,tail.ID)
colnames(tail.res) <- c("tail", "tarsus", "tail.res", "BGP_ID")
tail.res.all <- full_join(tail, tail.res, by=c("BGP_ID", "tail", "tarsus"))

nare.lm <- lm(as.numeric(nare) ~ as.numeric(tarsus), data=nare)
summary(nare.lm)
ggplot(nare, aes(x=as.numeric(tarsus), y=as.numeric(nare)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(nare.lm$residuals)
nare.res <- data.frame(nare.lm$model,nare.lm$residuals,nare.ID)
colnames(nare.res) <- c("nare", "tarsus", "nare.res", "BGP_ID")
nare.res.all <- full_join(nare, nare.res, by=c("BGP_ID", "nare", "tarsus"))

bristle_l.lm <- lm(as.numeric(bristle_length) ~ as.numeric(tarsus), data=bristle_l)
summary(bristle_l.lm)
ggplot(bristle_l, aes(x=as.numeric(tarsus), y=as.numeric(bristle_length)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bristle_l.lm$residuals)
bristle_l.res <- data.frame(bristle_l.lm$model,bristle_l.lm$residuals,bristle_l.ID)
colnames(bristle_l.res) <- c("bristle_length", "tarsus", "bristle_l.res", "BGP_ID")
bristle_l.res.all <- full_join(bristle_l, bristle_l.res, by=c("BGP_ID", "bristle_length", "tarsus"))

bill_l.lm <- lm(as.numeric(bill_length) ~ as.numeric(tarsus), data=bill_l)
summary(bill_l.lm)
ggplot(bill_l, aes(x=as.numeric(tarsus), y=as.numeric(bill_length)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_l.lm$residuals)
bill_l.res <- data.frame(bill_l.lm$model,bill_l.lm$residuals,bill_l.ID)
colnames(bill_l.res) <- c("bill_length", "tarsus", "bill_l.res", "BGP_ID")
bill_l.res.all <- full_join(bill_l, bill_l.res, by=c("BGP_ID", "bill_length", "tarsus"))

bill_d.lm <- lm(as.numeric(bill_depth) ~ as.numeric(tarsus), data=bill_d)
summary(bill_d.lm)
ggplot(bill_d, aes(x=as.numeric(tarsus), y=as.numeric(bill_depth)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_d.lm$residuals)
bill_d.res <- data.frame(bill_d.lm$model,bill_d.lm$residuals,bill_d.ID)
colnames(bill_d.res) <- c("bill_depth", "tarsus", "bill_d.res", "BGP_ID")
bill_d.res.all <- full_join(bill_d, bill_d.res, by=c("BGP_ID", "bill_depth", "tarsus"))

bill_w.lm <- lm(as.numeric(bill_width) ~ as.numeric(tarsus), data=bill_w)
summary(bill_w.lm)
ggplot(bill_w, aes(x=as.numeric(tarsus), y=as.numeric(bill_width)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_w.lm$residuals)
bill_w.res <- data.frame(bill_w.lm$model,bill_w.lm$residuals,bill_w.ID)
colnames(bill_w.res) <- c("bill_width", "tarsus", "bill_w.res", "BGP_ID")
bill_w.res.all <- full_join(bill_w, bill_w.res, by=c("BGP_ID", "bill_width", "tarsus"))
```

```{r}
body.residuals <- data.frame(morpho$site, morpho$BGP_ID, wing.res.all$wing.res, tail.res.all$tail.res, nare.res.all$nare.res, bristle_l.res.all$bristle_l.res, bill_l.res.all$bill_l.res, bill_d.res.all$bill_d.res, bill_w.res.all$bill_w.res, morpho$tarsus)
colnames(body.residuals) <- c("site", "BGP_ID", "wing.res", "tail.res", "nare.res", "bristle_lenght.res", "beak_lenght.res", "beak_depth.res", "beak_width.res", "tarsus")
write.csv(file="GCRF-Present/GenomicPipeline/GWAS input/GCRF_morphology_residuals.csv", body.residuals, col.names = T, row.names = F)
```

### This is without the linear regressions
```{r}
morpho <- read.csv("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_morphology_residuals.csv")

FID_file <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/locID_num.txt", header = FALSE, sep = " ", na.strings = c("", "NA"))

FID_file$V3 <- 0
FID_file$V4 <- 0
FID_file$V5 <- 0
#adding the bill_length as the 6th row, skipping tarsus for now
morpho_ped_file <- data.frame(FID_file, morpho[,c(3:10)])
write.table(morpho_ped_file, file="~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_morpho_residuals.txt", row.names=F, col.names=F)
```

```{}
sed 's/"//g' GCRF_morpho_residuals.txt > GCRF_morpho_clean.txt

cut -d ' ' -f 1-6 GCRF_morpho_clean.txt > wing
cut -d ' ' -f 1-5,13 GCRF_morpho_clean.txt > tarsus
cut -d ' ' -f 1-5,7 GCRF_morpho_clean.txt > tail_length
cut -d ' ' -f 1-5,8 GCRF_morpho_clean.txt > nare_length
cut -d ' ' -f 1-5,9 GCRF_morpho_clean.txt > bristle_length
cut -d ' ' -f 1-5,11 GCRF_morpho_clean.txt > beak_depth
cut -d ' ' -f 1-5,12 GCRF_morpho_clean.txt > beak_width
cut -d ' ' -f 1-5,10 GCRF_morpho_clean.txt > culmen_end_length

paste wing ../GCRF_base_file_genotypes.txt > wing.ped
paste tarsus ../GCRF_base_file_genotypes.txt > tarsus.ped
paste tail_length ../GCRF_base_file_genotypes.txt > tail_length.ped
paste nare_length ../GCRF_base_file_genotypes.txt > nare_length.ped
paste bristle_length ../GCRF_base_file_genotypes.txt > bristle_length.ped
paste beak_depth ../GCRF_base_file_genotypes.txt > beak_depth.ped
paste beak_width ../GCRF_base_file_genotypes.txt > beak_width.ped
paste culmen_end_length ../GCRF_base_file_genotypes.txt > culmen_end_length.ped

#checked column numbers for all of these and everything checks out!
```

```{}
plink --ped wing.ped --map ../GCRF_base_file.map_loc.map --out wing.d/wing --make-bed --aec --allow-no-sex

plink --ped tarsus.ped --map ../GCRF_base_file.map_loc.map --out tarsus.d/tarsus --make-bed --aec --allow-no-sex

plink --ped tail_length.ped --map ../GCRF_base_file.map_loc.map --out tail.d/tail_length --make-bed --aec --allow-no-sex

plink --ped nare_length.ped --map ../GCRF_base_file.map_loc.map --out nare.d/nare_length --make-bed --aec --allow-no-sex

plink --ped bristle_length.ped --map ../GCRF_base_file.map_loc.map --out bristle.d/bristle_length --make-bed --aec --allow-no-sex

plink --ped beak_depth.ped --map ../GCRF_base_file.map_loc.map --out beak_d.d/beak_depth --make-bed --aec --allow-no-sex

plink --ped beak_width.ped --map ../GCRF_base_file.map_loc.map --out beak_w.d/beak_width --make-bed --aec --allow-no-sex

plink --ped culmen_end_length.ped --map ../GCRF_base_file.map_loc.map --out culmen.d/culmen_end_length --make-bed --aec --allow-no-sex

```

```{}
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile wing.d/wing -gk 1 -o relate
#just creating one relatedness matrix for this folder that they can all pull from
/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile wing -k ../output/relate.cXX.txt -lmm 1 -o wing_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile tarsus -k ../output/relate.cXX.txt -lmm 1 -o tarsus_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile tail_length -k ../output/relate.cXX.txt -lmm 1 -o tail_length_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile nare_length -k ../output/relate.cXX.txt -lmm 1 -o nare_length_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile bristle_length -k ../output/relate.cXX.txt -lmm 1 -o bristle_length_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile beak_depth -k ../output/relate.cXX.txt -lmm 1 -o beak_depth_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile beak_width -k ../output/relate.cXX.txt -lmm 1 -o beak_width_lmm -w 500000 -s 5000000

/projects/ericacnr@colostate.edu/GEMMA/bin/gemma -bfile culmen_end_length -k ../output/relate.cXX.txt -lmm 1 -o culmen_end_length_lmm -w 500000 -s 5000000
```

# GETTING ALLELE FREQUENCIES FOR THE FLAGGED SNPS

## BSLMM SNPs
So with filtering using the bslmm output for feathers and bills, I was able to identify some relevant SNPS. With the first iteration I got a lot of stuff flagging as relevant. The second (correct) iteration, not as much.

### first iteration
For the bill morphology: (above a .1 threshold for PIP values, per Chaves 2016).
chr	rs	ps	n_miss	alpha	beta	gamma	eff
Scaffold_2__1_contigs__length_114774641	Scaffold_2__1_contigs__length_114774641.41203916	41203916	0	-6.108376e-06	-0.8424626	0.114384	0.0963642420384

For the feather morphology:
chr	rs	ps	n_miss	alpha	beta	gamma	eff
Scaffold_28__1_contigs__length_11240829	Scaffold_28__1_contigs__length_11240829.4798538	4798538	0	-3.014029e-06	-1.487668	0.389188	0.578982533584
Scaffold_1__1_contigs__length_151072562	Scaffold_1__1_contigs__length_151072562.91081692	91081692	0	-2.156909e-06	-1.643631	0.25956	0.42662086236
Scaffold_18__1_contigs__length_25274884	Scaffold_18__1_contigs__length_25274884.2074884	2074884	0	2.328089e-06	0.9979339	0.23866	0.238166904574

### second iteration
feather PC bslmm:
Threshold of 0.01 which is typical but not quite as cool.
ZFCHROM	ZFPOS	CHROM	mean.loc	sca.ori	length	rs	POS	n_miss	alpha	beta	gamma
6	10477013	Scaffold_8__1_contigs__length_35685229	18106371.1019173	-1	35685229	*Scaffold_8__1_contigs__length_35685229.25471973*	25471973	0	-3.820215e-06	-1.778742	0.039932
1.1	61341218	Scaffold_6__1_contigs__length_71876329	37081312.5403133	1	71876329	*Scaffold_6__1_contigs__length_71876329.60198070*	60198070	0	1.661721e-06	1.221628	0.03012
1	24356364	Scaffold_2__1_contigs__length_114774641	59305973.9656377	-1	114774641	*Scaffold_2__1_contigs__length_114774641.92336930*	92336930	0	-3.689599e-06	-1.475477	0.024338
4	5065576	Scaffold_5__1_contigs__length_71959612	34675573.4638522	-1	71959612	*Scaffold_5__1_contigs__length_71959612.65589803*	65589803	0	-5.10239e-06	-1.161962	0.022772
2	93419280	Scaffold_1__1_contigs__length_151072562	40329379.5914754	1	151072562	*Scaffold_1__1_contigs__length_151072562.128626182*	128626182	0	-2.930171e-06	-1.766972	0.02032
2	111387037	Scaffold_1__1_contigs__length_151072562	40329379.5914754	1	151072562	*Scaffold_1__1_contigs__length_151072562.146593939*	146593939	0	2.650819e-06	1.039644	0.017154
1	36058968	Scaffold_2__1_contigs__length_114774641	59305973.9656377	-1	114774641	*Scaffold_2__1_contigs__length_114774641.80634326*	80634326	0	-3.089245e-06	-1.881733	0.011898
3	514041	Scaffold_3__1_contigs__length_112408489	56640462.8173333	-1	112408489	*Scaffold_3__1_contigs__length_112408489.112330666*	112330666	0	-2.940688e-06	-1.968629	0.010818

for the bill PC bslmm, with the same threshold. 
ZFCHROM	ZFPOS	CHROM	mean.loc	sca.ori	length	rs	POS	n_miss	alpha	beta	gamma	eff
2	18925575	Scaffold_1__1_contigs__length_151072562	40329379.5914754	1	151072562	*Scaffold_1__1_contigs__length_151072562.54132477*	54132477	0	-5.036987e-06	-1.285083	0.093498	0.120152690334
5	31292946	Scaffold_7__1_contigs__length_62267870	30554770.2958274	1	62267870	*Scaffold_7__1_contigs__length_62267870.31872111*	31872111	0	-3.124287e-06	-0.9874937	0.014814	0.0146287316718
1	31504261	Scaffold_2__1_contigs__length_114774641	59305973.9656377	-1	114774641	*Scaffold_2__1_contigs__length_114774641.85189033*	85189033	0	-4.471717e-06	-0.5889336	0.012078	0.0071131400208
13	840405	Scaffold_23__1_contigs__length_18116493	8513441.31332796	1	18116493	*Scaffold_23__1_contigs__length_18116493.1385210*	1385210	0	-2.590394e-06	-1.239186	0.011302	0.014005280172
2	103663091	Scaffold_1__1_contigs__length_151072562	40329379.5914754	1	151072562	*Scaffold_1__1_contigs__length_151072562.138869993*	138869993	0	-3.084669e-06	-0.5464172	0.011268	0.0061570290096
1	13464638	Scaffold_2__1_contigs__length_114774641	59305973.9656377	-1	114774641	*Scaffold_2__1_contigs__length_114774641.103228656*	103228656	0	-3.151422e-06	-0.3542868	0.01125	0.0039857265
2	72985722	Scaffold_1__1_contigs__length_151072562	40329379.5914754	1	151072562	*Scaffold_1__1_contigs__length_151072562.108192624*	108192624	0	-3.304588e-06	-0.9396421	0.011168	0.0104939229728
13	1905123	Scaffold_23__1_contigs__length_18116493	8513441.31332796	1	18116493	*Scaffold_23__1_contigs__length_18116493.2449928*	2449928	0	3.487147e-06	0.4015281	0.010922	0.0043854899082
4	56410849	Scaffold_5__1_contigs__length_71959612	34675573.4638522	-1	71959612	*Scaffold_5__1_contigs__length_71959612.14244530*	14244530	0	1.907588e-06	0.6763174	0.01029	0.006959306046

So the goal now is to use plink to pull out these specific snps, get recode the alleles to 0 (homozygous reference), 1 (heterozygous), and 2 (homozygous alternative) for each individual. The idea then is to run a regression with those categorical variables again the continuous bill of feather PC values and see if there's a relationship.

So first, we're going to be pull the SNP info from the vcf that was put into the GWAS, so the imputed file: GCRF_merged_impute4.1.sort.rm_dup.vcf. PLINK should be able to pull out a SNP based on it's name, but because the SNPS are namemd after the scaffold and then given a position, I'm not sure how to extract the one I'm looking for. 

This is how the SNP is stored in the vcf file:
Scaffold_292__1_contigs__length_11690	8856	.	A	G	.	PASS	.	GT	1|1	0|0	0|0	0|0	0|1	0|1	0|0	1|1	1|1	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|1	0|1	1|1	0|0	0|1	0|0	1|1	0|1	0|0	1|1	0|1	1|1	0|0	0|0	0|0	0|1	0|0	0|0	0|1	0|0	0|1	0|0	0|0	0|0	0|0	0|0	0|1	0|0	1|1	0|0	0|0	0|1	0|0	0|0	0|0	0|1	0|1	0|0	0|0	0|0	1|1	0|0	0|1	0|0	1|1	1|1	0|0	0|0	0|0	0|0	1|1	0|0	0|0	0|0	0|0	1|1	0|0	0|0	0|1	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0

### Using PLINK to extract the values
So the names are all there, just gotta figure out the right command for plink...

```{}
plink --vcf ../impute_new/GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr_gl_impute4.1.vcf.gz --aec --recode --out allele_freq

cat allele_freq.map | awk '{print$1"\t"$1"."$4"\t"$3"\t"$4}' > allele_freq_loc.map
mv allele_freq_loc.map allele_freq.map

plink --ped allele_freq.ped --map allele_freq.map --out allele_freq --make-bed --aec --allow-no-sex

plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.54132477 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.54132477.bill
```

That worked!! I now have a .fam (.raw) file that has the info on the genotype! So I can import that here, merge it with the phenotype info and there ya go. Sick. So I'm just gonna do that again real quick with the other bill SNPs...

```{}
plink --bfile allele_freq --snp Scaffold_7__1_contigs__length_62267870.31872111 --aec --recodeA --out snps/Scaffold_7__1_contigs__length_62267870.31872111.bill
plink --bfile allele_freq --snp Scaffold_2__1_contigs__length_114774641.85189033 --aec --recodeA --out snps/Scaffold_2__1_contigs__length_114774641.85189033.bill
plink --bfile allele_freq --snp Scaffold_23__1_contigs__length_18116493.1385210 --aec --recodeA --out snps/Scaffold_23__1_contigs__length_18116493.1385210.bill
plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.138869993 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.138869993.bill
plink --bfile allele_freq --snp Scaffold_2__1_contigs__length_114774641.103228656 --aec --recodeA --out snps/Scaffold_2__1_contigs__length_114774641.103228656.bill
plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.108192624 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.108192624.bill
plink --bfile allele_freq --snp Scaffold_23__1_contigs__length_18116493.2449928 --aec --recodeA --out snps/Scaffold_23__1_contigs__length_18116493.2449928.bill
plink --bfile allele_freq --snp Scaffold_5__1_contigs__length_71959612.14244530 --aec --recodeA --out snps/Scaffold_5__1_contigs__length_71959612.14244530.bill
```
And the feather SNPs

```{}
plink --bfile allele_freq --snp Scaffold_8__1_contigs__length_35685229.25471973 --aec --recodeA --out snps/Scaffold_8__1_contigs__length_35685229.25471973.feather
plink --bfile allele_freq --snp Scaffold_6__1_contigs__length_71876329.60198070 --aec --recodeA --out snps/Scaffold_6__1_contigs__length_71876329.60198070.feather
plink --bfile allele_freq --snp Scaffold_5__1_contigs__length_71959612.65589803 --aec --recodeA --out snps/Scaffold_5__1_contigs__length_71959612.65589803.feather
plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.128626182 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.128626182.feather
plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.146593939 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.146593939.feather
plink --bfile allele_freq --snp Scaffold_2__1_contigs__length_114774641.80634326 --aec --recodeA --out snps/Scaffold_2__1_contigs__length_114774641.80634326.feather
plink --bfile allele_freq --snp Scaffold_3__1_contigs__length_112408489.112330666 --aec --recodeA --out snps/Scaffold_3__1_contigs__length_112408489.112330666.2074884.feather
```

### Importing data into R
#### old stuff
Moving back into R after I moved the files over, going to organizing everything here.
```{r}
feather.PC_ped_file <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_feather_PC.txt")
colnames(feather.PC_ped_file) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE")
bill.PC_ped_file <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_bill_PC.txt")
colnames(bill.PC_ped_file) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE")


S1.1.91081692.feather <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_1__1_contigs__length_151072562.91081692.feather.raw", sep = " ", header = T)
S1.1.91081692.feather <- data.frame(feather.PC_ped_file, S1.1.91081692.feather[,7])
colnames(S1.1.91081692.feather) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")

S18.1.2074884.feather <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_18__1_contigs__length_25274884.2074884.feather.raw", sep = " ", header = T)
S18.1.2074884.feather <- data.frame(feather.PC_ped_file, S18.1.2074884.feather[,7])
colnames(S18.1.2074884.feather) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")

S28.1.4798539.feather <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_28__1_contigs__length_11240829.4798538.feather.raw", sep = " ", header = T)
S28.1.4798539.feather <- data.frame(feather.PC_ped_file, S28.1.4798539.feather[,7])
colnames(S28.1.4798539.feather) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")

S2.1.41203916.bill <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_2__1_contigs__length_114774641.41203916.bill.raw", sep = " ", header = T)
S2.1.41203916.bill <- data.frame(bill.PC_ped_file, S2.1.41203916.bill[,7])
colnames(S2.1.41203916.bill) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")

```

#### new stuff
Taking advantage of this setback by putting everything into for loops.
```{r, getting data set up}
feather_file_names <- list.files(path="GCRF-Present/GenomicPipeline/data/allele_freq/feather/", pattern = ".feather.raw$", full.names = TRUE)
  feather_data_list <- lapply(feather_file_names, function(file) {
  read.table(file, header = TRUE, sep = " ")
})
bill_file_names <- list.files(path="GCRF-Present/GenomicPipeline/data/allele_freq/bill/", pattern = ".bill.raw$", full.names = TRUE)
bill_data_list <- lapply(bill_file_names, function(file) {
  read.table(file, header = TRUE, sep = " ")
})

feather.PC_ped_file <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_feather_PC.txt")
colnames(feather.PC_ped_file) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE")
bill.PC_ped_file <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_bill_PC.txt")
colnames(bill.PC_ped_file) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE")

feather_ped_files <- list()
for (i in seq_along(feather_data_list)){
  df <- feather_data_list[[i]]
  col_name <- colnames(df)[7]
  ped_file <- data.frame(feather.PC_ped_file, df[,7])
  colnames(ped_file) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")
  feather_ped_files[[col_name]] <- ped_file
}

bill_ped_files <- list()
for (i in seq_along(bill_data_list)){
  df <- bill_data_list[[i]]
  col_name <- colnames(df)[7]
  ped_file <- data.frame(bill.PC_ped_file, df[,7])
  colnames(ped_file) <- c("FID", "IID", "PAT", "MAT", "SEX", "PHENOTYPE", "GENOTYPE")
  bill_ped_files[[col_name]] <- ped_file
}
```

### Running and plotting linear regressions
#### old stuff
Going to try running these in a linear regression, there are some NA values that might throw errors...

```{r, S1.1.91081692}
library(ggplot2)
S1.feather <- lm(PHENOTYPE~GENOTYPE, data=S1.1.91081692.feather)
summary(S1.feather)
#p=5.19e-06!!!
S1.1.91081692.feather.no.na <- na.omit(S1.1.91081692.feather)

S1.feather.results <- data.frame(FID=as.factor(S1.1.91081692.feather.no.na$FID), GENOTYPE = as.numeric(S1.1.91081692.feather.no.na$GENOTYPE), PHENOTYPE = as.numeric(S1.1.91081692.feather.no.na$PHENOTYPE))
S1.feather.results$predicted <- predict(S1.feather)

y_breaks <- pretty(range(S1.feather.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S1.feather.results$GENOTYPE), max(S1.feather.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S1.91081692 <- ggplot(S1.feather.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle("Feather Morphology vs SNP S1.91081692") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("GG", "GA", "AA")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S1.91081692.png", plot=S1.91081692, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r, S18.1.2074884}
S18.feather <- lm(PHENOTYPE~GENOTYPE, data=S18.1.2074884.feather)
summary(S18.feather)
#p=0.001585!!!
S18.1.2074884.feather.no.na <- na.omit(S18.1.2074884.feather)

S18.feather.results <- data.frame(FID=as.factor(S18.1.2074884.feather.no.na$FID), GENOTYPE = as.numeric(S18.1.2074884.feather.no.na$GENOTYPE), PHENOTYPE = as.numeric(S18.1.2074884.feather.no.na$PHENOTYPE))
S18.feather.results$predicted <- predict(S18.feather)

y_breaks <- pretty(range(S18.feather.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S18.feather.results$GENOTYPE), max(S18.feather.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S18.2074884 <- ggplot(S18.feather.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle("Feather Morphology vs SNP S18.2074884") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("CC", "CT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S18.2074884.png", plot=S18.2074884, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r, S28.1.4798539}
S28.feather <- lm(PHENOTYPE~GENOTYPE, data=S28.1.4798539.feather)
summary(S28.feather)
#p=9.482e-10!!!
S28.1.4798539.feather.no.na <- na.omit(S28.1.4798539.feather)

S28.feather.results <- data.frame(FID=as.factor(S28.1.4798539.feather.no.na$FID), GENOTYPE = as.numeric(S28.1.4798539.feather.no.na$GENOTYPE), PHENOTYPE = as.numeric(S28.1.4798539.feather.no.na$PHENOTYPE))
S28.feather.results$predicted <- predict(S28.feather)

y_breaks <- pretty(range(S28.feather.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S28.feather.results$GENOTYPE), max(S28.feather.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S28.4798539 <- ggplot(S28.feather.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle("Feather Morphology vs SNP S28.4798539") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("AA", "AG", "GG")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S28.4798539.png", plot=S28.4798539, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)

```

```{r, S2.1.41203916}
S2.bill <- lm(PHENOTYPE~GENOTYPE, data=S2.1.41203916.bill)
summary(S2.bill)
#p=3.949e-09!!!
S2.1.41203916.bill.no.na <- na.omit(S2.1.41203916.bill)

S2.bill.results <- data.frame(FID=as.factor(S2.1.41203916.bill.no.na$FID), GENOTYPE = as.numeric(S2.1.41203916.bill.no.na$GENOTYPE), PHENOTYPE = as.numeric(S2.1.41203916.bill.no.na$PHENOTYPE))
S2.bill.results$predicted <- predict(S2.bill)

y_breaks <- pretty(range(S2.bill.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S2.bill.results$GENOTYPE), max(S2.bill.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S2.41203916 <- ggplot(S2.bill.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Bill Morphology PC1") +  # Add axis labels
  ggtitle("Bill Morphology vs SNP S2.41203916") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("TT", "TG", "GG")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S2.41203916.png", plot=S2.41203916, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)

```


#### new stuff

```{r, linear models}
feather.results_list <- list()
feather.p_values <- data.frame(snp=character(), p_value=numeric())
for (i in seq_along(feather_ped_files)){
  model <- lm(PHENOTYPE~GENOTYPE, data=feather_ped_files[[i]])
  summary_model <- summary(model)
  snp_name <- names(feather_ped_files[i])
  p_value <- summary_model$coefficients["GENOTYPE", "Pr(>|t|)"]
  feather.p_values <- rbind(feather.p_values, data.frame(snp_name ,p_value))
  df.no.na <- na.omit(feather_ped_files[[i]])
  
  df.results <- data.frame(FID=as.factor(df.no.na$FID), GENOTYPE=as.numeric(df.no.na$GENOTYPE), PHENOTYPE=as.numeric(df.no.na$PHENOTYPE))
  df.results$predicted <- predict(model)
  feather.results_list[[snp_name]] <- df.results
}

bill.results_list <- list()
bill.p_values <- data.frame(snp=character(), p_value=numeric())
for (i in seq_along(bill_ped_files)){
  model <- lm(PHENOTYPE~GENOTYPE, data=bill_ped_files[[i]])
  summary_model <- summary(model)
  p_value <- summary_model$coefficients["GENOTYPE", "Pr(>|t|)"]
  snp_name <- names(bill_ped_files[i])
  bill.p_values <- rbind(bill.p_values, data.frame(snp_name,p_value))
  df.no.na <- na.omit(bill_ped_files[[i]])
  
  df.results <- data.frame(FID=as.factor(df.no.na$FID), GENOTYPE=as.numeric(df.no.na$GENOTYPE), PHENOTYPE=as.numeric(df.no.na$PHENOTYPE))
  df.results$predicted <- predict(model)
  bill.results_list[[snp_name]] <- df.results
}
```

```{r, plotting}
for(i in seq_along(feather.results_list)){
  y_breaks <- pretty(range(feather.results_list[[i]]$PHENOTYPE), n=5)
  x_breaks <- seq(min(feather.results_list[[i]]$GENOTYPE), max(feather.results_list[[i]]$GENOTYPE), length.out = 3)
  group_colors <- c("1" = "lightblue3", "2" = "tomato2")

  plot <- ggplot(feather.results_list[[i]], aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
    geom_point() +   # Add data points with color aesthetic
    geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
    labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
    scale_y_continuous(breaks = y_breaks) +
    scale_x_continuous(breaks = x_breaks, labels = c("homozygous reference", "heterozygous", "homozygous alternative")) +
    scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
    theme_classic()
  snp_name <- names(feather.results_list[i])
  plot_name <- paste0(snp_name, "feather.png")
  ggsave(plot_name, plot=plot, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
}

for(i in seq_along(bill.results_list)){
  y_breaks <- pretty(range(bill.results_list[[i]]$PHENOTYPE), n=5)
  x_breaks <- seq(min(bill.results_list[[i]]$GENOTYPE), max(bill.results_list[[i]]$GENOTYPE), length.out = 3)
  group_colors <- c("1" = "lightblue3", "2" = "tomato2")

  plot <- ggplot(bill.results_list[[i]], aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
    geom_point() +   # Add data points with color aesthetic
    geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
    labs(x = "Genotype", y = "Bill Morphology PC1") +  # Add axis labels
    scale_y_continuous(breaks = y_breaks) +
    scale_x_continuous(breaks = x_breaks, labels = c("homozygous reference", "heterozygous", "homozygous alternative")) +
    scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
    theme_classic()
  snp_name <- names(bill.results_list[i])
  plot_name <- paste0(snp_name, "bill.png")
  ggsave(plot_name, plot=plot, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
}
```
## LMM SNPs
### old stuff
So I ran some LMM's on all of the morphological traits separately, and got some interesting results. I ran them first with a Bonferonii correction (0.05/7900000). I then also ran it with the widely accept 5e-8 significant threshold.

Here's what I got:
WING > 1754 SNP @<6.25e-9
TARSUS > 0 SNP @<6.25e-9 > 0 SNP @5e-8
TAIL LENGTH > 1 SNP @<6.25e-9 13 SNP @5e-8
NARE LENGTH > 0 SNP @<6.25e-9 2 SNP @5e-8
BRISTLE LENGTH > 0 SNP @<6.25e-9 0 SNP @5e-8
BEAK WIDTH > 1 SNP @<6.25e-9 2 SNP @5e-8
BEAK DEPTH > 0 SNP @<6.25e-9 0 SNP @5e-8
CULMEN-END LENGTH > 0 SNP @<6.25e-9 5 SNP @5e-8

So I don't think I'm going to try the allele regression method for all 1754 SNPs for the Wing cause that'd be crazy, but doing some for the bill dimensions would be cool because it makes it a bit clearer what's happening compared to the bill PC equivalent. So let's play with those two SNPs for Nare Length and Beak Width.

Nare Length: (ref>alt)
Scaffold_6__1_contigs__length_71876329.70598544 G>T
Scaffold_6__1_contigs__length_71876329.70598546 G>C

These are right next to each other which is interesting. 

Beak Width
Scaffold_1__1_contigs__length_151072562.25482370 C>T
Scaffold_2__1_contigs__length_114774641.86583423 C>T

So I'm going to use the same plink command and pull out a new VCF file.

#### Extracting with PLINK
```{}
plink --bfile allele_freq --snp Scaffold_6__1_contigs__length_71876329.70598544 --aec --recodeA --out snps/Scaffold_6__1_contigs__length_71876329.70598544.nare_length

plink --bfile allele_freq --snp Scaffold_6__1_contigs__length_71876329.70598546 --aec --recodeA --out snps/Scaffold_6__1_contigs__length_71876329.70598546.nare_length

########################
plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.25482370 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.25482370.beak_width

plink --bfile allele_freq --snp Scaffold_2__1_contigs__length_114774641.86583423 --aec --recodeA --out snps/Scaffold_2__1_contigs__length_114774641.86583423.beak_width
```


#### Importing everything into R
Moving stuff over from the cluster

```{r}
morpho.ped <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_morpho.ped")
morpho <- data.frame(morpho.ped$V1, morpho.ped$V2, morpho.ped[,6:19])
colnames(morpho) <- c("FID", "BGP_ID", "wing", "tarsus", "weight", "bill_color", "tail_length", "bristle_cov", "nare_length", "bristle_length", "beak_depth", "beak_width", "tarsus_hair_full", "tarsus_hair_jointdown", "nare_end_length", "culmen_end_length")

#Scaffold_6__1_contigs__length_71876329.70598544 nare length
S6.70598544.nare_length <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_6__1_contigs__length_71876329.70598544.nare_length.raw", sep = " ", header = T)
S6.70598544.nare_length <- data.frame(morpho[,1:2], morpho$nare_length, S6.70598544.nare_length[,7])
colnames(S6.70598544.nare_length) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")

#Scaffold_6__1_contigs__length_71876329.70598546 nare length
S6.70598546.nare_length <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_6__1_contigs__length_71876329.70598546.nare_length.raw", sep = " ", header = T)
S6.70598546.nare_length <- data.frame(morpho[,1:2], morpho$nare_length, S6.70598546.nare_length[,7])
colnames(S6.70598546.nare_length) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")

#Scaffold_1__1_contigs__length_151072562.25482370 beak width
S1.25482370.beak_width <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_1__1_contigs__length_151072562.25482370.beak_width.raw", sep = " ", header = T)
S1.25482370.beak_width <- data.frame(morpho[,1:2], morpho$beak_width, S1.25482370.beak_width[,7])
colnames(S1.25482370.beak_width) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")

#Scaffold_2__1_contigs__length_114774641.86583423 beak width
S2.86583423.beak_width <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_2__1_contigs__length_114774641.86583423.beak_width.raw", sep = " ", header = T)
S2.86583423.beak_width <- data.frame(morpho[,1:2], morpho$beak_width, S2.86583423.beak_width[,7])
colnames(S2.86583423.beak_width) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")
```

```{r, S6.70598544.nare_length}
library(ggplot2)
S6.544.nare <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=S6.70598544.nare_length)
summary(S6.544.nare)
#p=3.417e-08
S6.70598544.nare_length.no.na <- na.omit(S6.70598544.nare_length)

S6.544.nare.results <- data.frame(FID=as.factor(S6.70598544.nare_length.no.na$FID), GENOTYPE = as.numeric(S6.70598544.nare_length.no.na$GENOTYPE), PHENOTYPE = as.numeric(S6.70598544.nare_length.no.na$PHENOTYPE))
S6.544.nare.results$predicted <- predict(S6.544.nare)

y_breaks <- pretty(range(S6.544.nare.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S6.544.nare.results$GENOTYPE), max(S6.544.nare.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S6.70598544 <- ggplot(S6.544.nare.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Nare Length (mm)") +  # Add axis labels
  ggtitle("Nare Length vs SNP S6.70598544") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("GG", "GT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S6.70598544.png", plot=S6.70598544, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r, S6.70598546.nare_length}
S6.546.nare <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=S6.70598546.nare_length)
summary(S6.546.nare)
#p=3.417e-08
S6.70598546.nare_length.no.na <- na.omit(S6.70598546.nare_length)

S6.546.nare.results <- data.frame(FID=as.factor(S6.70598546.nare_length.no.na$FID), GENOTYPE = as.numeric(S6.70598546.nare_length.no.na$GENOTYPE), PHENOTYPE = as.numeric(S6.70598546.nare_length.no.na$PHENOTYPE))
S6.546.nare.results$predicted <- predict(S6.546.nare)

y_breaks <- pretty(range(S6.546.nare.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S6.546.nare.results$GENOTYPE), max(S6.546.nare.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S6.70598546 <- ggplot(S6.546.nare.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Nare Length (mm)") +  # Add axis labels
  ggtitle("Nare Length vs SNP S6.70598546") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("GG", "GC", "CC")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S6.70598546.png", plot=S6.70598546, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

Interestingly, these seem to be identical. I'm going to go ahead and say that's because they're next to each other and completely linked.

```{r, S1.25482370.beak_width}
S1.width <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=S1.25482370.beak_width)
summary(S1.width)
#p=5.87e-09
S1.25482370.beak_width[S1.25482370.beak_width == "-"] <- NA
S1.25482370.beak_width.no.na <- na.omit(S1.25482370.beak_width)

S1.width.results <- data.frame(FID=as.factor(S1.25482370.beak_width.no.na$FID), GENOTYPE = as.numeric(S1.25482370.beak_width.no.na$GENOTYPE), PHENOTYPE = as.numeric(S1.25482370.beak_width.no.na$PHENOTYPE))
S1.width.results$predicted <- predict(S1.width)

y_breaks <- pretty(range(S1.width.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S1.width.results$GENOTYPE), max(S1.width.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S1.25482370 <- ggplot(S1.width.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Beak Width (mm)") +  # Add axis labels
  ggtitle("Beak Width vs SNP S1.25482370") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("CC", "CT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S1.25482370.png", plot=S1.25482370, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r, S2.86583423.beak_width}
S2.width <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=S2.86583423.beak_width)
summary(S2.width)
#p=9.965e-09
S2.86583423.beak_width[S2.86583423.beak_width == "-"] <- NA
S2.86583423.beak_width.no.na <- na.omit(S2.86583423.beak_width)

S2.width.results <- data.frame(FID=as.factor(S2.86583423.beak_width.no.na$FID), GENOTYPE = as.numeric(S2.86583423.beak_width.no.na$GENOTYPE), PHENOTYPE = as.numeric(S2.86583423.beak_width.no.na$PHENOTYPE))
S2.width.results$predicted <- predict(S2.width)

y_breaks <- pretty(range(S2.width.results$PHENOTYPE), n=5)
x_breaks <- seq(min(S2.width.results$GENOTYPE), max(S2.width.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

S2.86583423 <- ggplot(S2.width.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Beak Width (mm)") +  # Add axis labels
  ggtitle("Beak Width vs SNP S2.86583423") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("CC", "CT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S2.86583423.png", plot=S2.86583423  , path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

#### Investigating SNP S28.4798539.feather

This one (Scaffold_28__1_contigs__length_11240829.4798538) is only showing on individual as a heterozygote. I'm going to investigate a bit deeper to see how well we can trust the calling on this to see if this is real, also if this was an imputed SNP or not.

Comparing 
GCRF/GWAS/impute/GCRF_merged.vcf
GCRF/GWAS/impute/GCRF_merged_impute4.1.sort.rm_dup.vcf

```{}
wc -l GCRF_merged.vcf
7953291 GCRF_merged.vcf

wc -l GCRF_merged_impute4.1.sort.rm_dup.vcf
7953293 GCRF_merged_impute4.1.sort.rm_dup.vcf
```

So that's weird, it's suggesting that the imputation only added 2 extra SNPs??

```{}
cat GCRF_merged.vcf | grep "Scaffold_28__1_contigs__length_11240829" | grep "4798538"
```

Scaffold_28__1_contigs__length_11240829	4798538	.	A	G	28042.20	PASS	AC=161;AF=0.904;AN=178;DP=629;ExcessHet=0;FS=0;InbreedingCoeff=0.44;MLEAC=176;MLEAF=0.989;MQ=60;MQRankSum=0;QD=22.95;SOR=0.261	GT:AD:DP:GQ:PGT:PID:PL:PS	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1/1:0,3:3:9:.:.:125,9,0:.	1|1:0,7:7:21:1|1:4798535_T_C:302,21,0:4798535	1/1:0,9:9:27:.:.:376,27,0:.	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1/1:0,9:9:27:.:.:342,27,0:.	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1|1:0,6:6:18:1|1:4798535_T_C:263,18,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:347,24,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:315,21,0:4798535	1/1:0,5:5:15:.:.:209,15,0:.	1|1:0,6:6:18:1|1:4798535_T_C:260,18,0:4798535	1/1:0,7:7:21:.:.:292,21,0:.	0/0:4,0:4:0:.:.:0,0,0:.	1|1:0,7:7:21:1|1:4798535_T_C:309,21,0:4798535	0/0:2,0:2:0:.:.:0,0,0:.	1/1:0,3:3:9:.:.:125,9,0:.	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1/1:0,6:6:18:.:.:227,18,0:.	1/1:0,4:4:12:.:.:167,12,0:.	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1|1:0,10:10:30:1|1:4798535_T_C:428,30,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:309,21,0:4798535	1/1:0,2:2:6:.:.:84,6,0:.	1|1:0,15:15:45:1|1:4798535_T_C:675,45,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:315,21,0:4798535	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1|1:0,5:5:15:1|1:4798535_T_C:225,15,0:4798535	1|1:0,10:10:30:1|1:4798535_T_C:450,30,0:4798535	1|1:0,5:5:15:1|1:4798535_T_C:225,15,0:4798535	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	1/1:0,5:5:15:.:.:209,15,0:.	1|1:0,7:7:21:1|1:4798535_T_C:315,21,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:334,24,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:311,21,0:4798535	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1|1:0,3:3:9:1|1:4798535_T_C:135,9,0:4798535	1/1:0,4:4:12:.:.:167,12,0:.	1|1:0,6:6:18:1|1:4798535_T_C:270,18,0:4798535	1|1:0,13:13:39:1|1:4798535_T_C:582,39,0:4798535	1|1:0,9:9:27:1|1:4798535_T_C:405,27,0:4798535	0/0:2,0:2:0:.:.:0,0,0:.	1|1:0,7:7:21:1|1:4798535_T_C:299,21,0:4798535	1|1:0,11:11:33:1|1:4798535_T_C:495,33,0:4798535	0/0:4,0:4:0:.:.:0,0,0:.	1|1:0,12:12:36:1|1:4798535_T_C:540,36,0:4798535	1|1:0,5:5:15:1|1:4798535_T_C:225,15,0:4798535	*0|1:1,4:5:26:0|1:4798535_T_C:165,0,26:4798535*	1|1:0,12:12:36:1|1:4798535_T_C:540,36,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:360,24,0:4798535	1/1:0,10:10:30:.:.:418,30,0:.	1/1:0,7:7:21:.:.:292,21,0:.	0/0:2,0:2:0:.:.:0,0,0:.	1|1:0,9:9:27:1|1:4798535_T_C:386,27,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:305,21,0:4798535	1|1:0,6:6:18:1|1:4798535_T_C:260,18,0:4798535	1|1:0,7:7:21:1|1:4798535_T_C:315,21,0:4798535	0/0:3,0:3:0:.:.:0,0,0:.	1|1:0,10:10:30:1|1:4798535_T_C:440,30,0:4798535	1|1:0,6:6:18:1|1:4798535_T_C:270,18,0:4798535	1/1:0,2:2:6:.:.:84,6,0:.1|1:0,12:12:36:1|1:4798535_T_C:540,36,0:4798535	1/1:0,1:1:3:.:.:42,3,0:.	0/0:5,0:5:0:.:.:0,0,0:.	1|1:0,6:6:18:1|1:4798535_T_C:270,18,0:4798535	0/0:4,0:4:0:.:.:0,0,0:.	1|1:0,11:11:33:1|1:4798535_T_C:495,33,0:4798535	1|1:0,13:13:39:1|1:4798535_T_C:585,39,0:4798535	1/1:0,4:4:12:.:.:167,12,0:.	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:360,24,0:4798535	1/1:0,4:4:12:.:.:167,12,0:.	1|1:0,11:11:33:1|1:4798535_T_C:463,33,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:360,24,0:4798535	1|1:0,11:11:33:1|1:4798535_T_C:470,33,0:4798535	1/1:0,5:5:15:.:.:209,15,0:.	1|1:0,7:7:21:1|1:4798535_T_C:302,21,0:4798535	1/1:0,11:11:33:.:.:459,33,0:.	1/1:0,6:6:18:.:.:251,18,0:.	1|1:0,13:13:39:1|1:4798535_T_C:547,39,0:4798535	1/1:0,3:3:9:.:.:125,9,0:.	1|1:0,8:8:24:1|1:4798535_T_C:340,24,0:4798535	1|1:0,15:15:45:1|1:4798535_T_C:675,45,0:4798535	1/1:0,5:5:15:.:.:209,15,0:.	1|1:0,12:12:36:1|1:4798535_T_C:540,36,0:4798535	1/1:0,7:7:21:.:.:292,21,0:.	1|1:0,8:8:24:1|1:4798535_T_C:350,24,0:4798535	1|1:0,4:4:12:1|1:4798535_T_C:180,12,0:4798535	1|1:0,8:8:24:1|1:4798535_T_C:360,24,0:4798535

```{}
bcftools view GCRF_merged.bcf Scaffold_28__1_contigs__length_11240829:4798538
bcftools view GCRF_merged_impute4.1.sort.rm_dup.bcf Scaffold_28__1_contigs__length_11240829:4798538
```
```{r}
weird.snp <- read.table("GCRF-Present/R data/testing", sep="\t")
View(weird.snp)

```

So from this I'm getting a couple things:
It has a very high quality score (28042.20)
The frequency of the Alt Allele (G) here is actually really high, which is weird because in the regression most of the individuals are homozygous reference
Total depth is 629? Which I also don't totally get because my average depth per individual was like 7x or something.
A positive inbreeding score, suggesting an excess of homozygotes (which is what we're seeing) and departure from HWE
Mapping quality if 60, which seems good.
THE QD (quality by depth) is also really higher, also suggesting the variant is a true positive

The stats for the one heterozygous individual are:
0|1:1,4:5:26:0|1:4798535_T_C:165,0,26:4798535

So the depth of the first allele isn't very much, but the second is good. Total read depth is good. Genotype quality is 26, which is good. 

So I'm going to go ahead and say that generally, I trust these calls which makes it seem like there actually is only one heterozygote for this snp, which means something cool might actually be going on.

I'm trying the regression from above as a quadratic instead to see if it fits better. It seems like the R-squared values are a bit higher but the confidence interval is getting a lot wider... not sure what the threshold for "better" is here.

```{r, S28.feather_quad}
# Fit quadratic regression model
S28.feather_quad <- lm(PHENOTYPE ~ GENOTYPE + I(GENOTYPE^2), data = S28.1.4798539.feather)

# Summary of the quadratic regression model
summary(S28.feather_quad)

# Removing NA values
S28.1.4798539.feather.no.na <- na.omit(S28.1.4798539.feather)

# Create a data frame for results
S28.feather.results <- data.frame(FID = as.factor(S28.1.4798539.feather.no.na$FID),
                                   GENOTYPE = as.numeric(S28.1.4798539.feather.no.na$GENOTYPE),
                                   PHENOTYPE = as.numeric(S28.1.4798539.feather.no.na$PHENOTYPE))

# Add predicted values from the quadratic regression model
S28.feather.results$predicted <- predict(S28.feather_quad)

# Define breaks and colors for plotting
y_breaks <- pretty(range(S28.feather.results$PHENOTYPE), n = 5)
x_breaks <- seq(min(S28.feather.results$GENOTYPE), max(S28.feather.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

# Plotting with quadratic regression line
S28.4798539.2 <- ggplot(S28.feather.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", formula = y ~ poly(x, 2), se = TRUE, color = "grey3") +  # Add quadratic trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle("Feather Morphology vs SNP S28.4798539") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("AA", "AG", "GG")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values = group_colors) +  # Customize legend labels
  theme_classic()

# Display the plot
print(S28.4798539.2)

ggsave("S28.4798539_quad.png", plot=S28.4798539, path = "~/Desktop/GCRF/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r}
# Summary of linear regression model
summary(S28.feather)
#Call:
#lm(formula = PHENOTYPE ~ GENOTYPE, data = S28.1.4798539.feather)

#Residuals:
#    Min      1Q  Median      3Q     Max 
#-4.2729 -0.6994  0.0462  0.7834  4.1695 

#Coefficients:
#            Estimate Std. Error t value Pr(>|t|)    
#(Intercept)   0.3120     0.1586   1.967   0.0529 .  
#GENOTYPE     -1.6949     0.2423  -6.995 9.48e-10 ***
#---
#Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1

#Residual standard error: 1.31 on 75 degrees of freedom
#  (12 observations deleted due to missingness)
#Multiple R-squared:  0.3948,	Adjusted R-squared:  0.3868 
#F-statistic: 48.93 on 1 and 75 DF,  p-value: 9.482e-10

######################################
# Summary of quadratic regression model
summary(S28.feather_quad)

#Call:
#lm(formula = PHENOTYPE ~ GENOTYPE + I(GENOTYPE^2), data = S28.1.4798539.feather)

#Residuals:
#    Min      1Q  Median      3Q     Max 
#-4.2538 -0.6802  0.0654  0.6641  4.1886 

#Coefficients:
#              Estimate Std. Error t value Pr(>|t|)  
#(Intercept)     0.2929     0.1556   1.883   0.0637 .
#GENOTYPE        3.6205     2.5861   1.400   0.1657  
#I(GENOTYPE^2)  -2.6936     1.3050  -2.064   0.0425 *
#---
#Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1

#Residual standard error: 1.283 on 74 degrees of freedom
#  (12 observations deleted due to missingness)
#Multiple R-squared:  0.4278,	Adjusted R-squared:  0.4123 
#F-statistic: 27.66 on 2 and 74 DF,  p-value: 1.072e-09
```

### new stuff
So I ran some LMM's on all of the morphological traits separately (on the residuals from the tarsus regression to correct for body size), and got some interesting results. I ran them first with a Bonferonii correction (0.05/7500000). I then also ran it with the widely accept 5e-8 significant threshold.

Here's what I got:
TAIL LENGTH 1 SNP @<6.88e-9 1 SNP @5e-8
Scaffold_22__1_contigs__length_19532606.14845898 C    T   4.360509e-09
CULMEN-END LENGTH 0 SNP @<6.88e-9 1 SNP @5e-8
Scaffold_1__1_contigs__length_151072562.39837818 G    A   4.635984e-08

So I'm going to use the same plink command and pull out a new VCF file.

#### Extracting with PLINK
```{}
plink --bfile allele_freq --snp Scaffold_22__1_contigs__length_19532606.14845898 --aec --recodeA --out snps/Scaffold_22__1_contigs__length_19532606.14845898.lmm

plink --bfile allele_freq --snp Scaffold_1__1_contigs__length_151072562.39837818 --aec --recodeA --out snps/Scaffold_1__1_contigs__length_151072562.39837818.lmm

```
#### Importing everything into R
Moving stuff over from the cluster

```{r}
morpho.ped <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/GWAS input/GCRF_morpho_residuals.txt")
morpho <- data.frame(morpho.ped$V1, morpho.ped$V2, morpho.ped[,7], morpho.ped[,10])
colnames(morpho) <- c("FID", "BGP_ID", "tail_length", "culmen_end_length")

#Scaffold_22__1_contigs__length_19532606.14845898 tail length
S22.14845898 <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_22__1_contigs__length_19532606.14845898.lmm.raw", sep = " ", header = T)
tail_length <- data.frame(morpho.ped[,1:2], morpho$tail_length, S22.14845898[,7])
colnames(tail_length) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")

#Scaffold_1__1_contigs__length_151072562.39837818 beak length
S1.39837818 <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq/Scaffold_1__1_contigs__length_151072562.39837818.lmm.raw", sep = " ", header = T)
beak_length <- data.frame(morpho.ped[,1:2], morpho$culmen_end_length, S1.39837818[,7])
colnames(beak_length) <- c("FID", "IID", "PHENOTYPE", "GENOTYPE")
```

```{r}
tail_length_model <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=tail_length)
summary(tail_length_model)
#p=3.865e-09
tail_length.no.na <- na.omit(tail_length)

tail_length.results <- data.frame(FID=as.factor(tail_length.no.na$FID), GENOTYPE = as.numeric(tail_length.no.na$GENOTYPE), PHENOTYPE = as.numeric(tail_length.no.na$PHENOTYPE))
tail_length.results$predicted <- predict(tail_length_model)

y_breaks <- pretty(range(tail_length.results$PHENOTYPE), n=5)
x_breaks <- seq(min(tail_length.results$GENOTYPE), max(tail_length.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

tail_length.plot <- ggplot(tail_length.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Tail Length (mm)") +  # Add axis labels
  ggtitle("Tail Length vs SNP S22.14845898") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("CC", "CT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S22.14845898_tail_length.png", plot=tail_length.plot, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

```{r}
beak_length_model <- lm(as.numeric(PHENOTYPE)~GENOTYPE, data=beak_length)
summary(beak_length_model)
#p=3.865e-09
beak_length.no.na <- na.omit(beak_length)

beak_length.results <- data.frame(FID=as.factor(beak_length.no.na$FID), GENOTYPE = as.numeric(beak_length.no.na$GENOTYPE), PHENOTYPE = as.numeric(beak_length.no.na$PHENOTYPE))
beak_length.results$predicted <- predict(beak_length_model)

y_breaks <- pretty(range(beak_length.results$PHENOTYPE), n=5)
x_breaks <- seq(min(beak_length.results$GENOTYPE), max(beak_length.results$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")

beak_length.plot <- ggplot(beak_length.results, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Beak Length (mm)") +  # Add axis labels
  ggtitle("Beak Length vs SNP S1.39837818") +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("GG", "GA", "AA")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()

ggsave("S1.39837818_beak_length.png", plot=beak_length.plot, path = "~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/allele_freq", width=9, height=7)
```

# IDENTIFYING GENES ASSOCIATED WITH THE SNPS
## Bedtools closest

```{}
zcat GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr_gl_impute4.1.vcf.gz | grep -v "^#" | awk '{print$1"\t"($2-1)"\t"$2}' > GCRF_merged_impute4.1.bed
head GCRF_merged_impute4.1.bed
Scaffold_1__1_contigs__length_151072562	750	751
Scaffold_1__1_contigs__length_151072562	2700	2701
Scaffold_1__1_contigs__length_151072562	3063	3064
Scaffold_1__1_contigs__length_151072562	4682	4683
cat GCRF_merged_impute4.1.bed | awk '{print$1"\t"$2"\t"$3"\t"($1"."$3)}' > GCRF_merged_impute4.1_position.bed
head GCRF_merged_impute4.1_position.bed
Scaffold_1__1_contigs__length_151072562	750	751	Scaffold_1__1_contigs__length_151072562.751
Scaffold_1__1_contigs__length_151072562	2700	2701	Scaffold_1__1_contigs__length_151072562.2701
Scaffold_1__1_contigs__length_151072562	3063	3064	Scaffold_1__1_contigs__length_151072562.3064
Scaffold_1__1_contigs__length_151072562	4682	4683	Scaffold_1__1_contigs__length_151072562.4683
```

Okay so that's making a .bed file, now I need to get this to be at a chromosomal level.... I'll try and do that in R. I used the reoder_VCF functions to take the vcf and remake it. I output that as GCRF_clean.SNP.imuted_gt.ZFchr.bed.

ZFCHROM   ZFPOS   rs    pos-1
1	1918977	Scaffold_2__1_contigs__length_114774641.114774317	114774316
1	1920329	Scaffold_2__1_contigs__length_114774641.114772965	114772964
1	1921417	Scaffold_2__1_contigs__length_114774641.114771877	114771876
1	1922933	Scaffold_2__1_contigs__length_114774641.114770361	114770360

I then reformatted that to be what I need for the .bed file....
cat GCRF_clean.SNP.imuted_gt.ZFchr.bed | awk '{print$1"\t"($2-1)"\t"$2"\t"$3}' > GCRF.SNP.imputed_gt.ZFchr.sorted.bed
ZFCHROM ZFPOS-1 ZFPOS rs
1	1918976	1918977	Scaffold_2__1_contigs__length_114774641.114774317
1	1920328	1920329	Scaffold_2__1_contigs__length_114774641.114772965
1	1921416	1921417	Scaffold_2__1_contigs__length_114774641.114771877
1	1922932	1922933	Scaffold_2__1_contigs__length_114774641.114770361

```{feather.bed}
#Pulling out the specific SNPs I want to look at for this trait
cat GCRF.SNP.imputed_gt.ZFchr.sorted.bed | grep -E "Scaffold_8__1_contigs__length_35685229.25471973|Scaffold_6__1_contigs__length_71876329.60198070|Scaffold_2__1_contigs__length_114774641.92336930|Scaffold_5__1_contigs__length_71959612.65589803|Scaffold_1__1_contigs__length_151072562.128626182|Scaffold_1__1_contigs__length_151072562.146593939|Scaffold_2__1_contigs__length_114774641.80634326|Scaffold_3__1_contigs__length_112408489.112330666" > ../new_input_files/snps/feather_PC_snps.bed

#Converting my annotation file to a .bed file to run in bedtools closest
conda activate bioinf
gzip -dc Taeniopygia_guttata.taeGut3.2.4.98.chr.gtf.gz | gff2bed > Taeniopygia_guttata.bed
bedtools sort -i Taeniopygia_guttata.bed > Taeniopygia_guttata_sorted.bed
bedtools sort -i feather_PC_snps.bed > feather_PC_snps_sorted.bed

#bedtools closest to get the genes
bedtools closest -a feather_PC_snps_sorted.bed -b /scratch/alpine/ericacnr@colostate.edu/GCRF/Reference/ZEFI_annotation/Taeniopygia_guttata_sorted.bed -d > feather_distance.bed
```

```{bill.bed}
cat GCRF.SNP.imputed_gt.ZFchr.sorted.bed | grep -E "Scaffold_1__1_contigs__length_151072562\.54132477|Scaffold_7__1_contigs__length_62267870\.31872111|Scaffold_2__1_contigs__length_114774641\.85189033|Scaffold_23__1_contigs__length_18116493\.1385210|Scaffold_1__1_contigs__length_151072562\.138869993|Scaffold_2__1_contigs__length_114774641\.103228656|Scaffold_1__1_contigs__length_151072562\.108192624|Scaffold_23__1_contigs__length_18116493\.2449928|Scaffold_5__1_contigs__length_71959612\.14244530" > ../new_input_files/snps/bill_PC_snps.bed

bedtools sort -i bill_PC_snps.bed > bill_PC_snps_sorted.bed
bedtools closest -a bill_PC_snps_sorted.bed -b /scratch/alpine/ericacnr@colostate.edu/GCRF/Reference/ZEFI_annotation/Taeniopygia_guttata_sorted.bed -d > bill_distance.bed
```

```{tail.length}
cat GCRF.SNP.imputed_gt.ZFchr.sorted.bed | grep -E "Scaffold_22__1_contigs__length_19532606.14845898" > ../new_input_files/snps/tail_length_snp.bed

bedtools closest -a tail_length_snp.bed -b /scratch/alpine/ericacnr@colostate.edu/GCRF/Reference/ZEFI_annotation/Taeniopygia_guttata_sorted.bed -d > tail_length_distance.bed

```

```{beak.length}
cat GCRF.SNP.imputed_gt.ZFchr.sorted.bed | grep -E "Scaffold_1__1_contigs__length_151072562.39837818" > ../new_input_files/snps/beak_length_snp.bed

bedtools closest -a beak_length_snp.bed -b /scratch/alpine/ericacnr@colostate.edu/GCRF/Reference/ZEFI_annotation/Taeniopygia_guttata_sorted.bed -d > beak_length_distance.bed
```
## Linkage Decay to determine filtering parameteres for genes

Using NGSLD to figure out the linkage disequilibrium. I think ideally I would be running this on a vcf that has been fixed to be at the chromosome level but I'm going to try and dial everything in before I do that.

Okay, so on second thought I'm going to try and new program (https://doi.org/10.1093/bioinformatics/bty875) called PopLDdecay.

```{}
module load gcc/10.3.0
module load zlib

/projects/ericacnr@colostate.edu/PopLDdecay/bin/PopLDdecay -InVCF GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr_gl_impute4.1.vcf.gz -OutStat GCRF.LD.decay.stat
```

Running this is as a script cause I think it's going to take up a lot of memory.

Also going to try using PLINK since that's such a delightful program. Following this tutorial: https://speciationgenomics.github.io/ld_decay/

```{}
mkdir ../ld_decay
VCF=/scratch/alpine/ericacnr@colostate.edu/GCRF/GWAS/impute_new/GCRF.mkdp0_merged_25missing_min.5max.95_depth4_vqsr_gl_impute4.1.vcf.gz
cd ../ld_decay/
```

Going to run this on just the first scaffold, Scaffold 1. Scaffold_1__1_contigs__length_151072562
```{}
plink --vcf $VCF --double-id --allow-extra-chr --allow-no-sex \
--set-missing-var-ids @:# \
--chr Scaffold_1__1_contigs__length_151072562 \
--thin 0.1 -r2 gz --ld-window 100 --ld-window-kb 300 \
--ld-window-r2 0 \
--make-bed --out GCRF_scf1
```
--vcf - specified the location of our VCF file.
--double-id - told plink to duplicate the id of our samples (this is because plink typically expects a family and individual id - i.e. for pedigree data - this is not necessary for us.
--allow-extra-chr - allow additional chromosomes beyond the human chromosome set. This is necessary as otherwise plink expects chromosomes 1-22 and the human X chromosome.
--set-missing-var-ids - also necessary to set a variant ID for our SNPs. Human and model organisms often have annotated SNP names and so plink will look for these. We do not have them so instead we set ours to default to chromosome:position which can be achieved in plink by setting the option @:# - see here for more info.
--chr - an option to choose the chromosome we are running the analysis on. Here we are only performing it on Scaffold 1.
--thin - thin randomly thins out the data - i.e. it randomly retains p proportion of the data. Here we set that to 0.1 or 10%. This means that each site has a 10% probability of remaining in the dataset. This is done to ensure the analysis runs quickly and our output isnt too big as it can quickly get out of hand!
--r2 - finally were on to the options for LD! This tells plink to produce squared correlation coefficients. We also provide the argument gz in order to ensure the output is compressed. This is very important as it is easy to produce EXTREMELY large files.
--ld-window - this allows us to set the size of the lower end of the LD window. In this case, we set it to 100 bp - i.e. any sites with < 100 sites between them are ignored.
--ld-window-kb - this is the upper end of the LD window. Here we set it to 300, meaning that we ignore any two sites more than 300kb apart in the genome.
--ld-window-r2 - the final LD command - this sets a filter on the final output but we want all values of LD to be written out, so we set it to 0.
--make-bed - this just makes a plink bedfile for future analyses
--out Produce the prefix for the output data.

Run the python script...
```{}
acompile
module load python
python ld_decay_calc.py -i GCRF_scf1.ld.gz -o GCRF_scf1
```

```{r}
rm(list = ls())

library(tidyverse)

# set path
my_bins <- "./GCRF-Present/LD_decay/GCRF_scf1.ld_decay_bins"

# read in data
ld_bins <- read_tsv(my_bins)

# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2)) + geom_line() +
  xlab("Distance (bp)") + ylab(expression(italic(r)^2))

```
Repeating this again for scaffold 2 and 3 to see if it's consistent across the long ones but based on this using 100kb as my window, which is great.

```{r}
rm(list = ls())

# set path
my_bins <- "~/Desktop/GCRF/GCRF-Present/LD_decay/GCRF_scf2.ld_decay_bins"

# read in data
ld_bins <- read_tsv(my_bins)

# plot LD decay
ggplot(ld_bins, aes(distance, avg_R2)) + geom_line() +
  xlab("Distance (bp)") + ylab(expression(italic(r)^2))

```

### OLD WINGCHORD STUFF
Now I'm going to try and tackle the wingchord snps...

```{wing.chord}
2	12993716	Scaffold_1__1_contigs__length_151072562.48200618
2	14041130	Scaffold_1__1_contigs__length_151072562.49248032
2	14308333	Scaffold_1__1_contigs__length_151072562.49515235
2	14308348	Scaffold_1__1_contigs__length_151072562.49515250
cat GCRF_wing_ulmm_clean_SNP.txt | awk '{print$1"\t"($2-1)"\t"$2"\t"$3}' > wing_snps.bed
bedtools sort -i wing_snps.bed > wing_snps_sorted.bed

bedtools closest -a wing_snps_sorted.bed -b /scratch/alpine/ericacnr@colostate.edu/GCRF/Reference/ZEFI_annotation/Taeniopygia_guttata_sorted.bed -d > wing_distance.bed
```


## Figuring out what the genes do
### extracting the informations
```{r}
bill_PC <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/genes/bill_distance.bed", sep="\t")
feather_PC <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/genes/feather_distance.bed", sep="\t")
beak_length <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/genes/beak_length_distance.bed", sep="\t")
tail_length <- read.table("~/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/genes/tail_length_distance.bed", sep="\t")
```

```{r, bill_genes}
head(bill_PC)
bill_PC_genes <- bill_PC %>% filter(bill_PC$V12=="gene")

threshold <- 100000

# Filter for SNPs within 100000kb of gene start or stop, or inside the gene
bill_close_genes <- bill_PC_genes %>%
  filter((V2 >= (V6 - threshold) & V2 <= (V6 + threshold)) |
         (V2 >= (V7 - threshold) & V2 <= (V7 + threshold)) |
         (V2 >= V6 & V2 <= V7))
nrow(bill_close_genes)

df <- separate(bill_close_genes, V14, into = paste0("V14_", 1:6), sep = "; ")

# Add column names
colnames(df)[grep("^V14_", colnames(df))] <- c("gene_id", "gene_version", "gene_name", "gene_source", "gene_biotype", "additional_info")

bill_close_genes <- data.frame(bill_close_genes[,1:13], df, bill_close_genes$V15)
head(bill_close_genes)

#bill_pt_coding <- bill_close_genes %>% filter(bill_close_genes$gene_biotype=="gene_biotype protein_coding;")
#nrow(bill_pt_coding)

bill_gene_names <- data.frame(bill_close_genes$V4,bill_close_genes$gene_id, bill_close_genes$gene_name )
write.table(bill_gene_names, "/Users/ericarobertson/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/genes/bill_PC_genes.txt", quote=F)
```
1 Scaffold_2__1_contigs__length_114774641.85189033 gene_id ENSTGUG00000010196 gene_name INPP4A
2 Scaffold_23__1_contigs__length_18116493.1385210 gene_id ENSTGUG00000000148 
3 Scaffold_23__1_contigs__length_18116493.2449928 gene_id ENSTGUG00000000288 gene_name UBTD2
4 Scaffold_1__1_contigs__length_151072562.54132477 gene_id ENSTGUG00000001172 gene_name KIAA1217
5 Scaffold_1__1_contigs__length_151072562.138869993 gene_id ENSTGUG00000009758 gene_name APCDD1
6 Scaffold_5__1_contigs__length_71959612.14244530 gene_id ENSTGUG00000009822 gene_name CPEB2
7 Scaffold_7__1_contigs__length_62267870.31872111 gene_id ENSTGUG00000011762 

Some of these don't have gene names already put in for them.

```{r, feather_genes}
head(feather_PC)
feather_PC_genes <- feather_PC %>% filter(feather_PC$V12=="gene")

threshold <- 100000

# Filter for SNPs within 100000kb of gene start or stop, or inside the gene
feather_close_genes <- feather_PC_genes %>%
  filter((V2 >= (V6 - threshold) & V2 <= (V6 + threshold)) |
         (V2 >= (V7 - threshold) & V2 <= (V7 + threshold)) |
         (V2 >= V6 & V2 <= V7))
nrow(feather_close_genes)

df <- separate(feather_close_genes, V14, into = paste0("V14_", 1:6), sep = "; ")

# Add column names
colnames(df)[grep("^V14_", colnames(df))] <- c("gene_id", "gene_version", "gene_name", "gene_source", "gene_biotype", "additional_info")

feather_close_genes <- data.frame(feather_close_genes[,1:13], df, feather_close_genes$V15)
head(feather_close_genes)

#feather_pt_coding <- feather_close_genes %>% filter(feather_close_genes$gene_biotype=="gene_biotype protein_coding;")
#nrow(feather_pt_coding)

feather_gene_names <- data.frame(feather_close_genes$V4,feather_close_genes$gene_id, feather_close_genes$gene_name)
write.table(feather_gene_names, "/Users/ericarobertson/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/genes/feather_PC_genes.txt", quote = F)
```
1 Scaffold_2__1_contigs__length_114774641.92336930 gene_id ENSTGUG00000008693 gene_name COL4A1
2 Scaffold_2__1_contigs__length_114774641.92336930 gene_id ENSTGUG00000008766 
3 Scaffold_2__1_contigs__length_114774641.80634326 gene_id ENSTGUG00000010798 
4 Scaffold_1__1_contigs__length_151072562.128626182 gene_id ENSTGUG00000008527 gene_name GABBR2
5 Scaffold_3__1_contigs__length_112408489.112330666 gene_id ENSTGUG00000002140 
6 Scaffold_5__1_contigs__length_71959612.65589803 gene_id ENSTGUG00000001929 gene_name QRFPR
7 Scaffold_8__1_contigs__length_35685229.25471973 gene_id ENSTGUG00000006250 gene_name RET

Again, the first one doesn't have a gene name.

```{r, beak_length_genes}
head(beak_length)
beak_length_genes <- beak_length %>% filter(beak_length$V12=="gene")

threshold <- 100000

# Filter for SNPs within 100000kb of gene start or stop, or inside the gene
beak_close_genes <- beak_length_genes %>%
  filter((V2 >= (V6 - threshold) & V2 <= (V6 + threshold)) |
         (V2 >= (V7 - threshold) & V2 <= (V7 + threshold)) |
         (V2 >= V6 & V2 <= V7))
nrow(beak_close_genes)

df <- separate(beak_close_genes, V14, into = paste0("V14_", 1:6), sep = "; ")

# Add column names
colnames(df)[grep("^V14_", colnames(df))] <- c("gene_id", "gene_version", "gene_ID", "gene_source", "gene_biotype", "additional_info")

beak_close_genes <- data.frame(beak_close_genes[,1:13], df, beak_close_genes$V15)
head(beak_close_genes)

#feather_pt_coding <- beak_close_genes %>% filter(feather_close_genes$gene_biotype=="gene_biotype protein_coding;")
#nrow(feather_pt_coding)

beak_gene_names <- data.frame(beak_close_genes$V4,beak_close_genes$gene_id)
beak_gene_names
write.table(beak_gene_names, "/Users/ericarobertson/Desktop/GCRF/GCRF-Present/GenomicPipeline/data/genes/beak_length_genes.txt", quote=F)
```
1 Scaffold_1__1_contigs__length_151072562.39837818 gene_id ENSTGUG00000000374 

### Bill PC Genes
1 Scaffold_2__1_contigs__length_114774641.85189033 gene_id ENSTGUG00000010196 gene_name INPP4A
2 Scaffold_23__1_contigs__length_18116493.1385210 gene_id ENSTGUG00000000148 gene_name ARHGAP26
3 Scaffold_23__1_contigs__length_18116493.2449928 gene_id ENSTGUG00000000288 gene_name UBTD2
4 Scaffold_1__1_contigs__length_151072562.54132477 gene_id ENSTGUG00000001172 gene_name KIAA1217
5 Scaffold_1__1_contigs__length_151072562.138869993 gene_id ENSTGUG00000009758 gene_name APCDD1
6 Scaffold_5__1_contigs__length_71959612.14244530 gene_id ENSTGUG00000009822 gene_name CPEB2
7 Scaffold_7__1_contigs__length_62267870.31872111 gene_id ENSTGUG00000011762 gene_source ensembl

### bill length genes
1 Scaffold_1__1_contigs__length_151072562.39837818 gene_id ENSTGUG00000000374

### Feather PC Genes
1 Scaffold_2__1_contigs__length_114774641.92336930 gene_id ENSTGUG00000008693 gene_name COL4A1
2 Scaffold_2__1_contigs__length_114774641.92336930 gene_id ENSTGUG00000008766 
3 Scaffold_2__1_contigs__length_114774641.80634326 gene_id ENSTGUG00000010798 
4 Scaffold_1__1_contigs__length_151072562.128626182 gene_id ENSTGUG00000008527 gene_name GABBR2
5 Scaffold_3__1_contigs__length_112408489.112330666 gene_id ENSTGUG00000002140 
6 Scaffold_5__1_contigs__length_71959612.65589803 gene_id ENSTGUG00000001929 gene_name QRFPR
7 Scaffold_8__1_contigs__length_35685229.25471973 gene_id ENSTGUG00000006250 gene_name RET

## Allele freq per population

```{}
plink --map GCRF_base_file.map_loc.map --ped GCRF_base_file_fin.ped --out GCRF_af_pop --freqx --family --allow-extra-chr --snp Scaffold_1__1_contigs__length_151072562.138869993
```

So this is just giving me a frequency but I'm not sure which allele it's actually reflecting. So this line of code is the get location 1 allele counts to figure out the frequency and then go back and see what the first line of code is actually showing me. It's the frequency of the least common allele, but it's not always clear which one that is.

```{}
plink --map GCRF_base_file.map_loc.map --ped GCRF_base_file_fin.ped --out GCRF_af_pop1 --keep-fam family1.txt --freqx --allow-extra-chr --snp Scaffold_1__1_contigs__length_151072562.138869993
```

CHR     SNP     A1	A2	C(HOM A1)	C(HET)  C(HOM A2)	C(HAP A1)	C(HAP A2)	C(MISSING)
Scaffold_1__1_contigs__length_151072562 Scaffold_1__1_contigs__length_151072562.138869993	T	C	4	9	30	0	0	0

```{r}
homT <- 4
het <- 9
homC <- 30
total <- 2*4 + 2*9 + 2*30

Tfreq <- ((homT*2)+het)/total
Cfreq <- ((homC*2)+het)/total
Tfreq+Cfreq
```
### beak PC allele frequencies
```{}
plink --map ../GCRF_base_file.map_loc.map --ped ../GCRF_base_file_fin.ped --out GCRF_af_pop_S2.85189033 --freq --family --allow-extra-chr --snp Scaffold_2__1_contigs__length_114774641.85189033 
```
CHR                                                SNP     CLST   A1   A2      MAF    MAC  NCHROBS
S2   Scaffold_2__1_contigs__length_114774641.85189033        1    T    A   0.3605     31       86
S2   Scaffold_2__1_contigs__length_114774641.85189033        2    T    A   0.2717     25       92

```{}
plink --map ../GCRF_base_file.map_loc.map --ped ../GCRF_base_file_fin.ped --out GCRF_af_pop_S2.85189033 --freqx --keep-fam family1.txt --allow-extra-chr --snp Scaffold_2__1_contigs__length_114774641.85189033
```
CHR     SNP     A1      A2      C(HOM A1)     C(HET)      C(HOM A2)       C(HAP A1)       C(HAP A2)       C(MISSING)
S2 S2.85189033   T       A       8              15          20          0       0       0
```{r}
Tfreq <- ((8*2)+15)/(8*2 + 2*15 + 20*2)
Tfreq
```

```{}
plink --map ../GCRF_base_file.map_loc.map --ped ../GCRF_base_file_fin.ped --out GCRF_af_pop.S23.1385210 --freq --family --allow-extra-chr --snp Scaffold_23__1_contigs__length_18116493.1385210
```
CHR                                               SNP     CLST   A1   A2      MAF    MAC  NCHROBS
S23   Scaffold_23__1_contigs__length_18116493.1385210        1    A    G   0.1163     10       86
S23   Scaffold_23__1_contigs__length_18116493.1385210        2    A    G  0.05435      5       92
```{}
plink --map ../GCRF_base_file.map_loc.map --ped ../GCRF_base_file_fin.ped --out GCRF_af_pop.S23.2449928 --freq --family --allow-extra-chr --snp Scaffold_23__1_contigs__length_18116493.2449928
```
CHR                                               SNP     CLST   A1   A2      MAF    MAC  NCHROBS
S23   Scaffold_23__1_contigs__length_18116493.2449928        1    T    C   0.2442     21       86
S23   Scaffold_23__1_contigs__length_18116493.2449928        2    T    C   0.2609     24       92
```{}
plink --map ../GCRF_base_file.map_loc.map --ped ../GCRF_base_file_fin.ped --out GCRF_af_pop.S1.54132477 --freq --family --allow-extra-chr --snp Scaffold_1__1_contigs__length_151072562.54132477
```
 CHR                                                SNP     CLST   A1   A2      MAF    MAC  NCHROBS
S1   Scaffold_1__1_contigs__length_151072562.54132477        1    A    G   0.1977     17       86
S1   Scaffold_1__1_contigs__length_151072562.54132477        2    A    G   0.1196     11       92
```{}
plink --map ../GCRF_base_file.map_loc.map --ped ../GCRF_base_file_fin.ped --out GCRF_af_pop.S3.112330666 --freq --family --allow-extra-chr --snp Scaffold_3__1_contigs__length_112408489.112330666
```
 CHR                                                 SNP     CLST   A1   A2      MAF    MAC  NCHROBS
S3   Scaffold_3__1_contigs__length_112408489.112330666        1    T    C  0.02326      2       86
S3   Scaffold_3__1_contigs__length_112408489.112330666        2    T    C  0.07609      7       92
```{}
plink --map ../GCRF_base_file.map_loc.map --ped ../GCRF_base_file_fin.ped --out GCRF_af_pop.S5.65589803 --freq --family --allow-extra-chr --snp Scaffold_5__1_contigs__length_71959612.65589803
```
 CHR                                               SNP     CLST   A1   A2      MAF    MAC  NCHROBS
S5   Scaffold_5__1_contigs__length_71959612.65589803        1    G    A   0.2442     21       86
S5   Scaffold_5__1_contigs__length_71959612.65589803        2    G    A   0.3804     35       92
```{}
plink --map ../GCRF_base_file.map_loc.map --ped ../GCRF_base_file_fin.ped --out GCRF_af_pop.S7.31872111 --freq --family --allow-extra-chr --snp Scaffold_7__1_contigs__length_62267870.31872111
```
CHR                                               SNP     CLST   A1   A2      MAF    MAC  NCHROBS
S7   Scaffold_7__1_contigs__length_62267870.31872111        1    A    G   0.1512     13       86
S7   Scaffold_7__1_contigs__length_62267870.31872111        2    A    G   0.1413     13       92


# FINAL FIGURES
## ARHGAP26 Gene
Rho GAP family also showed up in Matt's paper

Scaffold_23__1_contigs__length_18116493	1385210	.	G	A	.	PASS	.	GT	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	1|1	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|1	0|0	1|0	0|0	1|0	0|0	0|0	1|1	1|1	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|1	0|0	0|0	0|0	0|0	0|1	0|0	0|0	0|0	0|0	0|0	1|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	1|1	0|0	0|0	0|0	0|0	0|0	1|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0
```{r}
ARHGAP26 <- bill.results_list[[7]]
y_breaks <- pretty(range(ARHGAP26$PHENOTYPE), n=5)
x_breaks <- seq(min(ARHGAP26$GENOTYPE), max(ARHGAP26$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")
my_title <- expression(paste("Bill Morphology vs ", italic("ARHGAP26")))

ARHGAP26.plot <- ggplot(ARHGAP26, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle(my_title) +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("GG", "GA", "AA")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()
plot(ARHGAP26.plot)
ggsave("ARHGAP26_plot.png", plot=ARHGAP26.plot, path = "~/Desktop/GCRF/GCRF-Present/figures/", width=9, height=7)
```

## APCDDI Gene
Expressed in human hair follicles
Also found in Gamboa 2021

Scaffold_1__1_contigs__length_151072562	138869993	.	C	T	.	PASS	.	GT	0|0	0|0	0|0	0|0	0|0	1|1	0|1	0|0	0|0	1|1	0|0	0|0	0|0	0|1	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	1|0	1|0	0|0	1|0	0|0	1|1	0|0	0|0	0|0	1|1	0|0	0|0	1|0	0|0	0|1	0|0	0|0	0|0	0|0	0|0	0|1	0|0	0|0	0|0	0|0	0|1	0|0	0|0	0|0	0|0	0|0	1|0	0|0	0|1	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|0	0|1	1|0	0|0	1|0	0|0	0|0	0|0	0|0	1|0	0|1	0|0	0|0	0|0	0|1	0|0	0|1	0|0	1|0	1|0	0|0	1|1	1|1	0|0	0|0

```{r}
APCDDI <- bill.results_list[[2]]
y_breaks <- pretty(range(APCDDI$PHENOTYPE), n=5)
x_breaks <- seq(min(APCDDI$GENOTYPE), max(APCDDI$GENOTYPE), length.out = 3)
group_colors <- c("1" = "lightblue3", "2" = "tomato2")
my_title <- expression(paste("Bill Morphology vs ", italic("APCDDI")))

APCDDI.plot <- ggplot(APCDDI, aes(x = GENOTYPE, y = PHENOTYPE, color = FID)) +
  geom_point() +   # Add data points with color aesthetic
  geom_smooth(method = "lm", se = TRUE, color = "grey3") +  # Add smoothed trend line with confidence intervals
  labs(x = "Genotype", y = "Feather Morphology PC1") +  # Add axis labels
  ggtitle(my_title) +
  scale_y_continuous(breaks = y_breaks) +
  scale_x_continuous(breaks = x_breaks, labels = c("CC", "CT", "TT")) +
  scale_color_manual(labels = c("WMNT", "PIPA"), values=group_colors) +  # Customize legend labels
  theme_classic()
plot(ARHGAP26.plot)
ggsave("APCDDI_plot.png", plot=APCDDI.plot, path = "~/Desktop/GCRF/GCRF-Present/figures/", width=9, height=7)
```



