---
title: "GCRF PCA"
output: html_document
date: "2023-12-12"
---
```{r include=TRUE}
knitr::opts_chunk$set(include=TRUE)
```

```{r include=TRUE, echo=FALSE, message=FALSE}
library(MASS)
library(caret)
library(ggplot2)
library(dplyr)
library(FactoMineR)
library(factoextra)
library(corrplot)
library(ggcorrplot)
library(rstatix)
```

# MORPHOLOGY PCA (Undivided)
```{r}
morpho <- read.csv("~/Desktop/GCRF/R data/GCRF_morphology.csv")
morpho <- morpho[1:189,]

morpho_num <- as.data.frame(cbind(as.numeric(morpho$wing), as.numeric(morpho$tarsus), as.numeric(morpho$weight), as.numeric(morpho$tail_length), as.numeric(morpho$nare_length), as.numeric(morpho$bristle_length), as.numeric(morpho$beak_depth), as.numeric(morpho$beak_width), as.numeric(morpho$nare_end_length), as.numeric(morpho$culmen_end_length)))

colnames(morpho_num) <- c("wing", "tarsus", "weight", "tail_length", "nare_length", "bristle_length", "beak_depth", "beak_width", "nare_end_length", "culmen_end_length")
morpho2 <- as.data.frame(cbind(morpho$BGP_ID, as.factor(morpho$site),morpho_num))
colnames(morpho2) <- c("BGP_ID", "site", "wing", "tarsus", "weight", "tail_length", "nare_length", "bristle_length", "beak_depth", "beak_width", "nare_end_length", "culmen_end_length")

morpho_avg <- morpho2 %>% group_by(BGP_ID, site) %>% reframe(wing=mean(wing), tarsus=mean(tarsus), weight=mean(weight), tail_length=mean(tail_length), nare_length=mean(nare_length), bristle_length=mean(bristle_length), beak_depth=(beak_depth), beak_width=mean(beak_width), nare_end_length=mean(nare_end_length), culmen_end_length=mean(culmen_end_length))

morpho_avg <- morpho_avg[!duplicated(morpho_avg$BGP_ID),]

morpho_avg <- na.omit(morpho_avg)
colSums(is.na(morpho_avg))
morpho_num <- morpho_avg[,3:12]



corr_m <- cor(morpho_num)
ggcorrplot(corr_m)

```

```{r}
morpho_avg #the df with individuals (character) and groups (factor)
morpho_num #the df with just the variables
#nothing has been normalized

#taking the active variable, ie individuals and the measurements that go into the PCA
morpho.active <- morpho_avg[,c(1, 3:11)]

#now need to standardize the variables, but using the PCA function that's already done so don't need to do it prior to running the PCA in this case, otherwise use scale()... x-mean(x)/sd(x)
#need to remove the column for individual though...
morpho.active.1 <- morpho.active[,-1]
morpho.pca <- PCA(morpho.active.1,  scale.unit=T, graph=F)

eig.val <- get_eig(morpho.pca)
eig.val
fviz_eig(morpho.pca, addlabels = T)
```

```{r}
var <- get_pca_var(morpho.pca)
head(var$coord)

```

```{r}
fviz_pca_var(morpho.pca, col.var="black")
fviz_cos2(morpho.pca, choice="var", axes=1:4)
library("corrplot")
corrplot(var$contrib, is.corr=F)
```

```{r}
fviz_pca_ind(morpho.pca)
fviz_pca_ind(morpho.pca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )
```

```{r}
fviz_pca_ind(morpho.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = morpho_avg$site, # color by groups
             palette = c("#00AFBB", "#E7B800"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups"
             )
```

# LINEAR REGRESSIONS WITH BODY SIZE
To get residual values that will be corrected for body size and can be used in the PCA

```{r}
morp.sex.og <- read.csv("~/Desktop/GCRF/R data/GCRF_morpho_sex.csv")
#morp.sex$site, morp.sex$Sex, morp.sex$BGP_ID,
morp.sex <- as.data.frame(cbind(as.numeric(morp.sex.og$wing), as.numeric(morp.sex.og$tarsus), as.numeric(morp.sex.og$weight), as.numeric(morp.sex.og$tail_length), as.numeric(morp.sex.og$nare_length), as.numeric(morp.sex.og$bristle_length), as.numeric(morp.sex.og$beak_depth), as.numeric(morp.sex.og$beak_width), as.numeric(morp.sex.og$culmen_end_length)))
morp.sex <- as.data.frame(cbind(morp.sex.og$site, morp.sex.og$Sex, morp.sex.og$BGP_ID, morp.sex))
colnames(morp.sex) <- c("site", "sex", "BGP_ID", "wing", "tarsus", "weight", "tail_length", "nare_length", "bristle_length", "beak_depth", "beak_width", "culmen_end_length")
females <- na.omit(morp.sex %>% filter(sex=="F"))
females <- females %>% group_by(BGP_ID, site) %>% reframe(wing=mean(wing), tarsus=mean(tarsus), weight=mean(weight), tail_length=mean(tail_length), nare_length=mean(nare_length), bristle_length=mean(bristle_length), beak_depth=(beak_depth), beak_width=mean(beak_width), culmen_end_length=mean(culmen_end_length))

wing.f <- na.omit(as.data.frame(cbind(females$site, females$BGP_ID, females$wing, females$tarsus)))
colnames(wing.f) <- c("site", "BGP_ID", "wing", "tarsus")
tail.f <- na.omit(as.data.frame(cbind(females$site, females$BGP_ID, females$tail_length, females$tarsus)))
colnames(tail.f) <- c("site", "BGP_ID", "tail", "tarsus")
nare.f <- na.omit(as.data.frame(cbind(females$site, females$BGP_ID, females$nare_length, females$tarsus)))
colnames(nare.f) <- c("site", "BGP_ID", "nare", "tarsus")
bristle_l.f <- na.omit(as.data.frame(cbind(females$site, females$BGP_ID, females$bristle_length, females$tarsus)))
colnames(bristle_l.f) <- c("site", "BGP_ID", "bristle_length", "tarsus")
bill_l.f <- na.omit(as.data.frame(cbind(females$site, females$BGP_ID, females$culmen_end_length, females$tarsus)))
colnames(bill_l.f) <- c("site", "BGP_ID", "bill_length", "tarsus")
bill_d.f <- na.omit(as.data.frame(cbind(females$site, females$BGP_ID, females$beak_depth, females$tarsus)))
colnames(bill_d.f) <- c("site", "BGP_ID", "bill_depth", "tarsus")
bill_w.f <- na.omit(as.data.frame(cbind(females$site, females$BGP_ID, females$beak_width, females$tarsus)))
colnames(bill_w.f) <- c("site", "BGP_ID", "bill_width", "tarsus")
```

## Female linear models
```{r}
wing.f.lm <- lm(as.numeric(wing) ~ as.numeric(tarsus), data=wing.f)
summary(wing.f.lm)
ggplot(wing.f, aes(x=as.numeric(wing), y=as.numeric(tarsus)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(wing.f.lm$residuals)
wing.f.res <- wing.f.lm$residuals

tail.f.lm <- lm(as.numeric(tail) ~ as.numeric(tarsus), data=tail.f)
summary(tail.f.lm)
ggplot(tail.f, aes(x=as.numeric(tail), y=as.numeric(tarsus)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(tail.f.lm$residuals)
tail.f.res <- tail.f.lm$residuals

nare.f.lm <- lm(as.numeric(nare) ~ as.numeric(tarsus), data=nare.f)
summary(nare.f.lm)
ggplot(nare.f, aes(x=as.numeric(tarsus), y=as.numeric(nare)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(nare.f.lm$residuals)
nare.f.res <- nare.f.lm$residuals

bristle_l.f.lm <- lm(as.numeric(bristle_length) ~ as.numeric(tarsus), data=bristle_l.f)
summary(bristle_l.f.lm)
ggplot(bristle_l.f, aes(x=as.numeric(tarsus), y=as.numeric(bristle_length)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bristle_l.f.lm$residuals)
bristle_l.f.res <- bristle_l.f.lm$residuals

bill_l.f.lm <- lm(as.numeric(bill_length) ~ as.numeric(tarsus), data=bill_l.f)
summary(bill_l.f.lm)
ggplot(bill_l.f, aes(x=as.numeric(tarsus), y=as.numeric(bill_length)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_l.f.lm$residuals)
bill_l.f.res <- bill_l.f.lm$residuals

bill_d.f.lm <- lm(as.numeric(bill_depth) ~ as.numeric(tarsus), data=bill_d.f)
summary(bill_d.f.lm)
ggplot(bill_d.f, aes(x=as.numeric(tarsus), y=as.numeric(bill_depth)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_d.f.lm$residuals)
bill_d.f.res <- bill_d.f.lm$residuals

bill_w.f.lm <- lm(as.numeric(bill_width) ~ as.numeric(tarsus), data=bill_w.f)
summary(bill_w.f.lm)
ggplot(bill_w.f, aes(x=as.numeric(tarsus), y=as.numeric(bill_width)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_w.f.lm$residuals)
bill_w.f.res <- bill_w.f.lm$residuals

female.res <- as.data.frame(cbind(as.numeric(wing.f.res), as.numeric(tail.f.res), as.numeric(nare.f.res), as.numeric(bristle_l.f.res), as.numeric(bill_l.f.res), as.numeric(bill_d.f.res), as.numeric(bill_w.f.res), as.numeric(wing.f$tarsus)))
colnames(female.res) <- c("wing", "tail", "nare", "bristle_lenght", "bill_lenght", "bill_depth", "bill_width", "tarsus")
```

## Female PCA
```{r}
female.pca <- PCA(female.res,  scale.unit=T, graph=F)

eig.val.f <- get_eig(female.pca)
eig.val.f
fviz_eig(female.pca, addlabels = T)
```

```{r}
var.f <- get_pca_var(female.pca)
head(var.f$coord)

```

```{r}
fviz_pca_var(female.pca, col.var="black")
fviz_cos2(female.pca, choice="var", axes=1:2)

corrplot(var.f$contrib, is.corr=F)
```

```{r}
fviz_pca_ind(female.pca)
fviz_pca_ind(female.pca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping (slow if many points)
             )
```

```{r}
fviz_pca_ind(female.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = females$site, # color by groups
             palette = c("#00AFBB", "#E7B800"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups"
             )
```

## Male linear models
```{r}
males <- na.omit(morp.sex %>% filter(sex=="M"))
males <- males %>% group_by(BGP_ID, site) %>% reframe(wing=mean(wing), tarsus=mean(tarsus), weight=mean(weight), tail_length=mean(tail_length), nare_length=mean(nare_length), bristle_length=mean(bristle_length), beak_depth=(beak_depth), beak_width=mean(beak_width), culmen_end_length=mean(culmen_end_length))

wing.m <- na.omit(as.data.frame(cbind(males$site, males$BGP_ID, males$wing, males$tarsus)))
colnames(wing.m) <- c("site", "BGP_ID", "wing", "tarsus")
tail.m<- na.omit(as.data.frame(cbind(males$site, males$BGP_ID, males$tail_length, males$tarsus)))
colnames(tail.m) <- c("site", "BGP_ID", "tail", "tarsus")
nare.m<- na.omit(as.data.frame(cbind(males$site, males$BGP_ID, males$nare_length, males$tarsus)))
colnames(nare.m) <- c("site", "BGP_ID", "nare", "tarsus")
bristle_l.m <- na.omit(as.data.frame(cbind(males$site, males$BGP_ID, males$bristle_length, males$tarsus)))
colnames(bristle_l.m) <- c("site", "BGP_ID", "bristle_length", "tarsus")
bill_l.m <- na.omit(as.data.frame(cbind(males$site, males$BGP_ID, males$culmen_end_length, males$tarsus)))
colnames(bill_l.m) <- c("site", "BGP_ID", "bill_length", "tarsus")
bill_d.m <- na.omit(as.data.frame(cbind(males$site, males$BGP_ID, males$beak_depth, males$tarsus)))
colnames(bill_d.m) <- c("site", "BGP_ID", "bill_depth", "tarsus")
bill_w.m <- na.omit(as.data.frame(cbind(males$site, males$BGP_ID, males$beak_width, males$tarsus)))
colnames(bill_w.m) <- c("site", "BGP_ID", "bill_width", "tarsus")
```

```{r}
wing.m.lm <- lm(as.numeric(wing) ~ as.numeric(tarsus), data=wing.m)
summary(wing.m.lm)
ggplot(wing.f, aes(x=as.numeric(wing), y=as.numeric(tarsus)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(wing.m.lm$residuals)
wing.m.res <- wing.m.lm$residuals

tail.m.lm <- lm(as.numeric(tail) ~ as.numeric(tarsus), data=tail.m)
summary(tail.m.lm)
ggplot(tail.f, aes(x=as.numeric(tail), y=as.numeric(tarsus)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(tail.m.lm$residuals)
tail.m.res <- tail.m.lm$residuals

nare.m.lm <- lm(as.numeric(nare) ~ as.numeric(tarsus), data=nare.m)
summary(nare.m.lm)
ggplot(nare.f, aes(x=as.numeric(tarsus), y=as.numeric(nare)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(nare.m.lm$residuals)
nare.m.res <- nare.m.lm$residuals

bristle_l.m.lm <- lm(as.numeric(bristle_length) ~ as.numeric(tarsus), data=bristle_l.m)
summary(bristle_l.m.lm)
ggplot(bristle_l.f, aes(x=as.numeric(tarsus), y=as.numeric(bristle_length)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bristle_l.m.lm$residuals)
bristle_l.m.res <- bristle_l.m.lm$residuals

bill_l.m.lm <- lm(as.numeric(bill_length) ~ as.numeric(tarsus), data=bill_l.m)
summary(bill_l.m.lm)
ggplot(bill_l.f, aes(x=as.numeric(tarsus), y=as.numeric(bill_length)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_l.m.lm$residuals)
bill_l.m.res <- bill_l.m.lm$residuals

bill_d.m.lm <- lm(as.numeric(bill_depth) ~ as.numeric(tarsus), data=bill_d.m)
summary(bill_d.m.lm)
ggplot(bill_d.f, aes(x=as.numeric(tarsus), y=as.numeric(bill_depth)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_d.m.lm$residuals)
bill_d.m.res <- bill_d.m.lm$residuals

bill_w.m.lm <- lm(as.numeric(bill_width) ~ as.numeric(tarsus), data=bill_w.m)
summary(bill_w.m.lm)
ggplot(bill_w.f, aes(x=as.numeric(tarsus), y=as.numeric(bill_width)))+
  geom_point()+
  geom_smooth(method=lm)+
  theme_classic()
plot(bill_w.m.lm$residuals)
bill_w.m.res <- bill_w.m.lm$residuals

male.res <- as.data.frame(cbind(as.numeric(wing.m.res), as.numeric(tail.m.res), as.numeric(nare.m.res), as.numeric(bristle_l.m.res), as.numeric(bill_l.m.res), as.numeric(bill_d.m.res), as.numeric(bill_w.m.res), as.numeric(wing.f$tarsus)))
colnames(male.res) <- c("wing", "tail", "nare", "bristle_lenght", "bill_lenght", "bill_depth", "bill_width", "tarsus")
```

## Male PCA
```{r}
male.pca <- PCA(male.res,  scale.unit=T, graph=F)

eig.val.f <- get_eig(male.pca)
eig.val.f
fviz_eig(male.pca, addlabels = T)
```

```{r}
fviz_pca_ind(male.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = males$site, # color by groups
             palette = c("#00AFBB", "#E7B800"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups",
             title= "Males PCA"
             )

```

# T TESTS USING THE RESIDUALS
## Males
```{r m.wing}
male.morpho <- data.frame(males$site, male.res)
colnames(male.morpho) <- c("site", "wing", "tail", "nare", "bristle_lenght", "bill_lenght", "bill_depth", "bill_width", "tarsus")


qqnorm(male.morpho$wing)
qqline(male.morpho$wing)

shapiro.test(male.morpho$wing)

#p=0.4597
t.test(wing ~ site, data=male.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=wing), data=male.morpho)

ggplot(male.morpho, aes(x=site, y=wing))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")
```

```{r m.tail}
qqnorm(male.morpho$tail)
qqline(male.morpho$tail)

shapiro.test(male.morpho$tail)

#p=0.301
t.test(tail ~ site, data=male.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=tail), data=male.morpho)

ggplot(male.morpho, aes(x=site, y=tail))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r m.nare}
qqnorm(male.morpho$nare)
qqline(male.morpho$nare)

shapiro.test(male.morpho$nare)

#p=0.301
t.test(nare ~ site, data=male.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=nare), data=male.morpho)

ggplot(male.morpho, aes(x=site, y=nare))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r m.tarsus}
qqnorm(male.morpho$tarsus)
qqline(male.morpho$tarsus)

shapiro.test(male.morpho$tarsus)

#p=0.4919
t.test(tarsus ~ site, data=male.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=tail), data=male.morpho)

ggplot(male.morpho, aes(x=site, y=tail))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r m.bristle_length}
qqnorm(male.morpho$bristle_lenght)
qqline(male.morpho$bristle_lenght)

shapiro.test(male.morpho$bristle_lenght)

#p=0.3299
wilcox.test(bristle_lenght ~ site, data=male.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=bristle_lenght), data=male.morpho)

ggplot(male.morpho, aes(x=site, y=bristle_lenght))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r m.bill_lenght}
qqnorm(male.morpho$bill_lenght)
qqline(male.morpho$bill_lenght)

shapiro.test(male.morpho$bill_lenght)

#p=0.8286
t.test(bill_lenght ~ site, data=male.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=bill_lenght), data=male.morpho)

ggplot(male.morpho, aes(x=site, y=bill_lenght))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r m.bill_depth}
qqnorm(male.morpho$bill_depth)
qqline(male.morpho$bill_depth)

shapiro.test(male.morpho$bill_depth)

#p=0.4625
wilcox.test(bill_depth ~ site, data=male.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=bill_depth), data=male.morpho)

ggplot(male.morpho, aes(x=site, y=bill_depth))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r m.bill_width}
qqnorm(male.morpho$bill_width)
qqline(male.morpho$bill_width)

shapiro.test(male.morpho$bill_width)

#p=0.08037
t.test(bill_width ~ site, data=male.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=bill_width), data=male.morpho)

ggplot(male.morpho, aes(x=site, y=bill_width))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

## Females
```{r f.wing}
female.morpho <- data.frame(females$site, female.res)
colnames(female.morpho) <- c("site", "wing", "tail", "nare", "bristle_lenght", "bill_lenght", "bill_depth", "bill_width", "tarsus")


qqnorm(female.morpho$wing)
qqline(female.morpho$wing)

shapiro.test(female.morpho$wing)

#p=0.001526
t.test(wing ~ site, data=female.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=wing), data=female.morpho)

ggplot(female.morpho, aes(x=site, y=wing))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")
```

```{r f.tail}
qqnorm(female.morpho$tail)
qqline(female.morpho$tail)

shapiro.test(female.morpho$tail)

#p=0.3723
t.test(tail ~ site, data=female.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=tail), data=female.morpho)

ggplot(female.morpho, aes(x=site, y=tail))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r f.nare*}
qqnorm(female.morpho$nare)
qqline(female.morpho$nare)

shapiro.test(female.morpho$nare)

#p=0.01998
t.test(nare ~ site, data=female.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=nare), data=female.morpho)

ggplot(female.morpho, aes(x=site, y=nare))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r f.tarsus}
qqnorm(female.morpho$tarsus)
qqline(female.morpho$tarsus)

shapiro.test(female.morpho$tarsus)

#p=0.8496
t.test(tarsus ~ site, data=female.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=tail), data=female.morpho)

ggplot(female.morpho, aes(x=site, y=tail))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r f.bristle_length}
qqnorm(female.morpho$bristle_lenght)
qqline(female.morpho$bristle_lenght)

shapiro.test(female.morpho$bristle_lenght)

#p=0.6501
wilcox.test(bristle_lenght ~ site, data=female.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=bristle_lenght), data=female.morpho)

ggplot(female.morpho, aes(x=site, y=bristle_lenght))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r f.bill_lenght}
qqnorm(female.morpho$bill_lenght)
qqline(female.morpho$bill_lenght)

shapiro.test(female.morpho$bill_lenght)

#p=0.1041
t.test(bill_lenght ~ site, data=female.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=bill_lenght), data=female.morpho)

ggplot(female.morpho, aes(x=site, y=bill_lenght))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r f.bill_depth}
qqnorm(female.morpho$bill_depth)
qqline(female.morpho$bill_depth)

shapiro.test(female.morpho$bill_depth)

#p=0.9671
wilcox.test(bill_depth ~ site, data=female.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=bill_depth), data=female.morpho)

ggplot(female.morpho, aes(x=site, y=bill_depth))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```

```{r f.bill_width}
qqnorm(female.morpho$bill_width)
qqline(female.morpho$bill_width)

shapiro.test(female.morpho$bill_width)

#p=0.7416
wilcox.test(bill_width ~ site, data=female.morpho)
ggplot()+
  geom_boxplot(aes(x=site, y=bill_width), data=female.morpho)

ggplot(female.morpho, aes(x=site, y=bill_width))+
  geom_jitter()+
  theme_classic()+
  stat_summary(fun.y=mean, geom="point", shape=18,
                 size=3, color="red")

```


# FEATHER MORPHOLOGY
So the goal here is to get residuals that I can plug into the GWAS later on.

```{r}
feather <- read.csv("~/Desktop/GCRF/R data/GCRF_feather_morpho.csv")
site <- data.frame(morpho_avg$site, morpho_avg$BGP_ID, morpho_avg$tarsus)
colnames(site) <- c("site", "ID", "tarsus")

feather_avg <- feather %>% group_by(ID) %>% summarize(pen_length=mean(Penn_length), pen_num=mean(Penn_number), plum_length=mean(Plum_length), plum_number=mean(Plum_number), nodes=mean(Nodes))

feather_avg <- merge(feather_avg, site, by="ID")

```

```{r}
feather_res <- data.frame(ID = feather_avg$ID)  # Assuming ID is a common column

# Function to fit linear models and extract residuals
fit_and_get_residuals <- function(data, predictor_col, response_col) {
  lm_model <- lm(data[[response_col]] ~ data[[predictor_col]], data = feather_avg)
  residuals <- residuals(lm_model)
  return(residuals)
}

# Loop over each variable (X1, X2, X3, X4, X5)
for (col_name in colnames(feather_avg)[2:6]) {
  residuals <- fit_and_get_residuals(feather_avg, col_name, "tarsus")
  feather_res[[col_name]] <- residuals
}

# Print the new dataframe with residuals
print(feather_res)

```
```{r}
feather_res <- rbind(feather_res, 
                     c("23N00431", NA, NA, NA, NA, NA), 
                     c("23N00434", NA, NA, NA, NA, NA), 
                     c("23N00436", NA, NA, NA, NA, NA), 
                     c("23N00437", NA, NA, NA, NA, NA), 
                     c("23N00444", NA, NA, NA, NA, NA),  # Removed extra comma
                     c("23N00447", NA, NA, NA, NA, NA), 
                     c("23N00448", NA, NA, NA, NA, NA), 
                     c("23N00450", NA, NA, NA, NA, NA), 
                     c("23N00451", NA, NA, NA, NA, NA), 
                     c("23N00452", NA, NA, NA, NA, NA), 
                     c("23N00453", NA, NA, NA, NA, NA), 
                     c("23N00454", NA, NA, NA, NA, NA), 
                     c("23N00456", NA, NA, NA, NA, NA), 
                     c("23N00457", NA, NA, NA, NA, NA), 
                     c("23N00459", NA, NA, NA, NA, NA), 
                     c("23N00460", NA, NA, NA, NA, NA), 
                     c("23N00463", NA, NA, NA, NA, NA), 
                     c("23N00464", NA, NA, NA, NA, NA), 
                     c("23N00467", NA, NA, NA, NA, NA), 
                     c("23N00470", NA, NA, NA, NA, NA),  # Removed extra comma
                     c("23N00475", NA, NA, NA, NA, NA), 
                     c("23N0079", NA, NA, NA, NA, NA), 
                     c("23N00480", NA, NA, NA, NA, NA), 
                     c("23N00483", NA, NA, NA, NA, NA), 
                     c("23N00485", NA, NA, NA, NA, NA), 
                     c("23N00488", NA, NA, NA, NA, NA), 
                     c("23N00489", NA, NA, NA, NA, NA), 
                     c("23N00490", NA, NA, NA, NA, NA), 
                     c("23N00491", NA, NA, NA, NA, NA),  # Removed extra comma
                     c("23N00557", NA, NA, NA, NA, NA))
write.table(feather_res, file="~/Desktop/GCRF/GenomicPipeline/GWAS input/feather_residuals.txt", row.names=F, col.names=T)
```


Now my goal is to run a PCA with all of the feather stuff.

